---
title: "OysterEggMorphology"
author: "Camila Cortina"
date: "1/16/2019 updated by Elise 2/6/20"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Load libraries
```{r include=FALSE}
library(lmerTest)
library(car)
library(onewaytests)
library(e1071)
library(vioplot)
library(ggplot2)


```

#Load data
Name the data you export from FileMaker Pro by their exact table names and save them as CSVs, e.g. Egg_Morphology.csv
###This chunk of code gets the data and defines new categories, based on the Image number name, necessary for analysis such as the label, environment, and block 
```{r}
setwd("~/GitHub/LarvaeTransGen2018/data")
egg_data<- read.csv("Egg_Morphology.csv")
head(egg_data)

#Makes the FemaleIDs a string to use the strsplit function to get CrossType
egg_data$ImageID<- as.character(egg_data$ImageID)
egg_data$FemaleID<- as.character(egg_data$FemaleID)
#assign each row CrossType using the strsplit function on the FemaleID column
egg_data$CrossType<- sapply(strsplit(egg_data$FemaleID, "F"), '[',1)

#spilts the FemaleID string to get the block number or B1 or B2 or B3
egg_data$Block<- sapply(strsplit(egg_data$FemaleID, "_"), '[',2)
#changing block to a factor to arrange boxplots, etc.
egg_data$Block<-as.factor(egg_data$Block)

#rearranging the order of the strings so that block is first, to display all blocks together on a graph
egg_data$FemaleID<- paste(sapply(strsplit(egg_data$ImageID, "_"), '[',2),sapply(strsplit(egg_data$ImageID, "_"), '[',1),sep="_")

#look at the structure of the data
str(egg_data)

#make Female ID a factor again
egg_data$FemaleID<- as.factor(egg_data$FemaleID)

```


###Subsetting the data
```{r}
#subsetting egg_data to not have CF06_B2 because it only had 8 eggs total and they were all small ones that may have been excluded otherwise
#exclude variables CF06_B2
sub.eggs<- subset(egg_data, FemaleID != "B2_CF06", drop = TRUE)
sub.eggs$FemaleID<- as.factor(sub.eggs$FemaleID) #relevel FemaleID in sub.eggs to get rid of CF06_B2 from the dataset

#subsetting the eggs so that any area 100000 pixels or below is excluded, this takes out 46 data points out of 1314 total 
#unclear if this subset was done simply to fix the data skew or to remove underdeveloped eggs
sub.eggs2<- subset(sub.eggs, AreaPix2 >= 100000, drop = TRUE)

#subsetting to take out CF04_B3 because 26% of their eggs were taken out of the set when excluding eggs with an area under 100,000 pixels # Elise found that B1_EF07 had only 11 eggs left, but it was still kept in and B3_CF04 wasn't why? 
sub.eggs3<- subset(sub.eggs2, FemaleID != "B3_CF04", drop = TRUE) #is this really necessary? Elise doesn't think it is because there are still 22 eggs for that female, but only 11 eggs for EF07_B1, which we keep in
sub.eggs3$FemaleID<- as.factor(sub.eggs3$FemaleID) #relevel FemaleID in sub.eggs3 to get rid of CF06_B2 from the dataset, but this didn't work
Area.female.plot3<- boxplot(AreauM2~FemaleID, data = sub.eggs3, las = 2, xlab = "Female ID", ylab = "Area in uM2", title = "Area of eggs by female", cex.axis = .55)

diameter.female.plot3<- boxplot(DiameteruM~FemaleID, data = sub.eggs3, las = 2, xlab = "female ID", ylab = "Diameter in micrometers", main = "Diameter of eggs by female", cex.axis = .55)

```

###Everything in this chunk is for the final subsetted data (the dataset excluding eggs with an area below 100,000 and all of CF04_B3)
```{r}
#a linear mixed model incorporating CrossType (parent treatment) and block as fixed effects and and female ID as a random effect
egg.model.sub3<- lmer(AreauM2~CrossType + Block + (1|FemaleID), data = sub.eggs3, REML = FALSE)
summary(egg.model.sub3)

#The same model but without CrossType, so that we can compare the two models to see if the environment had a significant effect on the size
egg.model.sub3.woenvr<- lmer(AreauM2~Block + (1|FemaleID), data = sub.eggs3, REML = FALSE)
summary(egg.model.sub3.woenvr)

#The same model, but without the block to see how much of an effect block had on the size
egg.model.sub3.woblock<- lmer(AreauM2~CrossType + (1|FemaleID), data = sub.eggs3, REML = FALSE)
summary(egg.model.sub3.woblock)

#running an anova to compare the first two models - will tell us if the parental environment had a significant effect. 
anova(egg.model.sub3, egg.model.sub3.woenvr)
#Pr(>Chisq) = .4972, the environment does not have a significant effect on the egg size 

#running an anova to compare the first model with the model without block to tell us if block had a significant effect on the egg size
anova(egg.model.sub3, egg.model.sub3.woblock)
#Pr(>Chisq) = .137, so the blocks did not have a significant effect on the egg size after removing CF04_B3. When CF04_B3 was included (see chunk below) there was a slightly significant effect due to the block Pr(>Chisq) = .0342 

#plotting the residuals
par(mfrow = c(1,1))
plot(egg.model.sub3, main = "Residuals plot for the model")
#testing normality of the residuals
qqnorm(residuals(egg.model.sub3))
hist(residuals(egg.model.sub3), breaks = 50)
#residuals of the model are not normally distributed 
shapiro.test(residuals(egg.model.sub3))

#a linear model with female ID as another fixed effect - is this what you wanted to compare females?
linear.model.sub.eggs3<- lm(AreauM2 ~ CrossType + FemaleID, data = sub.eggs3)
summary(linear.model.sub.eggs3)
#gives some instances of significance by female ID, however it said that one of them could not be compared due to signularities, which does not really make sense to me..#Elise did not get anything about singularties when she ran it
#with a one way anova and the linear model from above both the environment and female are significant
anova(linear.model.sub.eggs3)#Elise got a significant effect of female

#Linear model to compare environment, block, and female?
egglm3<- lm(AreauM2~CrossType + Block + FemaleID, data = sub.eggs3)
summary(egglm3)
par(mfrow=c(2,2))
plot(egglm3)
#I am not sure why 3 of the females were not defined due to singularities...#Elise did not get this error

#not sure what to do for the TukeyHSD because it gives me an error saying that TukeyHSD cannot be applied to class lmerMod or lm
#TukeyHSD(linear.model.sub.eggs3)
#TukeyHSD(egg.model.sub3)
aov(linear.model.sub.eggs3)
summary(aov(AreauM2~FemaleID + CrossType, data = sub.eggs3))
par(mfrow = c(2,2))
plot(aov(AreauM2~FemaleID+CrossType, data = sub.eggs3))
#TukeyHSD(aov(AreauM2~FemaleID + CrossType, data = sub.eggs3))
par(mfrow = c(1,1))

#testing homogeneity of the data - the variances of eggs for each female are NOT the same
bartlett.test(AreauM2 ~ FemaleID, data = sub.eggs3)

#Ran a second test with the levene test because the data are left skewed, still indicating that the variances are significantly not equal. 
leveneTest(AreauM2 ~ FemaleID, data = sub.eggs3)

#testing normality of the raw data
qqnorm(sub.eggs3$AreauM2)
qqline(sub.eggs3$AreauM2)
#squaring it seems to make it a little more normal # elise disagrees, does not look more normal by squaring, looks worse
qqnorm((sub.eggs3$AreauM2)^2)
qqline((sub.eggs3$AreauM2)^2)

#trying a welch's anova because the data was not homogeneous 
welch.test(AreauM2 ~ FemaleID, data = sub.eggs3)

```


###square-transforming the orignial data seems to offset most of the abnormality of the data
```{r}
#the final subsetted data is still moderately skewed
skewness(sub.eggs3$AreauM2)
#squaring the subsetted data seems to have over-corrected it
sub.eggs3$sqrdarea<- (sub.eggs3$AreauM2)^2
hist(sub.eggs3$sqrdarea, breaks = 50)


#histogram of original data
hist(egg_data$AreauM2, breaks = 50, main = "Area of Eggs in the Original Data", xlab = "Area Per Egg in micrometers", ylab = "Frequency")
#the skewness of the orignial data is -1.291, suggesting the data is highly skewed
skewness(egg_data$AreauM2)
#squaring the original data
egg_data$areasqrd<- (egg_data$AreauM2)^2
#histogram of square-transformed data
hist(egg_data$areasqrd, breaks = 100)
#the skewness is now miniscule 
skewness(egg_data$areasqrd)
qqnorm(egg_data$areasqrd)

egg.model.sub.sq<- lmer(areasqrd~CrossType + Block + (1|FemaleID), data = egg_data, REML = FALSE)
plot(egg.model.sub.sq)
qqnorm(residuals(egg.model.sub.sq))
shapiro.test(residuals(egg.model.sub.sq)) #this is significant
#The same model but without parental environment, so that we can compare the two models to see if the environment had a significant effect on the size
egg.model.subsq.woenvr<- lmer(areasqrd~Block + (1|FemaleID), data = egg_data, REML = FALSE)

#The same model, but without the block to see how much of an effect block had on the size
egg.model.sub.woblock<- lmer(areasqrd~CrossType + (1|FemaleID), data = egg_data, REML = FALSE)

#running an anova to compare the first two models - will tell us if the parental environment had a significant effect. 
anova(egg.model.sub.sq, egg.model.subsq.woenvr)
```


###Potential models based on other measurements that were taken by the cell profiler

```{r}
#a mixed linear model comparing eccentricities
egg.model.sub.ecc3<- lmer(Eccentricity~CrossType + Block + (1|FemaleID), data = sub.eggs3, REML = FALSE)
summary(egg.model.sub.ecc3)
plot(egg.model.sub.ecc3)
qqnorm(residuals(egg.model.sub.ecc3))
qqline(residuals(egg.model.sub.ecc3))
#a mixed linear model comparing form factor - the residuals look very not normal
egg.model.sub.form3<- lmer(FormFactor~CrossType + Block + (1|FemaleID), data = sub.eggs3, REML = FALSE)
summary(egg.model.sub.form3)
plot(egg.model.sub.form3)
qqnorm(residuals(egg.model.sub.form3)) # data are not normal

#a mixed linear model comparing major axis 
egg.model.sub.maj3<- lmer(MajorAxisLengthPix~CrossType + Block + (1|FemaleID), data = sub.eggs3, REML = FALSE)
summary(egg.model.sub.maj3) 
plot(egg.model.sub.maj3)
qqnorm(residuals(egg.model.sub.form3))#data are not normal
```
##Final models
```{r}
par(mfrow=c(2,2))
## Fit linear mixed model to the data:
full_model<- lmer(AreauM2~CrossType * Block + (1|FemaleID), data = sub.eggs3)
## Summary including coefficient table with p-values for t-statistics using
## Satterthwaite's method for denominator degrees of freedom:
summary(full_model)

## Type III anova table with p-values for F-tests based on Satterthwaite's
## method:
aov.1 <- anova(full_model)
aov.1

## Inspect the contrast matrix for the Type III test of :
show_tests(aov.1, fractions = TRUE)$CrossType #Elise cannot get this to work

## Choose type II anova table with Kenward-Roger method for the F-test:
## Not run: don't use this unless better
if(requireNamespace("pbkrtest", quietly = TRUE))
anova(full_model, type=2, ddf="Kenward-Roger")

## Anova-like table of random-effect terms using likelihood ratio tests:
anova(full_model)#Camila originally had this as ranova()

## F-tests of 'single term deletions' for all marginal terms:
drop1(full_model)

## Least-Square means and pairwise differences:
(lsm <- ls_means(full_model)) #elise can't get this to work
ls_means(full_model, which = "block", pairwise = TRUE) 

## ls_means also have plot and as.data.frame methods:
## Not run:
plot(lsm, which=c("CrossType", "Block"))
as.data.frame(lsm)

## backward elimination of non-significant effects:
step_result <- step(full_model)

## Elimination tables for random- and fixed-effect terms:
step_result
 
# Extract the model that step found:
final_model <- get_model(step_result)
final_model

citation("lmerTest")
```

##Models that I did before we decided to subset further
```{r}
#a linear mixed model incorporating environment and block as fixed effects and and female ID as a random effect
egg.model.sub2<- lmer(AreauM2~CrossType + Block + (1|FemaleID), data = sub.eggs, REML = FALSE)
summary(egg.model.sub2)

#The same model but without environment, so that we can compare the two models to see if the environment had a significant effect on the size
egg.model.sub2.woenvr<- lmer(AreauM2~Block + (1|FemaleID), data = sub.eggs, REML = FALSE)
summary(egg.model.sub2.woenvr)

#The same model, but without the block to see how much of an effect block had on the size
egg.model.sub2.woblock<- lmer(AreauM2~CrossType + (1|FemaleID), data = sub.eggs, REML = FALSE)
summary(egg.model.sub2.woblock)

#a linear model (not mixed)
egglm<- lm(AreauM2~CrossType + Block, data = sub.eggs)
summary(egglm)
par(mfrow=c(2,2))
plot(egglm)

#comparing the model without environment as an effect - shows that environment does not have a significant effect on the area
anova(egg.model.sub2, egg.model.sub2.woenvr)
#Pr(>Chisq) = .9843

#comparing the model without block as an effect - shows that block does have a slightly significant effect on the area 
anova(egg.model.sub2, egg.model.sub2.woblock)
#Pr(>Chisq) = .0342

#plot of the residuals of the model
plot(egg.model.sub2)

#normality of the residuals
qqnorm(residuals(egg.model.sub2))
hist(residuals(egg.model.sub2), breaks = 50)

#normality of the raw data
qqnorm(egg_data$AreauM2)
hist(egg_data$AreauM2, breaks = 50)

#a mixed model with environment and block interacting
egg.model.sub.interact<- lmer(AreauM2~CrossType*Block + (1|FemaleID), data = sub.eggs, REML = FALSE)

anova(egg.model.sub2, egg.model.sub.interact)

#average size of egg by female ID
average_by_female<- aggregate(AreauM2 ~ FemaleID, data = sub.eggs, FUN = "mean")
head(average_by_female)

#a mixed linear model comparing eccentricities
egg.model.sub.ecc<- lmer(Eccentricity~CrossType + Block + (1|FemaleID), data = sub.eggs, REML = FALSE)
summary(egg.model.sub.ecc)
plot(egg.model.sub.ecc)
qqnorm(residuals(egg.model.sub.ecc))

#a mixed linear model comparing form factor
egg.model.sub.form<- lmer(FormFactor~CrossType + Block + (1|FemaleID), data = sub.eggs, REML = FALSE)
summary(egg.model.sub.form)
plot(egg.model.sub.form)
qqnorm(residuals(egg.model.sub.form))

#a mixed linear model comparing major axis 
par(mfrow = c(1,1))
egg.model.sub.maj<- lmer(MajorAxisLengthPix~CrossType + Block + (1|FemaleID), data = sub.eggs, REML = FALSE)
summary(egg.model.sub.maj) 
plot(egg.model.sub.maj)
qqnorm(residuals(egg.model.sub.form))
```

####Boxplots of the data
```{r}
#plots on the final subsetted data
Area.female.plot2<- boxplot(AreauM2~FemaleID, data = sub.eggs3, las = 2, main = "Area of eggs by each female - subsetted", xlab = "female ID", ylab = "Area in uM2", cex.axis = .55)
hist(sub.eggs3$AreauM2, breaks = 50, main = "Histogram of the area of the subsetted data", xlab = "Area in uM2")
#comparing with the max feret diameter, which is the maximum length across the egg
hist(sub.eggs3$DiameteruM, breaks = 50, main = "Histogram of the diameter of the subsetted data", xlab = "Diameter in uM")



par(mfrow = c(1,1))

#ggplot(egg_data, aes(x=FemaleID, y=AreauM2))+
#  geom_boxplot()

#boxplot of each female and their eggs area using ggplot
#ggplot(egg_data, aes(x=FemaleID, y=AreauM2), las = 2)+ geom_boxplot() 

#boxplot of each female and their eggs - not pretty 
#boxplot(AreauM2~FemaleID+Block, data = egg_data, las = 2)

#better looking boxplot of each female and the distribution of the area of their eggs with the FULL dataset
Area.female.plot<- boxplot(AreauM2~FemaleID, data = egg_data, las = 2, ylab = "Area in uM2", xlab = "Female ID", main = "Area of eggs for each female - full dataset", cex.axis = .5)

#boxplot of area by block (does not show treatment) for FULL dataset
boxplot(AreauM2~Block, data = egg_data, las = 2, main = "Area by block - full dataset", xlab = "Block", ylab = "Area in uM2")

#boxplot of area by block (does not show treatment) for final subsetted dataset
boxplot(AreauM2~Block, data = sub.eggs3, las = 2, main = "Area by block of subsetted data", xlab = "Block", ylab = "Area in uM2")

#A simple box plot to display the area relative to the environment for FULL dataset
control_vs_exposed<- boxplot(AreauM2~CrossType, data = egg_data, main = "Control vs. Exposed by Area", xlab = "Adult Environment", ylab = "Area in uM2")

#A simple box plot to display the area relative to the environment for final subsetted dataset
boxplot(AreauM2~CrossType, data = sub.eggs3, main = "Control vs. Exposed by area for subsetted data", xlab = "Adult Environment", ylab = "Area in uM2")

#boxplot of area by block with environment 
Area.envr.block.plot<- boxplot(AreauM2~CrossType+Block, data = egg_data, main = "Environment with block by area - full dataset", xlab = "Environment and Block", ylab = "Area in uM2")
#same boxplot excluding the areas under 100000 and CF04_B3
boxplot(AreauM2~CrossType+Block, data = sub.eggs3, main = "Environment with block by area - subsetted dataset", xlab = "Environment and Block", ylab = "Area in micrometers2")
legend(x = "topright", legend = c("C = control", "E = exposed", "B = block"))
boxplot(DiameteruM~CrossType+Block, data = sub.eggs3, main = "Environment with block by diameter", xlab = "Environment and Block", ylab = "Diameter in micrometers")

#boxplots for form factor
#FormFactor: (2D only) Calculated as 4*π*Area/Perimeter^2. Equals 1 for a perfectly circular object.
#environment and block
form.envr.block.plot<- boxplot(FormFactor~CrossType+Block, data = egg_data, main = "Form Factor for Block and Environment", xlab = "Block and Environment", ylab = "Form factor : 4*π*Area/Perimeter^2")
#showing just female 
form.female.plot<- boxplot(FormFactor~FemaleID, data = egg_data, las = 2, main = "Form Factor for each female", xlab = "Female ID", ylab = "Form factor : 4*π*Area/Perimeter^2", cex.axis = .55)

#boxplots for eccentricity 
#Eccentricity: (2D only) The eccentricity of the ellipse that has the same second-moments as the region. The eccentricity is the ratio of the distance between the foci of the ellipse and its major axis length. The value is between 0 and 1. (0 and 1 are degenerate cases; an ellipse whose eccentricity is 0 is actually a circle, while an ellipse whose eccentricity is 1 is a line segment.)
#environment and block
ecc.envr.block.plot<- boxplot(Eccentricity~CrossType+Block, data = egg_data, main = "Eccentricity for Block and Environment", xlab = "Block and Environment", ylab = "Eccentricity")
#showing just female 
ecc.female.plot<- boxplot(Eccentricity~FemaleID, data = egg_data, las = 2, main = "Eccentricity for each female", xlab = "Female ID", ylab = "Eccentricity", cex.axis = .55)

#boxplots for major axis length - looks very similar to the area because axis length and area are related
#environment and block
majaxis.envr.block.plot<- boxplot(MajorAxisLengthPix~CrossType+Block, data = egg_data, main = "Major Axis for Block and Environment", xlab = "Block and Environment", ylab = "Major Axis Length in pixels")
#showing just female 
majaxis.female.plot<- boxplot(MajorAxisLengthPix~FemaleID, data = egg_data, las = 2, main = "Major Axis Length for each female", xlab = "Female ID", ylab = "Major Axis Length", cex.axis = .55)
```

###Violin plots
```{r}
#Violin plot depicting each females egg distribution
ggplot(aes(FemaleID, AreauM2), data = sub.eggs3)+ geom_violin()

boxplot(AreauM2~CrossType+Block, data = sub.eggs3)
#I want to make a graph similar to this one, but I can't figure out how to with ggplot
#ggplot(aes(Block+CrossType, AreauM2), data = sub.eggs3)+ geom_violin() 

#tried getting the vioplot package to do this, but it said it could not find the "vioplot" function

#with(data = sub.eggs3, (vioplot(AreauM2 ~ CrossType + Block)))
```








