---
title: "Larvae Biomineralization Block 1"
output: html_document
---

Load libraries
```{r LoadLibraries, include=FALSE}
library(knitr)
library(cowplot)
library(plyr)
library(tidyverse)
library(sciplot)
library(reshape2)
library(lme4)
library(lmerTest)
library(blme)
library(grid)
library(data.table)
library(gridExtra)
library(lattice)
library(optimx)
library(car)
library(Hmisc)
library(corrplot)
library(emmeans)
library(chron)
library(GLMMadaptive)
library(calibrate)
library(merTools)
```

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

Load data
Name the data you export from FileMaker Pro by their exact table names and save them as CSVs, e.g. Larvae_Morphology.csv
```{r LoadData}
#setwd("~/Repos/LarvaeTransGen2018")
setwd("~/R/GitHub/LarvaeTransGen2018/data") #Elise's working directory
#upload all of the data tables for the larvae experiment
Larvae_Morphology <- read.csv("../data/Larvae_Morphology.csv") #contains data from CellProfiler from the larvae outlines
Barcode_Jar <- read.csv("../data/Barcode_Jar.csv") #contains data on CrossID, seatable, and whether or not larvae were present
Block_ID <- read.csv("../data/Block_ID.csv") #contains data on the block
Cross<- read.csv("../data/Cross.csv") #contains data on the female and male IDs for the cross and whether the cross was for QG or Meth
Fert_QG<- read.csv("../data/Fert_QG.csv") #contains data on fertilization counts, the JarIDs for the crosses
Header_WC<- read.csv("../data/Header_WC.csv") #contains data on the header tanks used to fill the jars
Larvae_Calcification <- read.csv("../data/Larvae_Calcification.csv") #contains data on filter weights and counts
Larvae_Counts <- read.csv("../data/Larvae_Counts.csv") #contains data on sedgewick rafter larvae counts
Larvae_WC <- read.csv("../data/Larvae_WC.csv") #contains data on larvae jar water chemistry
Parent_ID <- read.csv("../data/Parent_ID.csv") #contains data on the adults at the time of shucking. Includes Parent ID assignments
Larvae_Cilia<- read.csv("../data/Larvae_Cilia.csv") #contains data on cilia extrusion of the larvae photographed for morphology
Egg_Morphology<- read.csv("../data/Egg_Morphology.csv") #contains data on egg sizes from CellProfiler
Adult_Sample<- read.csv("../data/Adult_Sample.csv") #contains data on adult oysters at the time of collection
WC_Standard<- read.csv("../data/WC_Standard.csv") #contains data on the standards used for water chemistry
Tank_WC<- read.csv("../data/Tank_WC.csv") #contains data on the adult tank water chemistry
```

Make fields factors: 
```{r MakeFieldsFactors}
Adult_Sample$AccTankID<- as.factor(Adult_Sample$AccTankID)
Adult_Sample$TrtTankID<- as.factor(Adult_Sample$TrtTankID)
Barcode_Jar$JarSeatable<- as.factor(Barcode_Jar$JarSeatable)
Barcode_Jar<- subset(Barcode_Jar, JarSeatable =="2" | JarSeatable =="4" | JarSeatable =="6")
Tank_WC$ParentTrt<- as.factor(Tank_WC$ParentTrt)
Tank_WC$Tank<- as.factor(Tank_WC$Tank)
WC_Standard$CRM<- as.factor(WC_Standard$CRM)
Larvae_WC$DateFilter<- as.factor(Larvae_WC$DateFilter)
```

Look at the water chemistry data for adult tanks
```{r AdultTankChemistry}
#subset the tank data to only include B1 
Tank_WCB1<- subset(Tank_WC, UseForCalcite == "TRUE")
Tank_WCB1$Tank<-droplevels(Tank_WCB1)$Tank #drop unused levels
Tank_WCB1$ParentTrt<- droplevels(Tank_WCB1)$ParentTrt
#test to see if tanks are significantly different. 
boxplot(SatCalcite_ATDIC~Tank, data=Tank_WCB1, na.omit=TRUE)
#check for significant differences in saturation state between tanks within treatments
Ca1<- lmer(SatCalcite_ATDIC~ParentTrt + (1|Tank), data=Tank_WCB1)
Ca2<- lmer(SatCalcite_ATDIC~ 1 + (1|Tank), data= Tank_WCB1)
anova(Ca1, Ca2) #select Ca1
Ca3<- lm(SatCalcite_ATDIC~ParentTrt, data=Tank_WCB1)
anova(Ca1, Ca3) #select Ca3
#check assumptions
par(mfcol=c(2,2))
plot(Ca3) #seems to meet assumptions of homoscedasticity and linearity. Normality could look better. 
par(mfcol=c(1,1))
acf(resid(Ca3)) #this looks good. 
summary(Ca3) #There is a significant effect of ParentTrt, but not tank according to best model. 
#try a transformation to see if it helps normality
Ca1log<- lmer(log(SatCalcite_ATDIC)~ParentTrt+(1|Tank), data=Tank_WCB1)
Ca3log<- lm(log(SatCalcite_ATDIC)~ParentTrt, data=Tank_WCB1)
anova(Ca1log, Ca3log) #still choose Ca3log
par(mfcol=c(2,2))
plot(Ca3log) #now meets assumptions 
par(mfcol=c(1,1))
acf(resid(Ca3log))#looks good
summary(Ca3log)
#create dataframe that has the mean values for each tank
TankMeans<- ddply(Tank_WCB1, .(Tank, Block, ParentTrt), numcolwise(mean, na.rm=T))#can't use this for pH and salinity, remember, it is just for carb chem
```

Join the data into dataframe for Egg analysis
```{r JoinDataForEggAnalysis}
#First dataframe will be for examining egg morphology for all eggs traced; multiple values for each adult
ParentInfo<- merge(x=Adult_Sample, y=Parent_ID, by="AdultID", all.x=TRUE)
ParentInfo<- merge(x=ParentInfo, y=TankMeans, by.x="TrtTankID", by.y="Tank", all.x=TRUE)
Eggs<- merge(x=Egg_Morphology, y=ParentInfo, by.x="FemaleID", by.y="ParentBlockID", all.x=TRUE)
Eggs$FemaleID<- as.factor(Eggs$FemaleID)
#subset the data to only include B1
EggsB1<- subset(Eggs, Block =="B1")
```

Visualize the data for eggs to start
```{r EggVisualization}
#get the original range of the eggs
range(EggsB1$EggDiamum)
#look at the egg measurements to see outliers
boxplot(EggDiamum~FemaleID, data=EggsB1)
boxplot(EggDiamum~ParentTrt, data=EggsB1)
```

Remove egg outliers
```{r EggOutlierRemoval}
#check for outliers using code from https://www.r-bloggers.com/identify-describe-plot-and-remove-the-outliers-from-the-dataset/. The function uses Tukey's method to ID ouliers ranged above and below the 1.5*IQR. 
outlierKD <- function(dt, var) {
     var_name <- eval(substitute(var),eval(dt))
     na1 <- sum(is.na(var_name))
     m1 <- mean(var_name, na.rm = T)
     par(mfrow=c(2, 2), oma=c(0,0,3,0))
     boxplot(var_name, main="With outliers")
     hist(var_name, main="With outliers", xlab=NA, ylab=NA)
     outlier <- boxplot.stats(var_name)$out
     mo <- mean(outlier)
     var_name <- ifelse(var_name %in% outlier, NA, var_name)
     boxplot(var_name, main="Without outliers")
     hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
     title("Outlier Check", outer=TRUE)
     na2 <- sum(is.na(var_name))
     cat("Outliers identified:", na2 - na1, "n")
     cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "n")
     cat("Mean of the outliers:", round(mo, 2), "n")
     m2 <- mean(var_name, na.rm = T)
     cat("Mean without removing outliers:", round(m1, 2), "n")
     cat("Mean if we remove outliers:", round(m2, 2), "n")
     response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
     if(response == "y" | response == "yes"){
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     } else{
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     }
}
outlierKD(EggsB1, EggDiamum) #I opted to remove outliers for now. 
#Outliers identified: 48 nPropotion (%) of outliers: 11.5 nMean of the outliers: 57.63 nMean without removing outliers: 62.4 nMean if we remove outliers: 62.95 n
yes #to remove outliers from EggsB1
#count how many eggs are left per female after removing the outliers. 
outrem<- aggregate(EggDiamum~FemaleID, data=EggsB1, FUN=length)
#EF07_B1 had 10 eggs that were included the rest were removed. Keep EF07_B1 in the analysis
```

Visualize eggs without outliers
```{r EggVisualizationNoOutliers}
#Plot again without outliers
par(mfcol=c(1,1))
boxplot(EggDiamum~FemaleID, data=EggsB1) 
boxplot(EggDiamum~ParentTrt, data=EggsB1)
#get the range now
range(EggsB1$EggDiamum, na.rm=TRUE)
#note that the eggs were filtered through a 70 um filter, but it appears that larger ones made it through.We are going to keep these in for now because they aren't that outrageously large. 
#plot histograms of the egg data with outliers removed
ggplot(EggsB1, aes(x=EggDiamum, color=ParentTrt))+
  geom_histogram(alpha=0.5, position="identity")
eggmeans<- aggregate(EggDiamum~ParentTrt, data=EggsB1, FUN=mean) #400 = 63.45709; 2800= 62.36829
eggmedians<- aggregate(EggDiamum~ParentTrt, data=EggsB1, FUN=median)#400 = 62.91200; 2800= 62.06727
view(eggmeans)
view(eggmedians)
#I will use the mean egg size. 
```

Make datasets without the egg size outliers
```{r MakeLarvaeDataframes}
meanegg<- ddply(EggsB1, .(FemaleID), numcolwise(mean, na.rm=T)) #get mean egg size
#select only columns that we need for egg morph
meanegg<- subset(meanegg, select=c("FemaleID", "EggDiamum"))
FemAdults<- merge(x=ParentInfo, y= meanegg, by.x="ParentBlockID", by.y= "FemaleID", all.y=TRUE, all.x=TRUE)
#Next dataframe will be for one line per traced larva, includes all the data for that jar, including average egg size for that jar 
Larvae<-merge(x=Larvae_Morphology, y=Barcode_Jar, by="JarID", all.x=TRUE)
Larvae<- merge(x=Larvae, y=Larvae_WC, by="JarID", all.x=TRUE, all.y=TRUE)
#use gather function in tidyr to make the Fert_QG Jar ID columns in long form vs. wide form
Fert_QG_long <- Fert_QG %>% tidyr::gather(Key, JarID, JarID1:JarID6)
#use the gathered data frame to merge with Larvae
Larvae<-merge(x=Larvae, y=Fert_QG_long, by.x= c("JarID", "CrossID"), by.y=c("JarID", "CrossID"), all.x=TRUE) #did the join by two columns so CrossID column wasn't duplicated
Larvae<- merge(x=Larvae, y=Cross, by="CrossID", all.x=TRUE)
Larvae<- merge(x=Larvae, y=FemAdults, by.x= "FemaleID", by.y="ParentBlockID", all.x=TRUE)
Larvae<- merge(x=Larvae, y= Larvae_Counts, by="JarID", all.x=TRUE)
#Remove jars without larvae
Larvae<- subset(Larvae, Larvae=="TRUE")
#get the growth per day for the larva
Larvae$GrowthPerDay<- (Larvae$LarvaeDiamum-Larvae$EggDiamum)/3
#Final dataframe will be one line per Jar, includes all data for the jar, including the average larva size per jar and cilia
LarvByJar<-merge(x=Larvae_Calcification, y=Larvae_Cilia, by="JarID", all.x=TRUE, all.y=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Barcode_Jar, by="JarID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Larvae_WC, by="JarID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Fert_QG_long, by.x=c("JarID", "CrossID"), by.y=c("JarID","CrossID"), all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Cross, by="CrossID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=FemAdults, by.x="FemaleID", by.y="ParentBlockID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Larvae_Counts, by="JarID", all.x=TRUE)
#get the means for larvae morphology and add to LarvByJar
MeanLarvMorph<- ddply(Larvae_Morphology, .(JarID), numcolwise(mean, na.rm=T))
SELarvMorph<- ddply(Larvae_Morphology, .(JarID), numcolwise(se, na.rm=T)) #get standard error of larva size
MeanLarvMorph<- full_join(MeanLarvMorph,SELarvMorph, by=c("JarID", "JarID"),suffix=c("","SE")) #join the mean and se larvae data
LarvByJar<- merge(x=LarvByJar, y=MeanLarvMorph, by="JarID", all.x=TRUE)
#Now let's add the survival data. See notes on "Larvae Survival" for more info on how we decided to count total larvae in the jar
#Make columns for the two larvae counts for 1_5 and 2
LarvByJar$v1_5SurvCount<- LarvByJar$SWRTotal+LarvByJar$F2LarvaeCount
LarvByJar$v2SurvCount<- LarvByJar$F1LarvaeCount+LarvByJar$F2LarvaeCount
#use ifelse statement to define the LarvaeSurvived column
LarvByJar$LarvaeSurvived<- ifelse(LarvByJar$ProtocolVersion =="1", paste(LarvByJar$TotalLarvae), ifelse(LarvByJar$ProtocolVersion =="1_5", paste(LarvByJar$v1_5SurvCount), paste(LarvByJar$v2SurvCount)))
LarvByJar$LarvaeSurvived<- as.numeric(LarvByJar$LarvaeSurvived)
#get the total larvae for the control jars and exposed jars for each family and then get the ratio of exposed to control to get an idea of how good a male and female pair are at larvae surviving. Can also use this to create reaction norms. 
SurvDat<- ddply(LarvByJar, .(CrossID, ParentTrt, JarTrt, MaleID, FemaleID, TrtTankID, AccTankID, Block), numcolwise(mean, na.rm=TRUE))
SESurvDat<- ddply(LarvByJar, .(CrossID, ParentTrt, JarTrt, Block), numcolwise(se, na.rm=TRUE)) #get the standard errors for each mean count
SESurvDat<- subset(SESurvDat, select= c(CrossID, JarTrt, LarvaeSurvived, SatCalcite_ATDIC, JarSatCalcite_SW)) #rename the columns
colnames(SESurvDat)<- paste("SE", colnames(SESurvDat), sep="")
#merge the standard error data to the SurvDat dataframe
SurvDat<- merge(SurvDat, SESurvDat, by.x= c("CrossID", "JarTrt"), by.y= c("SECrossID", "SEJarTrt"))
#only keep the relevant fields because otherwise this is a huge dataframe
SurvDat<- subset(SurvDat, select= c(CrossID, JarTrt, ParentTrt, MaleID, FemaleID, TrtTankID, AccTankID, LarvaeSurvived, Block, SELarvaeSurvived, SatCalcite_ATDIC)) 
SurvDat<- subset(SurvDat, Block=="B1")
SurvDat<- subset(SurvDat, CrossID!="")
#use the spread function to go from long form to wide form for the survival data
SurvDatToWide<- unite(SurvDat, Value, c(LarvaeSurvived, SELarvaeSurvived), sep="_")
SurvWideDat<- spread(SurvDatToWide, JarTrt, value= Value)
#now split the mean and SE apart again
SurvWideDat<- separate(SurvWideDat, Control, into= c("MeanLarvSurvivedCon", "SELarvSurvivedCon"), sep="_")
SurvWideDat<- separate(SurvWideDat, Exposed, into= c("MeanLarvSurvivedExp", "SELarvSurvivedExp"), sep="_")
#make the means and ses numeric
SurvWideDat$MeanLarvSurvivedExp<- as.numeric(SurvWideDat$MeanLarvSurvivedExp)
SurvWideDat$MeanLarvSurvivedCon<- as.numeric(SurvWideDat$MeanLarvSurvivedCon)
SurvWideDat$SELarvSurvivedExp<- as.numeric(SurvWideDat$SELarvSurvivedExp)
SurvWideDat$SELarvSurvivedCon<- as.numeric(SurvWideDat$SELarvSurvivedCon)
SurvWideDat$RatSurv<- SurvWideDat$MeanLarvSurvivedExp/SurvWideDat$MeanLarvSurvivedCon
SurvWideDat$SurvChange<- (SurvWideDat$MeanLarvSurvivedCon-SurvWideDat$MeanLarvSurvivedExp)/SurvWideDat$MeanLarvSurvivedCon
SurvWideDat$CrossID<- as.factor(SurvWideDat$CrossID)
SurvWideDat$FemaleID<- as.factor(SurvWideDat$FemaleID)
SurvWideDat$MaleID<- as.factor(SurvWideDat$MaleID)
LarvByJar<- subset(LarvByJar, Larvae=="TRUE")
#now remove the data we do not want to analyze.
LarvaeDat<- subset(Larvae, Block == "B1")
LarvaeDat$JarID<- as.factor(LarvaeDat$JarID)
LarvaeDat$AccTankID<- as.factor(LarvaeDat$AccTankID)
LarvaeDat$TrtTankID<- as.factor(LarvaeDat$TrtTankID)
LarvaeDat$JarSeatable<- as.factor(LarvaeDat$JarSeatable)
LarvaeDat$JarTrt<- as.factor(LarvaeDat$JarTrt)
LarvaeDat$ParentTrt<- as.factor(LarvaeDat$ParentTrt)
LarvaeDat$FemaleID<- as.factor(LarvaeDat$FemaleID)
LarvaeDat$MaleID<- as.factor(LarvaeDat$MaleID)
LarvaeDat$CrossID<- as.factor(LarvaeDat$CrossID)
LarvByJarDat<- subset(LarvByJar, Block == "B1")
LarvByJarDat$JarID<- as.factor(LarvByJarDat$JarID)
LarvByJarDat$AccTankID<- as.factor(LarvByJarDat$AccTankID)
LarvByJarDat$TrtTankID<- as.factor(LarvByJarDat$TrtTankID)
LarvByJarDat$JarSeatable<- as.factor(LarvByJarDat$JarSeatable)
LarvByJarDat$ParentTrt<- as.factor(LarvByJarDat$ParentTrt)
LarvByJarDat$FemaleID<- as.factor(LarvByJarDat$FemaleID)
LarvByJarDat$MaleID<- as.factor(LarvByJarDat$MaleID)
LarvByJarDat$CrossID<- as.factor(LarvByJarDat$CrossID)
LarvByJarDat$JarTrt<- as.factor(LarvByJarDat$JarTrt)
#make a column for the ratio of major to minor axis
LarvaeDat$MajMinRat<- LarvaeDat$LarvaeMajorAxisLengthPix/LarvaeDat$LarvaeMinorAxisLengthPix
LarvaeDat$PerimDiamRat<-LarvaeDat$LarvaePerimeterum/LarvaeDat$LarvaeDiamum
LarvaeDat<- subset(LarvaeDat, JarSatCalcite_SW !="NA")
LarvByJarDat<- subset(LarvByJarDat, JarSatCalcite_SW !="NA")
```

Analyze jar chemistry data
```{r JarChemistryStat}
CalJar<- subset(LarvByJarDat, BottleChemistry=="TRUE" )#only use those that had their chem measured
#look at the saturation state for jars
boxplot(JarSatCalcite_SW~JarTrt*JarSeatable, data=CalJar)
#I was having trouble with making this model: JarSatCalcite_SW~JarTrt + (1|ParentTrt)+ (1|TimeFilter), so I looked at TimeFilter because I think that is the problem. 
Jar1<- lm(JarSatCalcite_SW~TimeFilter, data=CalJar)
par(mfcol=c(2,2))
plot(Jar1)
par(mfcol=c(1,1))
#looks like we can't use timefilter because it is too specific to each jar. Get the error message "observations with leverage one"; try using seatable instead
Jar2<- lmer(JarSatCalcite_SW~JarTrt + (1|JarSeatable), data=CalJar)
Jar2step<- step(Jar2)
#get error model not of class lmerMod. I looked up this issue and turns out that error results when all random effects are dropped. See https://stackoverflow.com/questions/55638476/how-to-do-stepwise-model-with-random-effect-lme4-lmertest for more info if you want. 
JarFin<- lm(JarSatCalcite_SW~JarTrt, data=CalJar)
#check assumptions 
par(mfcol=c(2,2))
plot(JarFin)
par(mfcol=c(1,1))
acf(resid(JarFin))
summary(JarFin)
#Conclusion: JarSeatable doesn't need to be accounted for as a random effect in terms of water chem. 
```

Test for significant effect of parent treatment on egg size. Start with measured water chem
```{r EggSizeStat}
CalEggs<- subset(EggsB1, SatCalcite_ATDIC != "NA")
CalEggs$TrtTankID<-droplevels(CalEggs)$TrtTankID
#start with models that have the response variable EggDiamum
eggdiam1<- lmer(EggDiamum~SatCalcite_ATDIC*AdultLength + (1|FemaleID)  + (1|AccTankID), data=CalEggs)
eggstep<-step(eggdiam1)
print(eggstep)
#final model chosen: EggDiamum~ 1 + (1|FemaleID)
eggdiamfin<- lmer(EggDiamum~1+(1|FemaleID), data=CalEggs)
#Check assumptions
par(mfcol=c(1,1))
qqnorm(resid(eggdiamfin))
qqline(resid(eggdiamfin))
plot(eggdiamfin)
acf(resid(eggdiamfin))
#the above looks fine, seems to meet all assumptions. 
summary(eggdiamfin)
#Conclusion: Female matters but not treatment or adult length. Acclimation tank and treatment tank also don't matter.  
plot(EggDiamum~FemaleID, data=CalEggs) #we need to account for FemaleID, but not egg size in Larvae models. Can also make parallel models one with EggDiamum as a covariate and one with Female ID as a random effect
```

```{r EggSizeStatFixedEffect, include=FALSE}
eggdiam1f<- lmer(EggDiamum~ParentTrt*AdultLength + (1|FemaleID) + (1|TrtTankID)+ (1|AccTankID), data=CalEggs)
eggstepf<-step(eggdiam1f)
print(eggstepf)
#final model chosen: EggDiamum~ 1 + (1|FemaleID), which is the same as the eggdiamfin model
summary(eggdiamfin)
```

Choosing mean or median for larvae length data
```{r LarvaeDiamMeanOrMedian}
ggplot(LarvaeDat, aes(x=LarvaeDiamum, color=JarTrt))+
  geom_histogram(alpha=0.5, position="identity") +
  facet_grid(~ParentTrt)
larvmeans<- aggregate(LarvaeDiamum~ParentTrt*JarTrt, data=LarvaeDat, FUN=mean)
larvmedians<- aggregate(LarvaeDiamum~ParentTrt*JarTrt, data=LarvaeDat, FUN=median)
view(larvmeans)
view(larvmedians)
#I think we can use the mean for these data 

#look for the number of jars in each treatment
larvjarn<- aggregate(LarvaeDiamum~ParentTrt*JarTrt, data=LarvByJarDat, FUN=length)
```

Look at the larvae diameter data
```{r LarvaeDiamStat}
boxplot(LarvaeDiamum~FemaleID, data=LarvaeDat) #I looked specifically for weird things with EF07_B1 because that individual only had 10 eggs after removing outliers. But all seems fine! 
#create most complex model and then use the step function to see which model is best fit for the data; these are for ParentTrt and JarTrt as covariates
larvlen<-lmer(LarvaeDiamum~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarID)+(1|JarSeatable)+(1|AccTankID), data=LarvaeDat)
steppedlarvlen<-step(larvlen)
print(steppedlarvlen)
#final model chosen by the lme4 step function: y~SatCalcite_ATDIC*JarSatCalcite_SW +(1|FemaleID)+(1|MaleID)+(1|JarID)+ (1|JarSeatable)
larvlenfin<- lmer(LarvaeDiamum~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|JarID)+ (1|JarSeatable), data=LarvaeDat)
#got warning message that model failed to converged. I read this stack overflow talking about checking to see if it is acceptable: https://stats.stackexchange.com/questions/97929/lmer-model-fails-to-converge
TestConvDiam <- with(larvlenfin@optinfo$derivs,solve(Hessian,gradient))
max(abs(TestConvDiam)) #this seems small enough that it's alright
#check the final model to see that it meets assumptions
plot(larvlenfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvlenfin))
qqline(resid(larvlenfin)) #has relatively long tails, doesn't seem to meet assumption of normality. Check on this
acf(resid(larvlenfin)) #this looks good 
summary(larvlenfin)
#Conclusion: signficant main effect of Parent and Jar calcite sat, and sig. interaction between parent and Jar saturation states
#add predicted length
LarvaeDat$PredLength<- predict(larvlenfin)
#get 95% confidence interval for the linear model
predint <- predictInterval(larvlenfin, LarvaeDat, type="linear.prediction", level=0.95)
head(predint)
LarvaeDat<- cbind(LarvaeDat, predint)
#I don't know why the predict and predictInterval aren't matching. I think predictInterval is better so use that fit
```

Use fixed effect for adult treatment and larvae treatment as covariate
```{r LarvDiamStatParentTrtCovariateJarTrtFixedEffect, include=FALSE}
larvlenJF<-lmer(LarvaeDiamum~ParentTrt*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarID)+(1|JarSeatable)+(1|AccTankID), data=LarvaeDat)
steppedlarvlenJF<-step(larvlenJF)
print(steppedlarvlenJF)
#final model chosen by the lme4 step function: y~SatCalcite_ATDIC*JarTrt +(1|FemaleID)+(1|MaleID)+(1|JarID)+ (1|JarSeatable)
larvlenfinJF<- lmer(LarvaeDiamum~ParentTrt*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|JarID)+ (1|JarSeatable), data=LarvaeDat)

#check the final model to see that it meets assumptions
plot(larvlenfinJF) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvlenfinJF))
qqline(resid(larvlenfinJF)) #has relatively long tails, I think this should be acceptable
acf(resid(larvlenfinJF)) #this looks good 
summary(larvlenfinJF)
```

Use fixed effect for both adult and larvae treatments
```{r LarvDiamStatBothFixedEffects, include=FALSE}
larvlenFE<-lmer(LarvaeDiamum~ParentTrt*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarID)+(1|JarSeatable)+(1|TrtTankID)+(1|AccTankID), data=LarvaeDat)
steppedlarvlenFE<-step(larvlenFE)
print(steppedlarvlenFE)
#final model chosen by the lme4 step function: y~ParentTrt*JarTrt +(1|FemaleID)+(1|MaleID)+(1|JarID)+ (1|JarSeatable)
larvlenfinFE<- lmer(LarvaeDiamum~ParentTrt*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|JarID)+ (1|JarSeatable), data=LarvaeDat)
#check the final model to see that it meets assumptions
plot(larvlenfinFE) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvlenfinFE))
qqline(resid(larvlenfinFE)) #has relatively long tails, I think this should be acceptable
acf(resid(larvlenfinFE)) #this looks good 
summary(larvlenfinFE)
```

Analyze at growth per day. This will likely be the main figure
```{r GrowthPerDayStat}
boxplot(GrowthPerDay~ParentTrt*JarTrt, data=LarvaeDat)
larvgrow<-lmer(GrowthPerDay~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarID) +(1|JarSeatable)+(1|AccTankID) , data=LarvaeDat)
steppedlarvgrow<- step(larvgrow)
print(steppedlarvgrow)
#final model chosen by the lme4 step function: y~SatCalcite_ATDIC*JarSatCalcite_SW+(1|FemaleID)+(1|MaleID)+ (1|CrossID)+ (1|JarID) + (1|JarSeatable)
larvgrowfin<- lmer(GrowthPerDay~SatCalcite_ATDIC*JarSatCalcite_SW+(1|FemaleID)+(1|MaleID)+(1|CrossID) + (1|JarID) + (1|JarSeatable), data=LarvaeDat)
#check the final model to see that it meets assumptions
plot(larvgrowfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvgrowfin))
qqline(resid(larvgrowfin)) #has relatively long tails, based on scale I think it reasonably meets assumption of normality. 
acf(resid(larvgrowfin))
TestConvGrow <- with(larvgrowfin@optinfo$derivs,solve(Hessian,gradient))
max(abs(TestConvGrow)) #this seems small enough that it's alright
summary(larvgrowfin)
#Conclusion: Signficant main effect of tank and jar calcite saturation state and significant interaction.
```

Analyze larvae growth with adult trt as covariate and jar as fixed effect
```{r LarvGrowthStatCovariateJarFixed, include=FALSE}
larvgrowJF<-lmer(GrowthPerDay~SatCalcite_ATDIC*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarID) +(1|JarSeatable)+(1|AccTankID) , data=LarvaeDat)
steppedlarvgrowJF<- step(larvgrowJF)
print(steppedlarvgrowJF)
#final model chosen by the lme4 step function: y~SatCalcite_ATDIC*JarTrt+(1|FemaleID)+(1|MaleID)+ (1|CrossID)+ (1|JarID) + (1|JarSeatable)
larvgrowfinJF<- lmer(GrowthPerDay~SatCalcite_ATDIC*JarTrt+(1|FemaleID)+(1|MaleID)+(1|CrossID) + (1|JarID) + (1|JarSeatable), data=LarvaeDat)
#check the final model to see that it meets assumptions
plot(larvgrowfinJF) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvgrowfinJF))
qqline(resid(larvgrowfinJF)) #has relatively long tails, based on scale I think it reasonably meets assumption of normality. 
acf(resid(larvgrowfinJF))
summary(larvgrowfinJF)
```

Analyze larvae growth with adult trt as covariate and jar as fixed effect
```{r LarvGrowthStatBothFixedEffects, include=FALSE}
larvgrowFE<-lmer(GrowthPerDay~ParentTrt*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarID) +(1|JarSeatable)+(1|TrtTankID)+(1|AccTankID) , data=LarvaeDat)
steppedlarvgrowFE<- step(larvgrowFE)
print(steppedlarvgrowFE)
#final model chosen by the lme4 step function: y~SatCalcite_ATDIC*JarTrt+(1|FemaleID)+(1|MaleID)+ (1|CrossID)+ (1|JarID) + (1|JarSeatable)
larvgrowfinFE<- lmer(GrowthPerDay~ParentTrt*JarTrt+(1|FemaleID)+(1|MaleID)+(1|CrossID) + (1|JarID) + (1|JarSeatable), data=LarvaeDat)
#check the final model to see that it meets assumptions
plot(larvgrowfinFE) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvgrowfinFE))
qqline(resid(larvgrowfinFE)) #has relatively long tails, based on scale I think it reasonably meets assumption of normality. 
acf(resid(larvgrowfinFE))
summary(larvgrowfinFE)
```

Look at larvae area
```{r LarvaeAreaStat}
boxplot(LarvaeAreaum2~ParentTrt*JarTrt, data=LarvaeDat)
#create most complex model and then use the step function to see which model is best fit for the data; these are for ParentTrt and JarTrt as covariates
larvarea<-lmer(LarvaeAreaum2~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) , data=LarvaeDat)
steppedlarvarea<-step(larvarea, alpha.fixed=0.05)
print(steppedlarvarea)
#final model chosen by the lme4 step function: y~SatCalcite_ATDIC*JarSatCalcite_SW+(1|FemaleID)+(1|CrossID)+(1|JarID)
larvareafin<- lmer(LarvaeAreaum2~SatCalcite_ATDIC*JarSatCalcite_SW+(1|FemaleID)+(1|CrossID)+(1|JarID), data=LarvaeDat)
#check the final model to see that it meets assumptions
plot(larvareafin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvareafin))
qqline(resid(larvareafin)) #has relatively long tails, doesn't seem to meet assumption of normality. 
acf(resid(larvareafin)) #this looks fine to me. 
summary(larvareafin) #try transforming the data
larvareat<-lmer(sqrt(LarvaeAreaum2)~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) , data=LarvaeDat)
#got warning message that model failed to converged. I read this stack overflow talking about checking to see if it is acceptable: https://stats.stackexchange.com/questions/97929/lmer-model-fails-to-converge
TestConverge <- with(larvareat@optinfo$derivs,solve(Hessian,gradient))
max(abs(TestConverge))
#this value seems small enough to be acceptable
steppedlarvareat<-step(larvareat, alpha.fixed = 0.051)#changed this because it was eliminating adult trt when that had a pvalue of 0.05043
print(steppedlarvareat)
#final model chosen by the lme4 step function: y~SatCalcite_ATDIC*JarSatCalcite_SW+(1|FemaleID)+(1|MaleID)+(1|JarID)+(1|JarSeatable)
larvareafint<- lmer(sqrt(LarvaeAreaum2)~SatCalcite_ATDIC*JarSatCalcite_SW+(1|FemaleID)+(1|MaleID)+ (1|JarSeatable)+(1|JarID), data=LarvaeDat)
TestConvergeT <- with(larvareafint@optinfo$derivs,solve(Hessian,gradient))
max(abs(TestConvergeT))
plot(larvareafint) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvareafint))
qqline(resid(larvareafint)) #this seems to have improved it
acf(resid(larvareafint)) #this looks fine to me. 
summary(larvareafint)
#significant main effects of jar treatments and significant interaction 0.0504 seems like we can call that significant it is so close
```

Analyze surface area with Parent Trt as covariate and Jar Treatment as fixed effect
```{r SurfaceAreaPrtTrtCovariateJarTrtFixed, include=FALSE}
larvareatJF<-lmer(sqrt(LarvaeAreaum2)~SatCalcite_ATDIC*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) , data=LarvaeDat)
steppedlarvareatJF<-step(larvareatJF)
print(steppedlarvareatJF)
#final model chosen by the lme4 step function: y~SatCalcite_ATDIC*JarTrt+(1|FemaleID)+(1|MaleID)+(1|JarID)+(1|JarSeatable)
larvareafintJF<- lmer(sqrt(LarvaeAreaum2)~SatCalcite_ATDIC*JarTrt+(1|FemaleID)+(1|MaleID)+(1|JarID)+(1|JarSeatable), data=LarvaeDat)
plot(larvareafintJF) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvareafintJF))
qqline(resid(larvareafintJF)) #this seems to have improved it
acf(resid(larvareafintJF)) #this looks fine to me. 
summary(larvareafintJF)
```

Analyze surface area with Parent and Jar Treatment as fixed effects
```{r SurfaceAreaBothFixedEffects, include=FALSE}
larvareatFE<-lmer(sqrt(LarvaeAreaum2)~ParentTrt*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) + (1|TrtTankID), data=LarvaeDat)
steppedlarvareatFE<-step(larvareatFE)
print(steppedlarvareatFE)
#final model chosen by the lme4 step function: y~ParentTrt*JarTrt+(1|FemaleID)+(1|MaleID)+(1|JarID)+(1|JarSeatable)
larvareafintFE<- lmer(sqrt(LarvaeAreaum2)~ParentTrt*JarTrt+(1|FemaleID)+(1|MaleID)+(1|JarID)+(1|JarSeatable), data=LarvaeDat)
plot(larvareafintFE) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvareafintFE))
qqline(resid(larvareafintFE)) #this seems to have improved it
acf(resid(larvareafintFE)) #this looks fine to me. 
summary(larvareafintFE)
```

Characterize the larvae by the ratio of major and minor axis
```{r LarvaeAxesStat}
boxplot(MajMinRat~JarTrt*ParentTrt, data=LarvaeDat)
larvrat<-lmer(MajMinRat~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) +(1|CrossID), data=LarvaeDat)
larvratstep<- step(larvrat)
print(larvratstep)
larvratfin<- lmer(MajMinRat~SatCalcite_ATDIC*JarSatCalcite_SW +(1|FemaleID)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larvratfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvratfin))
qqline(resid(larvratfin)) #has relatively long tails, doesn't seem to meet assumption of normality. try a transformation
acf(resid(larvratfin)) #this looks fine.
summary(larvratfin) #significant effect of parent trt and significant interaction between parent treatment and larval treatment
#try transforming the data
larvratt<-lmer(MajMinRat^1/3~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) +(1|CrossID), data=LarvaeDat)
larvratstept<- step(larvratt)
print(larvratstept)
larvratfint<- lmer((MajMinRat^1/3)~SatCalcite_ATDIC*JarSatCalcite_SW +(1|FemaleID)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larvratfint) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvratfint))
qqline(resid(larvratfint)) #has relatively long tails, doesn't seem to meet assumption of normality despite transformation
summary(larvratfint) #significant effect of parent trt and significant interaction between parent treatment and larval treatment
#I have tried transforming with reciprocal, square root, square, and cube root none of them have helped with normality. No matter what though, still get a significant effect of parent treatment and significant interaction
```

Analyze the larvae by the ratio of major and minor axis with parent as a covariate and larvae as a fixed effect
```{r LarvAxesStatParentCovariateLarvaeFixedEffect, include=FALSE}
larvratJF<-lmer(MajMinRat~SatCalcite_ATDIC*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) +(1|CrossID), data=LarvaeDat)
larvratstepJF<- step(larvratJF)
print(larvratstepJF)
larvratfinJF<- lmer(MajMinRat~SatCalcite_ATDIC*JarTrt +(1|FemaleID)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larvratfinJF) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvratfinJF))
qqline(resid(larvratfinJF)) #has relatively long tails, doesn't seem to meet assumption of normality. try a transformation
acf(resid(larvratfinJF)) #this looks fine.
summary(larvratfinJF)
```

Analyze the larvae by the ratio of major and minor axis with parent and larvae as fixed effects
```{r LarvAxesStatBothFixedEffects, include=FALSE}
larvratFE<-lmer(MajMinRat~ParentTrt*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)+(1|TrtTankID)+(1|AccTankID) +(1|CrossID), data=LarvaeDat)
larvratstepFE<- step(larvratFE)
print(larvratstepFE)
larvratfinFE<- lmer(MajMinRat~ParentTrt*JarTrt +(1|FemaleID)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larvratfinFE) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvratfinFE))
qqline(resid(larvratfinFE)) #has relatively long tails, doesn't seem to meet assumption of normality
acf(resid(larvratfinFE)) #this looks fine.
summary(larvratfinFE)
```

Look at the ratio of perimeter to diameter. For a circle, this would be equal to pi
```{r PerimeterToDiameterStat}
PerDiRat<- lmer(PerimDiamRat~SatCalcite_ATDIC*JarSatCalcite_SW+(1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) +(1|CrossID), data=LarvaeDat)
PerDiStep<- step(PerDiRat)
print(PerDiStep)
#final model chosen PerimDiamRat ~ SatCalcite_ATDIC * JarSatCalcite_SW + (1 | FemaleID) + (1 | JarID)+ (1|JarSeatable)
PerDiRatFin<- lmer(PerimDiamRat ~ SatCalcite_ATDIC * JarSatCalcite_SW + (1 | FemaleID) + (1 | JarID) +(1|JarSeatable), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(PerDiRatFin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(PerDiRatFin))
qqline(resid(PerDiRatFin)) #this looks fine
acf(resid(PerDiRatFin))
summary(PerDiRatFin) #significant effect of parent treatment, jar, and interaction, not an additive effect though. 
#larvae in control treatments from both parents are more circular. Larvae are shorter and squatter from OA treatments. If they were elongated, the value would be greater than pi. 

#do a t-test to see if they are significantly different than a circle
contLcontP <- subset(LarvaeDat, ParentTrt =="400")
contLcontP<- subset(contLcontP, JarTrt=="Control")
t.test(contLcontP$PerimDiamRat, mu=pi, alternative="two.sided") # not significantly different than a circle
contLexpP <- subset(LarvaeDat, ParentTrt =="2600")
contLexpP<- subset(contLexpP, JarTrt=="Control")
t.test(contLexpP$PerimDiamRat, mu=pi, alternative="two.sided") # not significantly different than a circle

expLcontP <- subset(LarvaeDat, ParentTrt =="400")
expLcontP<- subset(expLcontP, JarTrt=="Exposed")
t.test(expLcontP$PerimDiamRat, mu=pi, alternative="two.sided") #significantly different than a circle
expLexpP <- subset(LarvaeDat, ParentTrt =="2600")
expLexpP<- subset(expLexpP, JarTrt=="Exposed")
t.test(expLexpP$PerimDiamRat, mu=pi, alternative="two.sided") #significantly different than a circle
```

Look at the ratio of perimeter to diameter with Parent as a covariate and JarTrt as a fixed effect. For a circle, this would be equal to pi
```{r PerimeterToDiameterStatCovariateJarFixed, include=FALSE}
PerDiRatJF<- lmer(PerimDiamRat~SatCalcite_ATDIC*JarTrt+(1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) +(1|CrossID), data=LarvaeDat)
PerDiStepJF<- step(PerDiRatJF)
print(PerDiStepJF)
#final model chosen PerimDiamRat ~ SatCalcite_ATDIC * JarTrt + (1 | FemaleID) + (1 | JarID)
PerDiRatFinJF<- lmer(PerimDiamRat ~ SatCalcite_ATDIC * JarTrt + (1 | FemaleID) + (1 | JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(PerDiRatFinJF) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(PerDiRatFinJF))
qqline(resid(PerDiRatFinJF)) #this looks fine
acf(resid(PerDiRatFinJF))
summary(PerDiRatFinJF)
```

Look at the ratio of perimeter to diameter with Parent and JarTrt as a fixed effects
```{r PerimeterToDiameterStatBothFixedEffects, include=FALSE}
PerDiRatFE<- lmer(PerimDiamRat~ParentTrt*JarTrt+(1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) +(1|CrossID)+(1|TrtTankID), data=LarvaeDat)
PerDiStepFE<- step(PerDiRatFE)
print(PerDiStepFE)
#final model chosen PerimDiamRat ~ SatCalcite_ATDIC * JarTrt + (1 | FemaleID) + (1 | JarID)
PerDiRatFinFE<- lmer(PerimDiamRat ~ ParentTrt * JarTrt + (1 | FemaleID) + (1 | JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(PerDiRatFinFE) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(PerDiRatFinFE))
qqline(resid(PerDiRatFinFE)) #this looks fine
acf(resid(PerDiRatFinFE))
summary(PerDiRatFinFE)
```

```{r EccentricityStat}
larveccen<-lmer(LarvaeEccentricity~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID), data=LarvaeDat)
steppedlarveccen<- step(larveccen)
print(steppedlarveccen)
#final model only had JarTrt*TankTrt, female ID and jarID
larveccenfin<- lmer(LarvaeEccentricity~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larveccenfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larveccenfin))
qqline(resid(larveccenfin)) #has relatively long tails, doesn't meet assumption of normality. 
acf(resid(larveccenfin))
summary(larveccenfin) #significant effect of ParentTrt and interaction
#try a transformation
larveccent<-lmer(LarvaeEccentricity^3~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|JarID) , data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larveccent) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larveccent))
qqline(resid(larveccent)) #this looks much better than untransformed
summary(larveccent) #significant effect of parent treatment and interaction 
```

Analyze eccentricity with parent as a covariate and jar as fixed effect
```{r EccentricityStatParTrtCovariateJarFixed, include=FALSE}
larveccenJF<-lmer(LarvaeEccentricity~SatCalcite_ATDIC*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID), data=LarvaeDat)
steppedlarveccenJF<- step(larveccenJF)
print(steppedlarveccenJF)
#final model
larveccenfinJF<- lmer(LarvaeEccentricity~JarTrt+ (1|FemaleID)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larveccenfinJF) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larveccenfinJF))
qqline(resid(larveccenfinJF)) #has relatively long tails, doesn't meet assumption of normality. 
acf(resid(larveccenfinJF))
summary(larveccenfinJF)
```

Analyze eccentricity with parent and jar as fixed effects
```{r EccentricityStatBothFixed, include=FALSE}
larveccenFE<-lmer(LarvaeEccentricity~ParentTrt*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|TrtTankID)+(1|AccTankID), data=LarvaeDat)
steppedlarveccenFE<- step(larveccenFE)
print(steppedlarveccenFE)
#final model
larveccenfinFE<- lmer(LarvaeEccentricity~ParentTrt*JarTrt+ (1|FemaleID)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larveccenfinFE) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larveccenfinFE))
qqline(resid(larveccenfinFE)) #has relatively long tails, doesn't meet assumption of normality. 
acf(resid(larveccenfinFE))
summary(larveccenfinFE)
```

Test Larvae perimeter
```{r LarvaePerimStat}
ggplot(LarvaeDat, aes(x=LarvaePerimeterum, color=JarTrt))+
  geom_histogram(alpha=0.5, position="identity") +
  facet_grid(~ParentTrt)
#test to see if there is a significant difference
larvperim<-lmer(LarvaePerimeterum~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID), data=LarvaeDat)
steppedlarvperim<- step(larvperim)
print(steppedlarvperim)
#final model chosen by the step function: LarvaePerimeterum~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)
larvperimfin<- lmer(LarvaePerimeterum~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID), data=LarvaeDat)
#got warning message that model failed to converged. I read this stack overflow talking about checking to see if it is acceptable: https://stats.stackexchange.com/questions/97929/lmer-model-fails-to-converge
TestConvergePerim <- with(larvperimfin@optinfo$derivs,solve(Hessian,gradient))
max(abs(TestConvergePerim))
#this value seems small enough to be acceptable
#check the model to see that it meets assumptions
plot(larvperimfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvperimfin))
qqline(resid(larvperimfin)) #has relatively long tails, doesn't meet assumption of normality. 
acf(resid(larvperimfin)) #this looks fine to me.
summary(larvperimfin) #Main effects and interaction are significant. parent is nearly significant. 
```

Analyze larvae perimeter with adult as a covariate and jar as a fixed effect
```{r PerimeterStatAdultCovariateJarFixedEffect, include=FALSE}
larvperimJF<-lmer(LarvaePerimeterum~SatCalcite_ATDIC*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID), data=LarvaeDat)
steppedlarvperimJF<- step(larvperimJF)
print(steppedlarvperimJF)
#final model chosen by the step function: LarvaePerimeterum~SatCalcite_ATDIC*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)
larvperimfinJF<- lmer(LarvaePerimeterum~SatCalcite_ATDIC*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larvperimfinJF) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvperimfinJF))
qqline(resid(larvperimfinJF)) #has relatively long tails, doesn't meet assumption of normality. 
acf(resid(larvperimfinJF)) #this looks fine to me.
summary(larvperimfinJF)
```

Analyze larvae perimeter with adult and jar as a fixed effects
```{r PerimeterStatBothFixedEffects, include=FALSE}
larvperimFE<-lmer(LarvaePerimeterum~ParentTrt*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|TrtTankID)+(1|AccTankID), data=LarvaeDat)
steppedlarvperimFE<- step(larvperimFE)
print(steppedlarvperimFE)
#final model chosen by the step function: LarvaePerimeterum~ParentTrt*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)
larvperimfinFE<- lmer(LarvaePerimeterum~ParentTrt*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larvperimfinFE) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvperimfinFE))
qqline(resid(larvperimfinFE)) #has relatively long tails, doesn't meet assumption of normality. 
acf(resid(larvperimfinFE)) #this looks fine to me.
summary(larvperimfinFE)
```

Analyze survival data
```{r SurvivalStat}
hist(LarvByJarDat$LarvaeSurvived) #data are right skewed so try a log transformation
surv1<- lmer(log(LarvaeSurvived)~SatCalcite_ATDIC*JarSatCalcite_SW+ (JarSatCalcite_SW|CrossID)+(1|FemaleID)+(1|MaleID)+(1|AccTankID), data=LarvByJarDat) #model with random slope and intercept for CrossID
survstep<- step(surv1)
print(survstep)
#final model chosen by step function: log(LarvaeSurvived) ~ JarSatCalcite_SW + (1 | FemaleID) + (1 |CrossID)
survfin<- lmer(log(LarvaeSurvived) ~ JarSatCalcite_SW + (1 | FemaleID) + (1 |CrossID), data=LarvByJarDat)
plot(survfin) #this looks fine to me
qqnorm(resid(survfin))
qqline(resid(survfin)) #I think this is fine
acf(resid(survfin))
summary(survfin)

#repeat without jars that had <1000 
SurvSub<- subset(LarvByJarDat, LarvaeSurvived<1000)
surv1sub<- lmer(log(LarvaeSurvived)~SatCalcite_ATDIC*JarSatCalcite_SW+ (JarSatCalcite_SW|CrossID)+(1|FemaleID)+(1|MaleID)+(1|AccTankID), data=SurvSub) 
survsubstep<- step(surv1sub)
print(survsubstep)
#final model chosen was log(LarvaeSurvived) ~ JarSatCalcite_SW + (1 | CrossID)
survfinsub<- lmer(log(LarvaeSurvived) ~ JarSatCalcite_SW  + (1 |CrossID), data=SurvSub)
plot(survfinsub) #this looks fine to me
qqnorm(resid(survfinsub))
qqline(resid(survfinsub)) #I think this is fine
acf(resid(survfinsub))
summary(survfinsub)
```

Analyze survival data with parent as a fixed effect
```{r SurvivalStatFixedEffect, include=FALSE}
surv1FE<- lmer(RatSurv~ParentTrt+ (1|FemaleID)+(1|MaleID)+(1|AccTankID)+(1|TrtTankID), data=SurvWideDat)
survstepFE<- step(surv1FE)
print(survstepFE)
#final model chosen by step function: RatSurv~ (1|FemaleID), which is the same as survfin from above
survpropFE<- lmer(SurvChange~ParentTrt+ (1|FemaleID)+(1|MaleID)+(1|AccTankID)+(1|TrtTankID), data=SurvWideDat)
survsteppropFE<- step(survpropFE)
print(survsteppropFE)
#final model chosen by step function: SurvChange~ (1|FemaleID), which is the same as survfinprop
```

Cilia extrusion statistics
```{r CiliaExtrusionStat}
#Plot cilia extrusion by size to see if there is a relationship (see Waldbusser et al 2015)
#to do this Elise looked at a subset of Larvae for each jar treatment. 
par(mfcol=c(1,1))
testcil<- read.csv("../data/LarvMorphTest.csv")
testcil$CilExtruded<- as.factor(testcil$CilExtruded)
testcilsub<- subset(testcil, CilExtruded !="")
testcilsub$Flag<- as.factor(testcilsub$Flag)
testcilsub<- subset(testcilsub, Flag !="TRUE")
#take a look at the data
ggplot(testcilsub,
       aes(fill = JarTrt, y = MaxFeretDiamum, x = CilExtruded)) +
  geom_point(aes( shape=JarTrt, colour=JarTrt)) +
  ggtitle("Larvae cilia extrusion by size")+ 
  ylab("Larvae Diameter (um)")+
  facet_grid(~JarTrt)+
  theme_classic()
#test for a relationship between size and cilia extruded within each jar treatment type. 
exp<- subset(testcilsub, JarTrt=="Exposed")
con<- subset(testcilsub, JarTrt=="Control")

expcil1<-glm(CilExtruded~MaxFeretDiamum, data=exp, family=binomial)
summary(expcil1) #no significant relationship between size and cilia extrusion
boxplot(MaxFeretDiamum~CilExtruded,  data=exp)

#now test the control jars for relationship between size and cilia extrusion. 
concil1<-glm(CilExtruded~MaxFeretDiamum, data=con, family=binomial)
summary(concil1) #looks like there is a significant relationship between size and cilia extrusion

boxplot(MaxFeretDiamum~CilExtruded,  data=con)
```

Final Models to Use
```{r FinalModelsWithSameRandomEffects}
summary(larvgrowfin)
LarvMods<- LarvaeDat
predgrow<- predictInterval(larvgrowfin, LarvMods, type="linear.prediction", level=0.95)
colnames(predgrow)<- paste("grow", colnames(predgrow), sep="")
LarvMods<- cbind(LarvMods, predgrow)
Diam<-lmer(LarvaeDiamum~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarID)+(1|JarSeatable), data=LarvaeDat)
#check the final model to see that it meets assumptions
plot(Diam) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(Diam))
qqline(resid(Diam)) #has relatively long tails,but I think this is fine
acf(resid(Diam)) #this looks good 
summary(Diam)
#Conclusion: signficant main effect of Parent and Jar calcite sat, and sig. interaction between parent and Jar saturation states
#add predicted length
#get 95% confidence interval for the linear model
predint <- predictInterval(Diam, LarvMods, type="linear.prediction", level=0.95)
head(predint)
colnames(predint)<- paste("Diam", colnames(predint), sep="")
LarvMods<- cbind(LarvMods, predint)

Area<-lmer(sqrt(LarvaeAreaum2)~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID), data=LarvaeDat)
plot(Area) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(Area))
qqline(resid(Area)) #this seems to have improved it
acf(resid(Area)) #this looks fine to me. 
summary(Area)
predarea <- predictInterval(Area, LarvMods, type="linear.prediction", level=0.95)
colnames(predarea)<- paste("Area", colnames(predarea), sep="")
predarea$Areafit2<- predarea$Areafit^2
predarea$Arealwr2<- predarea$Arealwr^2
predarea$Areaupr2<- predarea$Areaupr^2
LarvMods<- cbind(LarvMods, predarea)

PerimDiam<- lmer(PerimDiamRat~SatCalcite_ATDIC*JarSatCalcite_SW+(1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID) +(1|CrossID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(PerimDiam) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(PerimDiam))
qqline(resid(PerimDiam)) #this looks fine
acf(resid(PerimDiam))
summary(PerimDiam) #crossID doesn't help at all and explains none of the variance
predperimdiam <- predictInterval(PerimDiam, LarvMods, type="linear.prediction", level=0.95)
colnames(predperimdiam)<- paste("PerimDiam", colnames(predperimdiam), sep="")
LarvMods<- cbind(LarvMods, predperimdiam)

Perim<-lmer(LarvaePerimeterum~SatCalcite_ATDIC*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(Perim) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(Perim))
qqline(resid(Perim)) #has relatively long tails, doesn't meet assumption of normality. 
acf(resid(Perim)) #this looks fine to me.
summary(Perim) 
#got warning message that model failed to converge. I read this stack overflow talking about checking to see if it is acceptable: https://stats.stackexchange.com/questions/97929/lmer-model-fails-to-converge
TestConvergePerim <- with(Perim@optinfo$derivs,solve(Hessian,gradient))
max(abs(TestConvergePerim))
#seems small enough to be acceptable
predperim <- predictInterval(Perim, LarvMods, type="linear.prediction", level=0.95)
colnames(predperim)<- paste("Perim", colnames(predperim), sep="")
LarvMods<- cbind(LarvMods, predperim)

#get means for the data
meanJarTrt<- ddply(LarvMods, .(JarTrt, ParentTrt), numcolwise(mean, na.rm=T))
seJarTrt<- ddply(LarvMods, .(JarTrt, ParentTrt), numcolwise(se, na.rm=T))

#predict values for survival
predsurv<- predictInterval(survfin, LarvByJarDat, type="linear.prediction", level=0.95)
colnames(predsurv)<- paste("surv", colnames(predsurv), sep="")
predsurvsub<- predictInterval(survfinsub, SurvSub, type="linear.prediction", level=0.95)
colnames(predsurvsub)<- paste("survsub", colnames(predsurvsub), sep="")
LarvaeDataSurv<- cbind(LarvByJarDat, predsurv)
LarvaeDataSurv$survfitBT<- exp(LarvaeDataSurv$survfit)
LarvaeDataSurv$survlwrBT<- exp(LarvaeDataSurv$survlwr)
LarvaeDataSurv$survuprBT<- exp(LarvaeDataSurv$survupr)
meansurv<- ddply(LarvaeDataSurv, .(JarTrt, ParentTrt), numcolwise(mean, na.rm=T))
sesurv<- ddply(LarvaeDataSurv, .(JarTrt, ParentTrt), numcolwise(se, na.rm=T))
meansurvsub<- ddply(LarvaeDataSurv, .(JarTrt), numcolwise(mean, na.rm=T))
sesurvsub<- ddply(LarvaeDataSurv, .(JarTrt), numcolwise(se, na.rm=T))
```

Create multipanel figure
```{r MultipanelFigure}
par(mfcol=c(1,1))
#this will be the largest in the panel
p1<-ggplot(data = meanJarTrt, aes(x = JarSatCalcite_SW, y = GrowthPerDay, group=ParentTrt, fill=ParentTrt, color=ParentTrt)) +
  geom_point(aes(colour=ParentTrt), size=2)+
  geom_ribbon( aes(ymin = growlwr, ymax = growupr, fill = ParentTrt, color = NULL), alpha = .15) +
     geom_line( aes(y = growfit, color=ParentTrt), size = 1)+
  geom_errorbar(aes(ymin= GrowthPerDay-seJarTrt$GrowthPerDay, ymax=GrowthPerDay+seJarTrt$GrowthPerDay),width=0)+
   geom_errorbarh(aes(xmin= JarSatCalcite_SW-seJarTrt$JarSatCalcite_SW, xmax=JarSatCalcite_SW+seJarTrt$JarSatCalcite_SW),height=0)+ 
  xlim(0.2, 1.2)+
  labs(x =expression(paste(Omega[Calcite])),
                  y = expression(paste("Growth (", mu, "m/day)")), tag="a.") +
  scale_color_manual(values = c("red2", "skyblue2"),  name="",labels= c("\nOA Exposed \nParents\n", "\nControl \nParents\n")) +
  scale_fill_manual(values = c("red2", "skyblue2"), name="", labels= c("\nOA Exposed \nParents\n", "\nControl \nParents\n"))+
 theme(legend.position = "none", legend.key = element_rect(size = 0.5),
        legend.key.size = unit(1.5, 'lines'),panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), panel.background = element_blank(), plot.margin = unit(c(1,1,1,1), "lines"),
        axis.line = element_line(colour = "black"))

#smaller diameter figure
p2<-ggplot(data = meanJarTrt, aes(x = JarSatCalcite_SW, y = LarvaeDiamum, group=ParentTrt, fill=ParentTrt, color=ParentTrt)) +
  geom_point(aes(colour=ParentTrt), size=2)+
  geom_ribbon( aes(ymin = Diamlwr, ymax = Diamupr, fill = ParentTrt, color = NULL), alpha = .15) +
     geom_line( aes(y = Diamfit), size = 1)+
  geom_errorbar(aes(ymin= LarvaeDiamum-seJarTrt$LarvaeDiamum, ymax=LarvaeDiamum+seJarTrt$LarvaeDiamum),width=0.03)+
   geom_errorbarh(aes(xmin= JarSatCalcite_SW-seJarTrt$JarSatCalcite_SW, xmax=JarSatCalcite_SW+seJarTrt$JarSatCalcite_SW),width=0.05, position=position_dodge(0))+ 
  xlim(0.2, 1.2)+
  labs(x =expression(paste(Omega[Calcite])),
                  y = expression(paste("Diameter (", mu, "m)")), tag="c.") +
  scale_color_manual(values = c("red2", "skyblue2"), name="Parent Treatment", labels= c("OA Exposed", "Control")) +
  scale_fill_manual(values = c("red2", "skyblue2"), name="Parent Treatment", labels= c("OA Exposed", "Control"))+
theme(legend.position = "none", panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), panel.background = element_blank(),  plot.margin = unit(c(1,1,1,1), "lines"),
        axis.line = element_line(colour = "black"))

#perimeter data
p3<-ggplot(data = meanJarTrt, aes(x = JarSatCalcite_SW, y = LarvaePerimeterum, group=ParentTrt, fill=ParentTrt, color=ParentTrt)) +
  geom_point(aes(colour=ParentTrt), size=2)+
  geom_ribbon( aes(ymin = Perimlwr, ymax = Perimupr, fill = ParentTrt, color = NULL), alpha = .15) +
     geom_line( aes(y = Perimfit, color=ParentTrt), size = 1)+
  geom_errorbar(aes(ymin= LarvaePerimeterum-seJarTrt$LarvaePerimeterum, ymax=LarvaePerimeterum+seJarTrt$LarvaePerimeterum),width=0)+
   geom_errorbarh(aes(xmin= JarSatCalcite_SW-seJarTrt$JarSatCalcite_SW, xmax=JarSatCalcite_SW+seJarTrt$JarSatCalcite_SW),height=0, position=position_dodge(0))+
  xlim(0.2, 1.2)+
  labs(x =expression(paste(Omega[Calcite])),
                  y = expression(paste("Perimeter (", mu, "m)")), tag="b.")+
  scale_color_manual(values = c("red2","skyblue2"), name="Parent Treatment", labels= c("OA Exposed", "Control"))+
  scale_fill_manual(values = c("red2","skyblue2"), name="Parent Treatment", labels= c("OA Exposed", "Control"))+
theme(legend.position = "none", panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), panel.background = element_blank(), plot.margin = unit(c(1,1,1,1), "lines"),
        axis.line = element_line(colour = "black"))

#surface area figure using back transformed predicted values
p4<-ggplot(data = meanJarTrt, aes(x = JarSatCalcite_SW, y = LarvaeAreaum2, group=ParentTrt, fill=ParentTrt, color=ParentTrt)) +
  geom_point(aes(colour=ParentTrt), size=2)+
  geom_ribbon( aes(ymin = Arealwr2, ymax = Areaupr2, fill = ParentTrt, color = NULL), alpha = .15) +
     geom_line(aes(y = Areafit2))+
  geom_errorbar(aes(ymin= LarvaeAreaum2-seJarTrt$LarvaeAreaum2, ymax=LarvaeAreaum2+seJarTrt$LarvaeAreaum2 ,width=0.03))+
   geom_errorbarh(aes(xmin= JarSatCalcite_SW-seJarTrt$JarSatCalcite_SW, xmax=JarSatCalcite_SW+seJarTrt$JarSatCalcite_SW ,width=0.05), position=position_dodge(0))+ 
  xlim(0.2, 1.2)+
  labs(x =expression(paste(Omega[Calcite])),
                  y = expression(paste("Surface Area (", mu, m^2,")")), tag="d.") +
  scale_color_manual(values = c("red2", "skyblue2"), name="Parent Treatment", labels= c("OA Exposed", "Control")) +
  scale_fill_manual(values = c("red2", "skyblue2"), name="Parent Treatment", labels= c("OA Exposed", "Control"))+
theme(legend.position = "none", panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), panel.background = element_blank(), plot.margin = unit(c(1,1,1,.7), "lines"),
        axis.line = element_line(colour = "black"))

#eccentricity figure
p5<-ggplot(data = meanJarTrt, aes(x = JarSatCalcite_SW, y = PerimDiamRat, group=ParentTrt, fill=ParentTrt, color=ParentTrt)) +
  geom_point(aes(colour=ParentTrt), size=2)+
  geom_ribbon( aes(ymin = PerimDiamlwr, ymax = PerimDiamupr, fill = ParentTrt, color = NULL), alpha = .15) +
     geom_line( aes(y = PerimDiamfit, color=ParentTrt, group=ParentTrt), size = 1)+
  geom_errorbar(aes(ymin= PerimDiamRat-seJarTrt$PerimDiamRat, ymax=PerimDiamRat+seJarTrt$PerimDiamRat),width=0)+
   geom_errorbarh(aes(xmin= JarSatCalcite_SW-seJarTrt$JarSatCalcite_SW, xmax=JarSatCalcite_SW+seJarTrt$JarSatCalcite_SW, height=0))+
  xlim(0.2, 1.2)+
  labs(x =expression(paste(Omega[Calcite])), y = "Perimeter:Diameter", tag="e.") +
  scale_color_manual(values = c("red2","skyblue2"), name="Parent Treatment", labels= c("OA Exposed", "Control"))+
  scale_fill_manual(values = c("red2","skyblue2"), name="Parent Treatment",labels= c("OA Exposed", "Control"))+
  theme(legend.position = "none", panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), panel.background = element_blank(), plot.margin = unit(c(1,1,1,1), "lines"), 
        axis.line = element_line(colour = "black"))
p1
p2
p3
p4
p5
legendmulti <- get_legend(p1 + theme(legend.position="right", plot.margin = unit(c(3,-1,-1,-1), "lines")))
m <- rbind(c(1, 2), c(1, 2),c(3, 4), c(3, 4), c(5,5))
grid.arrange(grobs = list(p1,p3,p2,p4, legendmulti), layout_matrix = m, common.legend=TRUE)
#or this way
n<- rbind(c(1,1,2,2,3,3,4,4), c(1,1,2,2,3,3,4,4), c(NA,NA,NA,5,5,NA,NA,NA))
grid.arrange(grobs = list(p1,p3,p2,p4, legendmulti), layout_matrix = n, common.legend=TRUE)

#or option 3
o<- rbind(c(1,1,2,2), c(3,3,4,4), c(5,5,6,6))
grid.arrange(grobs = list(p1,p3,p2,p4,p5, legendmulti), layout_matrix = o, common.legend=TRUE)
```

Visualize larvae survival
```{r SurvivalVisualization}
ggplot(data = meansurv, aes(x = JarSatCalcite_SW, y = LarvaeSurvived, group=ParentTrt, fill=ParentTrt, color=ParentTrt)) +
  geom_point(aes(colour=ParentTrt), size=2)+
  geom_ribbon( aes(ymin = survlwrBT, ymax = survuprBT, fill = ParentTrt, color = NULL), alpha = .15) +
     geom_line( aes(y = survfitBT, color=ParentTrt), size = 1)+
  geom_errorbar(aes(ymin= LarvaeSurvived-sesurv$LarvaeSurvived, ymax=LarvaeSurvived+sesurv$LarvaeSurvived),width=0)+
   geom_errorbarh(aes(xmin= JarSatCalcite_SW-sesurv$JarSatCalcite_SW, xmax=JarSatCalcite_SW+sesurv$JarSatCalcite_SW),height=0, position=position_dodge(0))+
  labs(x = "Larvae Jar Saturation State (Calcite)", y = "Larvae Survived") +
  scale_color_manual(values = c("red2","skyblue2"), name="Parent Treatment", labels= c("OA Exposed", "Control"))+
  scale_fill_manual(values = c("red2","skyblue2"), name="Parent Treatment", labels= c("OA Exposed", "Control"))+
theme_classic()
ggplot(data = meansurvsub, aes(x = JarSatCalcite_SW, y = LarvaeSurvived)) +
  geom_point(size=2)+
  geom_ribbon( aes(ymin = survlwrBT, ymax = survuprBT, color = NULL), alpha = .15) +
     geom_line( aes(y = survfitBT), size = 1)+
  ylim(c(200,1400))+
  geom_errorbar(aes(ymin= LarvaeSurvived-sesurvsub$LarvaeSurvived, ymax=LarvaeSurvived+sesurvsub$LarvaeSurvived),width=0)+
   geom_errorbarh(aes(xmin= JarSatCalcite_SW-sesurvsub$JarSatCalcite_SW, xmax=JarSatCalcite_SW+sesurvsub$JarSatCalcite_SW),height=0, position=position_dodge(0))+
  labs(x =expression(paste(Omega[Calcite])),
                  y = expression(paste("Number of Surviving Larvae")))+
theme_classic()
#this figure doesn't account for parent treatment because it wasn't significant

boxplot(RatSurv~as.factor(ParentTrt), data=SurvWideDat)
boxplot(SurvChange~as.factor(ParentTrt), data= SurvWideDat)
#make reaction norms for survival color will be parent treatment. X will be jar treatment. Each cross will be its own line
ggplot(SurvDat,aes(x=JarTrt, y=LarvaeSurvived, color = ParentTrt, group= CrossID)) +
    geom_point(aes(color=ParentTrt))+
  geom_line(aes(color=ParentTrt, group=CrossID))+
  geom_errorbar(aes(ymin= LarvaeSurvived-SELarvaeSurvived, ymax=LarvaeSurvived+SELarvaeSurvived),width=0.05, position=position_dodge(0))+
    labs(x = "Larvae Treatment", y = "Number of Larvae Survived") +
  scale_color_manual(values = c("skyblue2", "red2")) +
  theme_classic()
ggplot(SurvWideDat,
       aes(y = SurvChange, x = ParentTrt)) +
  geom_point() +
  ggtitle("Larvae survival by parent treatment")+ 
  ylab("Proportion change in the number of larvae that survived")+
  theme_classic()
meansurv2<- ddply(SurvWideDat, .(ParentTrt), numcolwise(mean, na.rm=T))
sesurv<- ddply(SurvWideDat, .(ParentTrt), numcolwise(se, na.rm=T))
#make a shorter name for the CrossIDs in SurvWideDat
SurvWideDat$Cross<- substr(SurvWideDat$CrossID, 1, 9)
ggplot(data = SurvWideDat, aes(x = Cross, y = RatSurv, fill=ParentTrt)) +
  geom_bar(stat="identity", position="dodge")+
  labs(x = "Cross ID", y = "Proportion of Larvae that Survived") +
  scale_fill_manual(name = "Parent Treatment", labels = c("Control", "OA Exposed"), values = c("skyblue2", "red2")) +
  theme(axis.text.x=element_text(angle=90))
ggplot(data = meansurv2, aes(x = ParentTrt, y = RatSurv)) +
  geom_bar(stat="identity", position="dodge")+
  labs(x = "Parent Treatment", y = "Proportion of Larvae that Survived")+
  geom_errorbar(aes(ymin= RatSurv-sesurv$RatSurv, ymax=RatSurv+sesurv$RatSurv),width=0.2, position=position_dodge(.9))+
  theme_classic()
```

```{r GetDescriptiveStats}
mgpd<- subset(meanJarTrt, select= c(JarTrt, ParentTrt, GrowthPerDay, LarvaeDiamum, LarvaeAreaum2, LarvaePerimeterum, PerimDiamRat))
(mgpd$GrowthPerDay[3]-mgpd$GrowthPerDay[4])/mgpd$GrowthPerDay[4]
(mgpd$GrowthPerDay[1]-mgpd$GrowthPerDay[2])/mgpd$GrowthPerDay[2]
(mgpd$LarvaeDiamum[3]-mgpd$LarvaeDiamum[4])/mgpd$LarvaeDiamum[4]
(mgpd$LarvaeDiamum[1]-mgpd$LarvaeDiamum[2])/mgpd$LarvaeDiamum[2]
mean(mgpd$LarvaeDiamum[1:2])/mean(mgpd$LarvaeDiamum[3:4])
(mgpd$LarvaeAreaum2[3]-mgpd$LarvaeAreaum2[4])/mgpd$LarvaeAreaum2[4]
(mgpd$LarvaeAreaum2[1]-mgpd$LarvaeAreaum2[2])/mgpd$LarvaeAreaum2[2]
mean(mgpd$LarvaeAreaum2[1:2])/mean(mgpd$LarvaeAreaum2[3:4])

(mgpd$LarvaePerimeterum[3]-mgpd$LarvaePerimeterum[4])/mgpd$LarvaePerimeterum[4]
(mgpd$LarvaePerimeterum[1]-mgpd$LarvaePerimeterum[2])/mgpd$LarvaePerimeterum[2]
mean(mgpd$LarvaePerimeterum[1:2])/mean(mgpd$LarvaePerimeterum[3:4])

(meansurvsub$LarvaeSurvived[1]-meansurvsub$LarvaeSurvived[2])/meansurvsub$LarvaeSurvived[1]
#exponentiate the slope of the best fit model. and that is the ratio of the difference. 
exp(0.19820)

```

Export Figures
```{r ExportFigures, include=FALSE}
plots.dir.path<- list.files(tempdir(), pattern="rs-graphics", full.names= TRUE)
plots.png.paths<- list.files(plots.dir.path, pattern=".png", full.names=TRUE)
file.copy(from=plots.png.paths, to="../results")
```

Look at weight per larvae
```{r WeightPerLarvaeData, include=FALSE}
#we can check these data in a few ways: 1) look to see if replicate jars have similar weights; 2) check to see if first and second filters measured come up with similar values for weight per larva
#Start with 1) 
#start by using all data: 
boxplot(MeanWtPerLarvae~CrossID*JarTrt, data=LarvByJarDat)
#looks like some of the crosses have high variance. could we just eliminate those? 
meanwtall<- ddply(LarvByJarDat, .(CrossID,ParentTrt, JarTrt, FemaleID, MaleID), numcolwise(mean, na.rm=T))
sewtall<- ddply(LarvByJarDat, .(CrossID, ParentTrt, JarTrt, FemaleID, MaleID), numcolwise(se, na.rm=T))
ggplot(data = meanwtall, aes(x = CrossID, y = MeanWtPerLarvae)) + 
  geom_errorbar(aes(ymin= MeanWtPerLarvae-sewtall$MeanWtPerLarvae, ymax=MeanWtPerLarvae+sewtall$MeanWtPerLarvae),width=0.2)+
  geom_point(aes(colour=JarTrt)) + 
  labs(x = "Cross ID", y = "Mean Weight Per Larvae ug") + 
   theme_classic()
#let's see if removing filters with fewer than 100 individuals helps this out and those with crystals.
subdat<- LarvByJarDat
subdat$MeanWtPerLarvae<- ifelse(subdat$F1Crystals=="FALSE", paste(subdat$MeanWtPerLarvae), paste("NA"))
subdat$MeanWtPerLarvae<- ifelse(subdat$F1LarvaeCount>100, paste(subdat$MeanWtPerLarvae), paste("NA"))
subdat$MeanWtPerLarvae<- ifelse(subdat$JarID!="J004_B1", paste(subdat$MeanWtPerLarvae), paste("NA"))#this is the only F2 filter after subsetting that has <100 larvae
subdat$JarSeatable<- as.factor(subdat$JarSeatable)
subdat$MeanWtPerLarvae<- as.numeric(subdat$MeanWtPerLarvae)
subdat$JarTrt<- as.factor(subdat$JarTrt)
subdat$LarvDensity<- subdat$MeanWtPerLarvae/subdat$LarvaeAreaum2
boxplot(MeanWtPerLarvae~CrossID*JarTrt, data=subdat)
#looks like some of the crosses have high variance. could we just eliminate those? 
submeanwtall<- ddply(subdat, .(Block, ParentTrt, JarTrt, CrossID, FemaleID, MaleID), numcolwise(mean, na.rm=T))
subsewtall<- ddply(subdat, .(Block, ParentTrt, JarTrt, CrossID, FemaleID, MaleID), numcolwise(se, na.rm=T))
ggplot(data = submeanwtall, aes(x = CrossID, y = MeanWtPerLarvae)) + 
  geom_errorbar(aes(ymin= MeanWtPerLarvae-subsewtall$MeanWtPerLarvae, ymax=MeanWtPerLarvae+subsewtall$MeanWtPerLarvae),width=0.2)+
  geom_point(aes(colour=JarTrt)) + 
  labs(x = "Cross ID", y = "Mean Weight Per Larvae ug") + 
  theme(axis.text.x = element_text(angle = 90))
#there are some crosses that only have a single point. Those should be removed. Do this by looking at se
#first join SE to the dataset
colnames(subsewtall)<- paste("SE", colnames(subsewtall), sep="")
#merge the standard error data to the SurvDat dataframe
SubWt<- merge(submeanwtall, subsewtall, by.x= c("CrossID", "JarTrt"), by.y= c("SECrossID", "SEJarTrt"))
relcrosses<- SubWt
relcrosses$MeanWtPerLarvae<- ifelse(relcrosses$SEMeanWtPerLarvae>0, paste(relcrosses$MeanWtPerLarvae), paste("NA"))
relcrosses$MeanWtPerLarvae<- as.numeric(relcrosses$MeanWtPerLarvae)
ggplot(data = relcrosses, aes(x = CrossID, y = MeanWtPerLarvae)) + 
  geom_errorbar(aes(ymin= MeanWtPerLarvae-SEMeanWtPerLarvae, ymax=MeanWtPerLarvae+SEMeanWtPerLarvae),width=0.2)+
  geom_point(aes(colour=JarTrt)) + 
  labs(x = "Cross ID", y = "Mean Weight Per Larvae ug") + 
  theme(axis.text.x = element_text(angle = 90))
boxplot(SEMeanWtPerLarvae~JarTrt, data= relcrosses)
#I don't know how to choose which ones to include. For now, let's say if the SE is greater than mean(MeanWtPerLarvae) then I will remove it
MeanWtVal<- mean(relcrosses$MeanWtPerLarvae, na.rm=TRUE)
relcrosses$MeanWtPerLarvae<- ifelse(relcrosses$SEMeanWtPerLarvae<= MeanWtVal, paste(relcrosses$MeanWtPerLarvae), paste("NA"))
relcrosses$MeanWtPerLarvae<- as.numeric(relcrosses$MeanWtPerLarvae)
boxplot(MeanWtPerLarvae~ParentTrt*as.factor(JarTrt), data=relcrosses)
#now try to create a linear model 
Density1<- lm(LarvDensity~ParentTrt*as.factor(JarTrt), data=relcrosses)
summary(Density1)
#but from the plot perhaps something is relevant for the difference going from control to exposed conditions. 
#use spread function
relcrossestowide<- unite(relcrosses, Value, c(MeanWtPerLarvae, SEMeanWtPerLarvae, LarvDensity, SELarvDensity), sep="_")
relcrossestowide<- subset(relcrossestowide, select=c("CrossID", "Value", "JarTrt", "ParentTrt", "MaleID", "FemaleID"))
#use spread function in tidyr to make the column in long form vs. wide form
relwide<- spread(relcrossestowide, JarTrt, value= Value)
#now split the mean and SE apart again
relwide<- separate(relwide, Control, into= c("MeanWtPerLarvaeCon", "SEMeanWtPerLarvaeCon", "LarvDensityCon", "SELarvDensityCon"), sep="_")
relwide<- separate(relwide, Exposed, into= c("MeanWtPerLarvaeExp", "SEMeanWtPerLarvaeExp", "LarvDensityExp", "SELarvDensityExp"), sep="_")
#make the means and ses numeric
relwide$MeanWtPerLarvaeCon<- as.numeric(relwide$MeanWtPerLarvaeCon)
relwide$MeanWtPerLarvaeExp<- as.numeric(relwide$MeanWtPerLarvaeExp)
relwide$SEMeanWtPerLarvaeCon<- as.numeric(relwide$SEMeanWtPerLarvaeCon)
relwide$SEMeanWtPerLarvaeExp<- as.numeric(relwide$SEMeanWtPerLarvaeExp)
relwide$LarvDensityCon<- as.numeric(relwide$LarvDensityCon)
relwide$SELarvDensityCon<- as.numeric(relwide$SELarvDensityCon)
relwide$LarvDensityExp<- as.numeric(relwide$LarvDensityExp)
relwide$SELarvDensityExp<- as.numeric(relwide$SELarvDensityExp)
#get the difference from control to OA
relwide$WtDiff<- (relwide$MeanWtPerLarvaeExp-relwide$MeanWtPerLarvaeCon)
plot(WtDiff~ParentTrt, data=relwide)
boxplot(WtDiff~FemaleID, data=relwide)
abline(a=0, b=0)
relwide$DensDiff<- (relwide$LarvDensityExp-relwide$LarvDensityCon)
plot(DensDiff~ParentTrt, data=relwide)
#now try a linear model. 
Dens1<- lm(DensDiff~ParentTrt, data=relwide)
summary(Dens1)
#Look at the difference between F1 and F2 measurements, so subset the data for only those with F2 measurements
#get the mean of weight per larva 
mean(LarvByJarDat$MeanWtPerLarvae)
F1F2<- subset(LarvByJarDat, F2WtPerLarvae >0)
F1F2<- subset(F1F2, F1WtPerLarvae>0)
F1F2$JarTrt<- as.factor(F1F2$JarTrt)
F1F2$F1FilterID<- as.factor(F1F2$F1FilterID)
F1F2$F2FilterID<- as.factor(F1F2$F2FilterID)
F1F2$ParentTrt<- as.factor(F1F2$ParentTrt)
F1F2$CrossID<- as.factor(F1F2$CrossID)
F1F2$TrtTankID<- as.factor(F1F2$TrtTankID)
F1F2$JarSeatable<- as.factor(F1F2$JarSeatable)
F1F2$MaleID<- as.factor(F1F2$MaleID)
F1F2$FemaleID<- as.factor(F1F2$FemaleID)
F1F2$JarID<-droplevels(F1F2)$JarID
F1F2<- subset(F1F2, select=c("JarID", "F1WtPerLarvae", "F1LarvaeCount", "F2WtPerLarvae", "F2LarvaeCount", "F1FilterID", "F2FilterID", "JarTrt", "ParentTrt", "CrossID", "TrtTankID", "JarSeatable", "MaleID", "FemaleID", "F1Crystals", "MeanWtPerLarvae"))
F1F2$LarvWtDiff<- abs(F1F2$F1WtPerLarvae-F1F2$F2WtPerLarvae)
F1F2$LarvWtDiffProp<- (F1F2$LarvWtDiff)/ (F1F2$F2WtPerLarvae)
F1F2$TotalLarvCount<- F1F2$F1LarvaeCount+F1F2$F2LarvaeCount
F1F2$LarvaeCountDiff<- abs(F1F2$F1LarvaeCount- F1F2$F2WtPerLarvae)
F1F2sub<- subset(F1F2, F1LarvaeCount>100)
F1F2sub<- subset(F1F2sub, F2LarvaeCount>100)
F1F2sub<- subset(F1F2sub, F1Crystals=="FALSE")
F1F2sub$JarID<-droplevels(F1F2sub)$JarID
#Now look at the differences.In weights
mean(F1F2sub$F1WtPerLarvae)
mean(F1F2sub$F2WtPerLarvae)
plot(LarvWtDiff~TotalLarvCount, data=F1F2sub, col="black", pch=19) 
abline(a= 1.59e-06, b=0, col="blue")
abline(a=(2* 1.59e-06), b=0, col="red")
plot(F1LarvaeCount~LarvWtDiffProp, data=F1F2sub, pch=19)
points(F2LarvaeCount~LarvWtDiffProp, data=F1F2sub, pch=19, col="orange")
#maybe I should just remove those with larvae counts <200
F200sub<- subset(F1F2, F1LarvaeCount>200)
F200sub<- subset(F200sub, F2LarvaeCount>200)
F200sub$JarID<-droplevels(F200sub)$JarID
mean(F200sub$F1WtPerLarvae)
mean(F200sub$F2WtPerLarvae)
plot(LarvWtDiff~TotalLarvCount, data=F200sub, col="black", pch=19) 
plot(F1LarvaeCount~LarvWtDiffProp, data=F200sub, pch=19)
points(F2LarvaeCount~LarvWtDiffProp, data=F200sub, pch=19, col="orange")
plot(F1WtPerLarvae~JarID, data=F200sub)
points(F2WtPerLarvae~JarID, data=F200sub)
points(LarvWtDiff~JarID, data=F200sub, col="red", pch=19) 
boxplot(LarvWtDiffProp~ParentTrt*JarTrt, data=F1F2sub)
#I'm going to just look at the jars that had proportion difference <1
Less1<- subset(F1F2, LarvWtDiffProp<1)
boxplot(MeanWtPerLarvae~ParentTrt*JarTrt, data=Less1)
boxplot(LarvWtDiffProp~ParentTrt*JarTrt, data=Less1)
Wtdiff<- lm(LarvWtDiff~TotalLarvCount, data=F1F2)
par(mfcol=c(2,2))
plot(Wtdiff)
par(mfcol=c(1,1))
summary(Wtdiff)
#Use the Wt diff model to decide which filters to include
#first get the average larva weight overall: 
mean(F1F2$MeanWtPerLarvae)
boxplot(F1WtPerLarvae~ParentTrt+JarTrt, data=LarvByJarDat)
#I don't know whether or not to trust these data. Let's look at F1 and F2 together
Calcmean<- ddply(LarvByJarDat, .(ParentTrt, JarTrt), numcolwise(mean, na.rm=TRUE))
Calcse<- ddply(LarvByJarDat, .(ParentTrt, JarTrt), numcolwise(se, na.rm=TRUE))
ggplot(data = Calcmean, aes(x = ParentTrt , y = F1WtPerLarvae, fill=JarTrt)) +
  geom_bar(stat="identity", aes(fill=JarTrt), position="dodge")+
  labs(x = "Parent Treatment", y = "Weight per Larvae") +
  scale_fill_manual(name = "Larvae Treatment", labels = c("Control", "OA Exposed"), values = c("skyblue2", "red2")) +
  geom_errorbar(aes(ymin= F1WtPerLarvae-Calcse$F1WtPerLarvae, ymax=F1WtPerLarvae+Calcse$F1WtPerLarvae),width=0.2, position=position_dodge(.9))+
  theme_classic()
calcexp<- subset(LarvByJarDat, JarTrt=="Exposed")
calcexp$JarTrt<- as.factor(calcexp$JarTrt)
calccon<- subset(LarvByJarDat, JarTrt=="Control")
calccon$JarTrt<- as.factor(calccon$JarTrt)
plot(F1WtPerLarvae~F1LarvaeCount, data=calccon, col="blue", pch=19)
points(F1WtPerLarvae~F1LarvaeCount, data=calcexp, col="red", pch=19)
summary(lm(F1WtPerLarvae~F1LarvaeCount+JarTrt+ParentTrt, data=LarvByJarDat))
```
Do simple t-tests on the means for oa-exposed and control exposed parents within each larvae treatment
```{r t-tests}
#get Bonferoni adjusted p-crit: 
0.05/10

Lexp<- subset(LarvaeDat, JarTrt=="Exposed")
Lcon<- subset(LarvaeDat, JarTrt=="Control")

t.test(GrowthPerDay~ParentTrt, data=Lexp)
t.test(GrowthPerDay~ParentTrt, data=Lcon)

t.test(LarvaeDiamum~ParentTrt, data=Lexp)
t.test(LarvaeDiamum~ParentTrt, data=Lcon)#not significantly different for control larvae

t.test(LarvaeAreaum2~ParentTrt, data=Lexp)
t.test(LarvaeAreaum2~ParentTrt, data=Lcon) #not signficantly different for control larvae

t.test(LarvaePerimeterum~ParentTrt, data=Lexp)
t.test(LarvaePerimeterum~ParentTrt, data=Lcon) #not significantly different for control larvae

t.test(PerimDiamRat~ParentTrt, data=Lexp)
t.test(PerimDiamRat~ParentTrt, data=Lcon)#not significantly different for control larvae
```

