---
title: "Larvae Biomineralization Block 1"
output: html_document
---

```{r LoadLibraries, include=FALSE}
library(knitr)
library(cowplot)
library(plyr)
library(tidyverse)
library(sciplot)
library(reshape2)
library(lme4)
library(lmerTest)
library(blme)
library(grid)
library(data.table)
library(gridExtra)
library(lattice)
library(optimx)
library(car)
library(Hmisc)
library(DescTools)
library(corrplot)
library(emmeans)
library(chron)
library(GLMMadaptive)
library(calibrate)
library(merTools)
library(bootpredictlme4)
library(visreg)

```

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

```{r Create Functions}
coefvariation<- function(x){
  sd(x,na.rm=T)/abs(mean(x, na.rm=T))
}

logitTransform<- function(p){
  log(p/(1-p))
}
```

## Load data
Name the data you export from FileMaker Pro by their exact table names and save them as CSVs, e.g. Larvae_Morphology.csv
```{r LoadData}
#setwd("~/Repos/LarvaeTransGen2018")
setwd("~/R/GitHub/LarvaeTransGen2018/data") #Elise's working directory
#upload all of the data tables for the larvae experiment
Larvae_Morphology <- read.csv("../data/Larvae_morphology.csv") #contains data from CellProfiler from the larvae outlines
Barcode_Jar <- read.csv("../data/Barcode_Jar.csv") #contains data on CrossID, seatable, and whether or not larvae were present
Block_ID <- read.csv("../data/Block_ID.csv") #contains data on the block
Cross<- read.csv("../data/Cross.csv") #contains data on the female and male IDs for the cross and whether the cross was for QG or Meth
Fert_QG<- read.csv("../data/Fert_QG.csv") #contains data on fertilization counts, the JarIDs for the crosses
Header_WC<- read.csv("../data/Header_WC.csv") #contains data on the header tanks used to fill the jars
Larvae_Calcification <- read.csv("../data/Larvae_Calcification.csv") #contains data on filter weights and counts
Larvae_Calcification<- subset(Larvae_Calcification, JarID != "")
Larvae_Counts <- read.csv("../data/Larvae_Counts.csv") #contains data on sedgewick rafter larvae counts
Larvae_WC <- read.csv("../data/Larvae_WC.csv") #contains data on larvae jar water chemistry
Parent_ID <- read.csv("../data/Parent_ID.csv") #contains data on the adults at the time of shucking. Includes Parent ID assignments
Larvae_Cilia<- read.csv("../data/Larvae_Cilia.csv") #contains data on cilia extrusion of the larvae photographed for morphology
Egg_Morphology<- read.csv("../data/Egg_Morphology.csv") #contains data on egg sizes from CellProfiler
Egg_Morphology$EggPerimeterum<- Egg_Morphology$EggPerimeterPix/8.892  #add a column to convert egg perimeter to um based on the scaled slide
Egg_Morphology$EggPerimDiamRat<- Egg_Morphology$EggPerimeterum/Egg_Morphology$EggDiamum #add a column for the ratio of perimeter to diameter
Adult_Sample<- read.csv("../data/Adult_Sample.csv") #contains data on adult oysters at the time of collection
WC_Standard<- read.csv("../data/WC_Standard.csv") #contains data on the standards used for water chemistry
Tank_WC<- read.csv("../data/Tank_WC.csv") #contains data on the adult tank water chemistry
CiliaArea<- read.csv("../data/Cilia_AreaLength.csv")
```

Make fields factors: 
```{r MakeFieldsFactors}
Adult_Sample$AccTankID<- factor(Adult_Sample$AccTankID)
Adult_Sample$TrtTankID<- factor(Adult_Sample$TrtTankID)
Barcode_Jar$JarSeatable<- factor(Barcode_Jar$JarSeatable)
Barcode_Jar<- subset(Barcode_Jar, JarSeatable =="2" | JarSeatable =="4" | JarSeatable =="6")
Tank_WC$ParentTrt<- factor(Tank_WC$ParentTrt)
Tank_WC$Tank<- factor(Tank_WC$Tank)
WC_Standard$CRM<- factor(WC_Standard$CRM)
Larvae_WC$DateFilter<- factor(Larvae_WC$DateFilter)
CiliaArea$LarvaeMorphID<- factor(CiliaArea$LarvaeMorphID)
CiliaArea$CilType<- factor(CiliaArea$CilType)
Larvae_Morphology$LarvaeMorphID<- factor(Larvae_Morphology$LarvaeMorphID)
```

Look at the water chemistry data for adult tanks
```{r AdultTankChemistry}
#subset the tank data to only include B1 
Tank_WCB1<- subset(Tank_WC, UseForCalcite == "TRUE")
Tank_WCB1$Tank<-droplevels(Tank_WCB1)$Tank #drop unused levels
Tank_WCB1$ParentTrt<- droplevels(Tank_WCB1)$ParentTrt
#test to see if tanks are significantly different. 
boxplot(SatArag_ATDIC~Tank, data=Tank_WCB1, na.omit=TRUE)
#check for significant differences in saturation state between tanks within treatments
Ca1<- lmer(SatArag_ATDIC~ParentTrt + (1|Tank), data=Tank_WCB1)
Ca2<- lmer(SatArag_ATDIC~ 1 + (1|Tank), data= Tank_WCB1)
anova(Ca1, Ca2) #select Ca1
Ca3<- lm(SatArag_ATDIC~ParentTrt, data=Tank_WCB1)
anova(Ca1, Ca3) #select Ca3
#check assumptions
par(mfcol=c(2,2))
plot(Ca3) #seems to meet assumptions of homoscedasticity and linearity. 
shapiro.test(resid(Ca3)) #meets assumption of normality. 
par(mfcol=c(1,1))
acf(resid(Ca3)) #this looks good. 
summary(Ca3) #There is a significant effect of ParentTrt, but not tank according to best model. 
```

## Get Tank Means
```{r TankMeans}
#create dataframe that has the mean values for each tank
TankMeans<- ddply(Tank_WCB1, .(Tank, Block, ParentTrt), numcolwise(mean, na.rm=T))#can't use this for pH and salinity, remember, it is just for carb chem
TankSes<- ddply(Tank_WCB1, .(Tank, Block, ParentTrt), numcolwise(se, na.rm=T))#can't use this for pH and salinity, remember, it is just for carb chem
```

## Carbonate Chemistry Summary Table for Tanks
Compile all of the data that I will need for the summary table. 
```{r CarbChemSummaryTable}
#get parameters for the carb chem summary table
TankCarbTrtMeans<- ddply(Tank_WCB1, .(Block, ParentTrt), numcolwise(mean, na.rm=T))
TankCarbTrtMeans[,c(24,27,35,42:45)]<- round(TankCarbTrtMeans[,c(24,27,35,42:45)]) #round to the correct number of decimal places. 
TankCarbTrtMeans[,c(46,47)]<- round(TankCarbTrtMeans[,c(46,47)],3) #round saturation states to the correct number of decimal places. 
TankCarbTrtSD<- ddply(Tank_WCB1, .(Block, ParentTrt), numcolwise(sd, na.rm=T))
colnames(TankCarbTrtSD)<- paste(colnames(TankCarbTrtSD), "SD", sep="_")
TankCarbTrtSD[,c(24,27,35,42:45)]<- round(TankCarbTrtSD[,c(24,27,35,42:45)]) #round to the correct number of decimal places. 
TankCarbTrtSD[,c(46,47)]<- round(TankCarbTrtSD[,c(46,47)],3) #round saturation states to the correct number of decimal places. 
TankCarbTrtRangeMin<- ddply(Tank_WCB1, .(Block, ParentTrt), numcolwise(min, na.rm=T))
colnames(TankCarbTrtRangeMin)<- paste(colnames(TankCarbTrtRangeMin),"min", sep="_")
TankCarbTrtRangeMin[,c(24,27,35,42:45)]<- round(TankCarbTrtRangeMin[,c(24,27,35,42:45)]) #round to the correct number of decimal places. 
TankCarbTrtRangeMin[,c(46,47)]<- round(TankCarbTrtRangeMin[,c(46,47)],3) #round saturation states to the correct number of decimal places. 
TankCarbTrtRangeMax<- ddply(Tank_WCB1, .(Block, ParentTrt), numcolwise(max, na.rm=T))
colnames(TankCarbTrtRangeMax)<- paste(colnames(TankCarbTrtRangeMax),"max", sep="_")
TankCarbTrtRangeMax[,c(24,27,35,42:45)]<- round(TankCarbTrtRangeMax[,c(24,27,35,42:45)]) #round to the correct number of decimal places. 
TankCarbTrtRangeMax[,c(46,47)]<- round(TankCarbTrtRangeMax[,c(46,47)],3) #round saturation states to the correct number of decimal places. 

#Tank Sal and pH SW parameters
TankSalpHTemp<- filter(Tank_WC, Block=="B1") %>% 
  filter(ParentTrt =="2600" | ParentTrt=="400")
TankSalpHTempMean<- ddply(TankSalpHTemp, .(Block, ParentTrt), numcolwise(mean, na.rm=T))
TankSalpHTempMean[,c(15,17)]<- round(TankSalpHTempMean[,c(15,17)], 2) #round to the correct number of decimal places. 
TankSalpHTempMean[,c(19,20)]<- round(TankSalpHTempMean[,c(19,20)],1) #round saturation states to the correct number of decimal places. 
TankSalpHTempSD<- ddply(TankSalpHTemp, .(Block, ParentTrt), numcolwise(sd, na.rm=T))
colnames(TankSalpHTempSD)<- paste(colnames(TankSalpHTempSD), "SD", sep="_")
TankSalpHTempSD[,c(15,17)]<- round(TankSalpHTempSD[,c(15,17)], 2) #round to the correct number of decimal places. 
TankSalpHTempSD[,c(19,20)]<- round(TankSalpHTempSD[,c(19,20)],1) #round saturation states to the correct number of decimal places. 
TankSalpHTempMin<- ddply(TankSalpHTemp, .(Block, ParentTrt), numcolwise(min, na.rm=T))
colnames(TankSalpHTempMin)<- paste(colnames(TankSalpHTempMin),"min", sep="_")
TankSalpHTempMin[,c(15,17)]<- round(TankSalpHTempMin[,c(15,17)], 2) #round to the correct number of decimal places. 
TankSalpHTempMin[,c(19,20)]<- round(TankSalpHTempMin[,c(19,20)],1) #round saturation states to the correct number of decimal places. 
TankSalpHTempMax<- ddply(TankSalpHTemp, .(Block, ParentTrt), numcolwise(max, na.rm=T))
colnames(TankSalpHTempMax)<- paste(colnames(TankSalpHTempMax),"max", sep="_")
TankSalpHTempMax[,c(15,17)]<- round(TankSalpHTempMax[,c(15,17)], 2) #round to the correct number of decimal places. 
TankSalpHTempMax[,c(19,20)]<- round(TankSalpHTempMax[,c(19,20)],1) #round saturation states to the correct number of decimal places.
#subset the salpHTemp data: 
TankSalpHTempMean<-subset(TankSalpHTempMean, select= c("ParentTrt","pHNBS", "pHSW", "salinity", "temperature"))
TankSalpHTempSD<- subset(TankSalpHTempSD, select= c("ParentTrt_SD", "pHNBS_SD", "pHSW_SD", "salinity_SD", "temperature_SD"))
TankSalpHTempMin<- subset(TankSalpHTempMin, select= c("ParentTrt_min", "pHNBS_min", "pHSW_min", "salinity_min", "temperature_min"))
TankSalpHTempMax<- subset(TankSalpHTempMax, select= c("ParentTrt_max", "pHNBS_max", "pHSW_max", "salinity_max", "temperature_max"))
#make a table to get carb chem params from. 
CarbChemSummary<- merge(TankCarbTrtMeans, TankCarbTrtSD, by.x="ParentTrt", by.y="ParentTrt_SD") %>% 
  merge(TankCarbTrtRangeMin, by.x="ParentTrt", by.y="ParentTrt_min") %>% 
  merge(TankCarbTrtRangeMax, by.x="ParentTrt", by.y="ParentTrt_max") %>% 
  subset(select= c("ParentTrt", "pCO2_ATDIC", "AT_Calc_SW", "CO3_ATDIC", "HCO3_ATDIC", "CO2_ATDIC", "SatCalcite_ATDIC", "SatArag_ATDIC", "DIC_corr", "Alk_corr", "pCO2_ATDIC_SD", "AT_Calc_SW_SD", "CO3_ATDIC_SD", "HCO3_ATDIC_SD", "CO2_ATDIC_SD", "SatCalcite_ATDIC_SD", "SatArag_ATDIC_SD", "DIC_corr_SD", "Alk_corr_SD", "pCO2_ATDIC_min", "AT_Calc_SW_min", "CO3_ATDIC_min", "HCO3_ATDIC_min", "CO2_ATDIC_min", "SatCalcite_ATDIC_min", "SatArag_ATDIC_min", "DIC_corr_min", "Alk_corr_min", "pCO2_ATDIC_max", "AT_Calc_SW_max", "CO3_ATDIC_max", "HCO3_ATDIC_max", "CO2_ATDIC_max", "SatCalcite_ATDIC_max", "SatArag_ATDIC_max", "DIC_corr_max", "Alk_corr_max")) %>% 
 merge(TankSalpHTempMean, by="ParentTrt") %>% 
  merge(TankSalpHTempSD, by.x="ParentTrt", by.y="ParentTrt_SD") %>% 
  merge(TankSalpHTempMin, by.x="ParentTrt", by.y="ParentTrt_min") %>% 
  merge(TankSalpHTempMax, by.x="ParentTrt", by.y="ParentTrt_max") %>% 
  t() %>% 
  as.data.frame()
colnames(CarbChemSummary)<- c("AdultExposed", "AdultControl")
write.csv(CarbChemSummary,"~/R/GitHub/LarvaeTransGen2018/data/TankCarbChemSummary.csv")
```

# Join the data into dataframe for Egg analysis
```{r JoinDataForEggAnalysis}
#First dataframe will be for examining egg morphology for all eggs traced; multiple values for each adult
AdultSampParID<- merge(x=Adult_Sample, y=Parent_ID, by="AdultID", all.x=TRUE)
AdultSampParID<- merge(x=AdultSampParID, y=TankMeans, by.x="TrtTankID", by.y="Tank", all.x=TRUE)
AdultSampParID<- subset(AdultSampParID, ParentBlockID !="NA")
AdultSampParID<- subset(AdultSampParID, Block =="B1")
MaleInfo<- subset(AdultSampParID, Sex=="M")
FemaleInfo<- subset(AdultSampParID, Sex=="F")
colnames(MaleInfo)<- paste("Male", colnames(MaleInfo), sep="")
Eggs<- merge(x=Egg_Morphology, y=FemaleInfo, by.x="FemaleID", by.y="ParentBlockID", all.x=TRUE)
Eggs$FemaleID<- as.factor(Eggs$FemaleID)
#subset the data to only include B1
EggsB1<- subset(Eggs, Block =="B1")
```

Visualize the data for eggs to start
```{r EggVisualization}
#get the original range of the eggs
range(EggsB1$EggDiamum)
#look at the egg measurements to see outliers
boxplot(EggDiamum~FemaleID, data=EggsB1)
boxplot(EggDiamum~ParentTrt, data=EggsB1)
```

Remove egg outliers
```{r EggOutlierRemoval}
#check for outliers using code from https://www.r-bloggers.com/identify-describe-plot-and-remove-the-outliers-from-the-dataset/. The function uses Tukey's method to ID ouliers ranged above and below the 1.5*IQR. 
outlierKD <- function(dt, var) {
     var_name <- eval(substitute(var),eval(dt))
     na1 <- sum(is.na(var_name))
     m1 <- mean(var_name, na.rm = T)
     par(mfrow=c(2, 2), oma=c(0,0,3,0))
     boxplot(var_name, main="With outliers")
     hist(var_name, main="With outliers", xlab=NA, ylab=NA)
     outlier <- boxplot.stats(var_name)$out
     mo <- mean(outlier)
     var_name <- ifelse(var_name %in% outlier, NA, var_name)
     boxplot(var_name, main="Without outliers")
     hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
     title("Outlier Check", outer=TRUE)
     na2 <- sum(is.na(var_name))
     cat("Outliers identified:", na2 - na1, "n")
     cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "n")
     cat("Mean of the outliers:", round(mo, 2), "n")
     m2 <- mean(var_name, na.rm = T)
     cat("Mean without removing outliers:", round(m1, 2), "n")
     cat("Mean if we remove outliers:", round(m2, 2), "n")
     response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
     if(response == "y" | response == "yes"){
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     } else{
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     }
}
outlierKD(EggsB1, EggDiamum) #I opted to remove outliers for now. 
#Outliers identified: 48 nPropotion (%) of outliers: 11.5 nMean of the outliers: 57.63 nMean without removing outliers: 62.4 nMean if we remove outliers: 62.95 n
yes #to remove outliers from EggsB1
#count how many eggs are left per female after removing the outliers. 
outrem<- aggregate(EggDiamum~FemaleID, data=EggsB1, FUN=length)
#EF07_B1 had 10 eggs that were included the rest were removed. Keep EF07_B1 in the analysis
```

Visualize eggs without outliers
```{r EggVisualizationNoOutliers}
#Plot again without outliers
par(mfcol=c(1,1))
boxplot(EggDiamum~FemaleID, data=EggsB1) 
boxplot(EggDiamum~ParentTrt, data=EggsB1)
#get the range now
range(EggsB1$EggDiamum, na.rm=TRUE)
#note that the eggs were filtered through a 70 um filter, but it appears that larger ones made it through.We are going to keep these in for now because they aren't that outrageously large. 
#plot histograms of the egg data with outliers removed
ggplot(EggsB1, aes(x=EggDiamum, color=ParentTrt))+
  geom_histogram(alpha=0.5, position="identity")
eggmeans<- aggregate(EggDiamum~ParentTrt, data=EggsB1, FUN=mean) #400 = 63.45709; 2800= 62.36829
eggmedians<- aggregate(EggDiamum~ParentTrt, data=EggsB1, FUN=median)#400 = 62.91200; 2800= 62.06727
view(eggmeans)
view(eggmedians)
#I will use the mean egg size. 
```

Make datasets without the egg size outliers
```{r MakeLarvaeDataframes}
meanegg<- ddply(EggsB1, .(FemaleID), numcolwise(mean, na.rm=T)) #get mean egg size
#select only columns that we need for egg morph
meanegg<- subset(meanegg, select=c("FemaleID", "EggDiamum"))
FemAdults<- merge(x=FemaleInfo, y= meanegg, by.x="ParentBlockID", by.y= "FemaleID", all.y=TRUE, all.x=TRUE)
#get the cilia area and extent data ready
CiliaExtent<- subset(CiliaArea, CilType=="Extent")
CiliaArea<- subset(CiliaArea, CilType=="Area")
#now use ddply to get only one value for each morphID
CiliaExtentAll<- ddply(CiliaExtent, .(LarvaeMorphID), numcolwise(mean, na.rm=TRUE))
CiliaAreaAll<- ddply(CiliaArea, .(LarvaeMorphID), numcolwise(mean, na.rm=TRUE))
colnames(CiliaExtentAll)<- paste(colnames(CiliaExtentAll), "ciliaextent", sep="")
colnames(CiliaAreaAll)<- paste(colnames(CiliaAreaAll), "ciliaarea", sep="")
#merge the cilia info to larvae morphology
Larvae_Morphology<- merge(Larvae_Morphology, CiliaExtentAll, by.x="LarvaeMorphID", by.y="LarvaeMorphIDciliaextent", all.x=TRUE)
Larvae_Morphology<- merge(Larvae_Morphology, CiliaAreaAll, by.x="LarvaeMorphID", by.y="LarvaeMorphIDciliaarea", all.x=TRUE)
#Next dataframe will be for one line per traced larva, includes all the data for that jar, including average egg size for that jar 
Larvae<-merge(x=Larvae_Morphology, y=Barcode_Jar, by="JarID", all.x=TRUE)
Larvae<- merge(x=Larvae, y=Larvae_WC, by="JarID", all.x=TRUE, all.y=TRUE)
#use gather function in tidyr to make the Fert_QG Jar ID columns in long form vs. wide form
Fert_QG_long <- Fert_QG %>% tidyr::gather(Key, JarID, JarID1:JarID6)
#use the gathered data frame to merge with Larvae
Larvae<-merge(x=Larvae, y=Fert_QG_long, by.x= c("JarID", "CrossID"), by.y=c("JarID", "CrossID"), all.x=TRUE) #did the join by two columns so CrossID column wasn't duplicated
Larvae<- merge(x=Larvae, y=Cross, by="CrossID", all.x=TRUE)
Larvae<- merge(x=Larvae, y=FemAdults, by.x= "FemaleID", by.y="ParentBlockID", all.x=TRUE)
Larvae<- merge(x=Larvae, y=MaleInfo, by.x= "MaleID", by.y="MaleParentBlockID", all.x=TRUE)
Larvae<- merge(x=Larvae, y= Larvae_Counts, by="JarID", all.x=TRUE)
#Remove jars without larvae
Larvae<- subset(Larvae, Larvae=="TRUE")
#get the growth per day for the larva
Larvae$GrowthPerDay<- (Larvae$LarvaeDiamum-Larvae$EggDiamum)/3
Larvae$Areaciliaareaum<- Larvae$Areaciliaarea/Larvae$LarvaePixPerum^2
Larvae$Perimeterciliaareaum<- Larvae$Perimeterciliaarea/Larvae$LarvaePixPerum
Larvae$Perimeterciliaextentum<- Larvae$Perimeterciliaextent/Larvae$LarvaePixPerum
Larvae$MajorAxisLengthciliaextentum<- Larvae$MajorAxisLengthciliaextent/Larvae$LarvaePixPerum
#Final dataframe will be one line per Jar, includes all data for the jar, including the average larva size per jar and cilia
LarvByJar<- merge(x=Larvae_Calcification, y=Larvae_Cilia, by="JarID", all.x=TRUE, all.y=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Barcode_Jar, by="JarID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Larvae_WC, by="JarID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Fert_QG_long, by.x=c("JarID", "CrossID"), by.y=c("JarID","CrossID"), all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Cross, by="CrossID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=FemAdults, by.x="FemaleID", by.y="ParentBlockID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=MaleInfo, by.x="MaleID", by.y="MaleParentBlockID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Larvae_Counts, by="JarID", all.x=TRUE)
#get the means for larvae morphology and add to LarvByJar
MeanLarvMorph<- ddply(Larvae_Morphology, .(JarID), numcolwise(mean, na.rm=T))
SELarvMorph<- ddply(Larvae_Morphology, .(JarID), numcolwise(se, na.rm=T)) #get standard error of larva size
MeanLarvMorph<- full_join(MeanLarvMorph,SELarvMorph, by="JarID", suffix=c("", "SE")) #join the mean and se larvae data
LarvByJar<- merge(x=LarvByJar, y=MeanLarvMorph, by="JarID", all.x=TRUE)
coVars<- ddply(Larvae, .(JarID), numcolwise(coefvariation)) #get the coefficients of variations at the 95% confidence levels for each jar. 
colnames(coVars)<- paste("coeffvar", colnames(coVars), sep="_")
LarvByJar<- merge(x=LarvByJar, y=coVars, by.x="JarID", by.y="coeffvar_JarID", all.x=TRUE)
#Now let's add the survival data. See notes on "Larvae Survival" for more info on how we decided to count total larvae in the jar
#Make columns for the two larvae counts for 1_5 and 2
LarvByJar$v1_5SurvCount<- rowSums(LarvByJar[,c("SWRTotal", "F2LarvaeCount")], na.rm=TRUE)
LarvByJar$v2SurvCount<- rowSums(LarvByJar[,c("F1LarvaeCount", "F2LarvaeCount")], na.rm=TRUE)
#use ifelse statement to define the LarvaeSurvived column
LarvByJar$LarvaeSurvived<- ifelse(LarvByJar$ProtocolVersion =="1", paste(LarvByJar$TotalLarvae), ifelse(LarvByJar$ProtocolVersion =="1_5", paste(LarvByJar$v1_5SurvCount), paste(LarvByJar$v2SurvCount)))
LarvByJar$LarvaeSurvived<- as.numeric(LarvByJar$LarvaeSurvived)
LarvByJar$GPD<- (LarvByJar$LarvaeDiamum-LarvByJar$EggDiamum)/3
#get the total larvae for the control jars and exposed jars for each family and then get the ratio of exposed to control to get an idea of how good a male and female pair are at larvae surviving. Can also use this to create reaction norms. 
SurvDat<- ddply(LarvByJar, .(CrossID, ParentTrt, JarTrt, MaleID, FemaleID, TrtTankID,MaleTrtTankID, AccTankID, Block), numcolwise(mean, na.rm=TRUE))
SESurvDat<- ddply(LarvByJar, .(CrossID, ParentTrt, JarTrt, Block), numcolwise(se, na.rm=TRUE)) #get the standard errors for each mean count
SESurvDat<- subset(SESurvDat, select= c(CrossID, JarTrt, LarvaeSurvived, SatArag_ATDIC, JarSatArg_SW)) #rename the columns
colnames(SESurvDat)<- paste("SE", colnames(SESurvDat), sep="")
#merge the standard error data to the SurvDat dataframe
SurvDat<- merge(SurvDat, SESurvDat, by.x= c("CrossID", "JarTrt"), by.y= c("SECrossID", "SEJarTrt"))
#only keep the relevant fields because otherwise this is a huge dataframe
SurvDat<- subset(SurvDat, select= c(CrossID, JarTrt, ParentTrt, MaleID, FemaleID, TrtTankID,MaleTrtTankID, AccTankID, LarvaeSurvived, Block, SELarvaeSurvived, SatArag_ATDIC, GPD )) 
SurvDat<- subset(SurvDat, Block=="B1")
SurvDat<- subset(SurvDat, CrossID!="")
#use the spread function to go from long form to wide form for the survival data
SurvDatToWide<- unite(SurvDat, Value, c(LarvaeSurvived, SELarvaeSurvived, GPD), sep="_")
SurvWideDat<- spread(SurvDatToWide, JarTrt, value= Value)
#now split the mean and SE apart again
SurvWideDat<- separate(SurvWideDat, Control, into= c("MeanLarvSurvivedCon", "SELarvSurvivedCon", "MeanGPDCon"), sep="_")
SurvWideDat<- separate(SurvWideDat, Exposed, into= c("MeanLarvSurvivedExp", "SELarvSurvivedExp", "MeanGPDExp"), sep="_")
#make the means and ses numeric
SurvWideDat$MeanLarvSurvivedExp<- as.numeric(SurvWideDat$MeanLarvSurvivedExp)
SurvWideDat$MeanLarvSurvivedCon<- as.numeric(SurvWideDat$MeanLarvSurvivedCon)
SurvWideDat$SELarvSurvivedExp<- as.numeric(SurvWideDat$SELarvSurvivedExp)
SurvWideDat$SELarvSurvivedCon<- as.numeric(SurvWideDat$SELarvSurvivedCon)
SurvWideDat$MeanGPDCon<- as.numeric(SurvWideDat$MeanGPDCon)
SurvWideDat$MeanGPDExp<- as.numeric(SurvWideDat$MeanGPDExp)
SurvWideDat$RatSurv<- SurvWideDat$MeanLarvSurvivedExp/SurvWideDat$MeanLarvSurvivedCon
SurvWideDat$RatGPD<- SurvWideDat$MeanGPDExp/SurvWideDat$MeanGPDCon
SurvWideDat$SurvChange<- (SurvWideDat$MeanLarvSurvivedCon-SurvWideDat$MeanLarvSurvivedExp)/SurvWideDat$MeanLarvSurvivedCon
SurvWideDat$CrossID<- as.factor(SurvWideDat$CrossID)
SurvWideDat$FemaleID<- as.factor(SurvWideDat$FemaleID)
SurvWideDat$MaleID<- as.factor(SurvWideDat$MaleID)
LarvByJar<- subset(LarvByJar, Larvae=="TRUE")

#now remove the data we do not want to analyze.
LarvaeDat<- subset(Larvae, Block == "B1")
#get a column without B1 in the female and male IDs 
LarvaeDat$FemIDPlain<- vapply(strsplit(as.character(LarvaeDat$FemaleID),"_"), `[`, 1, FUN.VALUE=character(1))
LarvaeDat$MaleIDPlain<- vapply(strsplit(as.character(LarvaeDat$MaleID),"_"), `[`, 1, FUN.VALUE=character(1))
#now make things factors and drop levels
LarvaeDat$JarID<- as.factor(LarvaeDat$JarID)
LarvaeDat$JarID<-droplevels(LarvaeDat)$JarID
LarvaeDat$AccTankID<- as.factor(LarvaeDat$AccTankID)
LarvaeDat$AccTankID<-droplevels(LarvaeDat)$AccTankID
LarvaeDat$TrtTankID<- as.factor(LarvaeDat$TrtTankID)
LarvaeDat$TrtTankID<-droplevels(LarvaeDat)$TrtTankID
LarvaeDat$MaleTrtTankID<- as.factor(LarvaeDat$MaleTrtTankID)
LarvaeDat$MaleTrtTankID<-droplevels(LarvaeDat)$MaleTrtTankID
LarvaeDat$MaleAccTankID<- as.factor(LarvaeDat$MaleAccTankID)
LarvaeDat$MaleAccTankID<-droplevels(LarvaeDat)$MaleAccTankID
LarvaeDat$JarSeatable<- as.factor(LarvaeDat$JarSeatable)
LarvaeDat$JarSeatable<-droplevels(LarvaeDat)$JarSeatable
LarvaeDat$JarTrt<- as.factor(LarvaeDat$JarTrt)
LarvaeDat$ParentTrt<- as.factor(LarvaeDat$ParentTrt)
LarvaeDat$FemaleID<- as.factor(LarvaeDat$FemaleID)
LarvaeDat$FemaleID<-droplevels(LarvaeDat)$FemaleID
LarvaeDat$MaleID<- as.factor(LarvaeDat$MaleID)
LarvaeDat$MaleID<-droplevels(LarvaeDat)$MaleID
LarvaeDat$CrossID<- as.factor(LarvaeDat$CrossID)
LarvaeDat$CrossID<-droplevels(LarvaeDat)$CrossID
LarvByJarDat<- subset(LarvByJar, Block == "B1")

LarvByJarDat$JarID<- as.factor(LarvByJarDat$JarID)
LarvByJarDat$AccTankID<- as.factor(LarvByJarDat$AccTankID)
LarvByJarDat$TrtTankID<- as.factor(LarvByJarDat$TrtTankID)
LarvByJarDat$MaleAccTankID<- as.factor(LarvByJarDat$MaleAccTankID)
LarvByJarDat$MaleTrtTankID<- as.factor(LarvByJarDat$MaleTrtTankID)
LarvByJarDat$JarSeatable<- as.factor(LarvByJarDat$JarSeatable)
LarvByJarDat$ParentTrt<- as.factor(LarvByJarDat$ParentTrt)
LarvByJarDat$FemaleID<- as.factor(LarvByJarDat$FemaleID)
LarvByJarDat$MaleID<- as.factor(LarvByJarDat$MaleID)
LarvByJarDat$CrossID<- as.factor(LarvByJarDat$CrossID)
LarvByJarDat$JarTrt<- as.factor(LarvByJarDat$JarTrt)
LarvByJarDat$MaleID<-droplevels(LarvByJarDat)$MaleID
LarvByJarDat$FemaleID<-droplevels(LarvByJarDat)$FemaleID
LarvByJarDat$CrossID<-droplevels(LarvByJarDat)$CrossID
LarvByJarDat$FemIDPlain<- vapply(strsplit(as.character(LarvByJarDat$FemaleID),"_"), `[`, 1, FUN.VALUE=character(1))
LarvByJarDat$MaleIDPlain<- vapply(strsplit(as.character(LarvByJarDat$MaleID),"_"), `[`, 1, FUN.VALUE=character(1))
#make a column for the ratio of major to minor axis
LarvaeDat$MajMinRat<- LarvaeDat$LarvaeMajorAxisLengthPix/LarvaeDat$LarvaeMinorAxisLengthPix
LarvaeDat$PerimLenRat<-LarvaeDat$LarvaePerimeterum/LarvaeDat$LarvaeDiamum
LarvByJarDat<- subset(LarvByJarDat, JarSatArg_SW !="NA")
```

Analyze jar chemistry data
```{r JarChemistryStat}
CalJar<- subset(LarvByJarDat, BottleChemistry=="TRUE" )#only use those that had their chem measured
#look at the saturation state for jars
boxplot(JarSatArg_SW~JarTrt*JarSeatable, data=CalJar)
#I was having trouble with making this model: JarSatArg_SW~JarTrt + (1|ParentTrt)+ (1|TimeFilter), so I looked at TimeFilter because I think that is the problem. 
Jar1<- lm(JarSatArg_SW~TimeFilter, data=CalJar)
par(mfcol=c(2,2))
plot(Jar1)
par(mfcol=c(1,1))
#looks like we can't use timefilter because it is too specific to each jar. Get the error message "observations with leverage one"; try using seatable instead
Jar2<- lmer(JarSatArg_SW~JarTrt + (1|JarSeatable), data=CalJar)
Jar2step<- step(Jar2)
#get error model not of class lmerMod. I looked up this issue and turns out that error results when all random effects are dropped. See https://stackoverflow.com/questions/55638476/how-to-do-stepwise-model-with-random-effect-lme4-lmertest for more info if you want. 
JarFin<- lm(JarSatArg_SW~JarTrt, data=CalJar)
#check assumptions 
par(mfcol=c(2,2))
plot(JarFin)
par(mfcol=c(1,1))
acf(resid(JarFin))
summary(JarFin)
#Conclusion: JarSeatable doesn't need to be accounted for as a random effect in terms of water chem. 
```

# Get Jar Chemistry to add to Carb summary table
```{r Jar Carb Sum}
#add the jar chemistry
CarbChemJar<- subset(LarvByJarDat, select= c("Trt", "BottleChemistry", "JarpHNBS", "JarpHSW", "JarSalinity", "JarTemp", "JarDICCorr", "JarATCorr", "JarATCalc_SW", "JarpCO2_SW", "JarHCO3_SW", "JarCO3_SW", "JarCO2_SW", "JarSatCalcite_SW", "JarSatArg_SW" ))
JarChemMean<- ddply(CarbChemJar, .(Trt), numcolwise(mean, na.rm=T)) 
JarChemMean[,c(2,3)]<- round(JarChemMean[,c(2,3)], 2) #round pHs to two decimal places. 
JarChemMean[,c(13,14)]<- round(JarChemMean[,c(13,14)],3) #round saturation states to three decimal places.
JarChemMean[,c(6:12)]<- round(JarChemMean[,c(6:12)]) #round AT, DIC, pCO2, CO3, Co2, and HCO3 to no decimal places. 
JarChemMean[,c(4,5)]<- round(JarChemMean[,c(4,5)], 1) #round temp and salinity to one decimal places. 

JarChemSD<- ddply(CarbChemJar, .(Trt), numcolwise(sd, na.rm=T)) 
JarChemSD[,c(2,3)]<- round(JarChemSD[,c(2,3)], 2) #round pHs to two decimal places. 
JarChemSD[,c(13,14)]<- round(JarChemSD[,c(13,14)],3) #round saturation states to three decimal places.
JarChemSD[,c(6:12)]<- round(JarChemSD[,c(6:12)]) #round AT, DIC, pCO2, CO3, Co2, and HCO3 to no decimal places. 
JarChemSD[,c(4,5)]<- round(JarChemSD[,c(4,5)], 1) #round temp and salinity to one decimal places. 
JarChemMin<- ddply(CarbChemJar, .(Trt), numcolwise(min, na.rm=T)) 
JarChemMin[,c(2,3)]<- round(JarChemMin[,c(2,3)], 2) #round pHs to two decimal places. 
JarChemMin[,c(13,14)]<- round(JarChemMin[,c(13,14)],3) #round saturation states to three decimal places.
JarChemMin[,c(6:12)]<- round(JarChemMin[,c(6:12)]) #round AT, DIC, pCO2, CO3, Co2, and HCO3 to no decimal places. 
JarChemMin[,c(4,5)]<- round(JarChemMin[,c(4,5)], 1) #round temp and salinity to one decimal places. 
JarChemMax<- ddply(CarbChemJar, .(Trt), numcolwise(max, na.rm=T)) 
JarChemMax[,c(2,3)]<- round(JarChemMax[,c(2,3)], 2) #round pHs to two decimal places. 
JarChemMax[,c(13,14)]<- round(JarChemMax[,c(13,14)],3) #round saturation states to three decimal places.
JarChemMax[,c(6:12)]<- round(JarChemMax[,c(6:12)]) #round AT, DIC, pCO2, CO3, Co2, and HCO3 to no decimal places. 
JarChemMax[,c(4,5)]<- round(JarChemMax[,c(4,5)], 1) #round temp and salinity to one decimal places. 
#rename columns
colnames(JarChemSD)<- paste(colnames(JarChemSD),"sd", sep="_")
colnames(JarChemMin)<- paste(colnames(JarChemMin),"min", sep="_")
colnames(JarChemMax)<- paste(colnames(JarChemMax),"max", sep="_")

#merge all of the jar data together. 
JarCarbChem<- merge(JarChemMean, JarChemSD, by.x="Trt", by.y="Trt_sd") %>% 
  merge(JarChemMin, by.x="Trt", by.y="Trt_min") %>% 
  merge(JarChemMax, by.x="Trt", by.y= "Trt_max") %>% 
  t() %>% 
  as.data.frame()
colnames(JarCarbChem)<- paste(c("LarvaeControl", "LarvaeExposed"))
#write the csv to get the parameters for the carb table for Jars
write.csv(JarCarbChem,"~/R/GitHub/LarvaeTransGen2018/data/JarCarbChemSummary.csv")
```


Test for significant effect of parent treatment on egg size. Start with measured water chem
```{r EggSizeStat}
CalEggs<- subset(EggsB1, SatArag_ATDIC != "NA")
CalEggs$TrtTankID<-droplevels(CalEggs)$TrtTankID
#start with models that have the response variable EggDiamum
eggdiam1<- lmer(EggDiamum~SatArag_ATDIC*AdultLength + (1|FemaleID)  + (1|AccTankID), data=CalEggs)
eggstep<-step(eggdiam1)
print(eggstep)
#final model chosen: EggDiamum~ 1 + (1|FemaleID)
eggdiamfin<- lmer(EggDiamum~1+(1|FemaleID), data=CalEggs)
#Check assumptions
par(mfcol=c(1,1))
qqnorm(resid(eggdiamfin))
qqline(resid(eggdiamfin))
plot(eggdiamfin)
shapiro.test(resid(eggdiamfin)) #does not meet assumption of normality for shapiro test. 
acf(resid(eggdiamfin))
#the above looks fine, seems to meet all assumptions. 
summary(eggdiamfin)
#Conclusion: Female matters but not treatment or adult length. Acclimation tank and treatment tank also don't matter.  
plot(EggDiamum~FemaleID, data=CalEggs) #we need to account for FemaleID, but not egg size in Larvae models. Can also make parallel models one with EggDiamum as a covariate and one with Female ID as a random effect

#also test eccentricity
eggperimdiam1<- lmer(EggPerimDiamRat~SatArag_ATDIC*AdultLength + (1|FemaleID)  + (1|AccTankID), data=CalEggs)
eggperimdiamstep<-step(eggperimdiam1)
print(eggperimdiamstep)
eggperimdiamfin<- lmer(EggPerimDiamRat^3~(1|FemaleID), data=CalEggs)
par(mfcol=c(1,1))
qqnorm(resid(eggperimdiamfin))
qqline(resid(eggperimdiamfin))
plot(eggperimdiamfin)
acf(resid(eggperimdiamfin))
shapiro.test(resid(eggperimdiamfin))
summary(eggperimdiamfin)
plot(EggPerimDiamRat~FemaleID, data=CalEggs)
```
Egg data don't meet assumption of normality according to Shapiro-Wilk, what is the best next step? 

```{r EggSizeStatFixedEffect, include=FALSE}
eggdiam1f<- lmer(EggDiamum~ParentTrt*AdultLength + (1|FemaleID) + (1|TrtTankID)+ (1|AccTankID), data=CalEggs)
eggstepf<-step(eggdiam1f)
print(eggstepf)
#final model chosen: EggDiamum~ 1 + (1|FemaleID), which is the same as the eggdiamfin model
summary(eggdiamfin)
shapiro.test(resid(eggdiamfin))
```

Choosing mean or median for larvae length data
```{r LarvaeDiamMeanOrMedian}
ggplot(LarvaeDat, aes(x=LarvaeDiamum, color=JarTrt))+
  geom_histogram(alpha=0.5, position="identity") +
  facet_grid(~ParentTrt)
larvmeans<- aggregate(LarvaeDiamum~ParentTrt*JarTrt, data=LarvaeDat, FUN=mean)
larvmedians<- aggregate(LarvaeDiamum~ParentTrt*JarTrt, data=LarvaeDat, FUN=median)
view(larvmeans)
view(larvmedians)
#I think we can use the mean for these data 

#look for the number of jars in each treatment
larvjarn<- aggregate(LarvaeDiamum~ParentTrt*JarTrt, data=LarvByJarDat, FUN=length)
```

Look at the larvae diameter data
Use fixed effect for adult treatment and larvae treatment as covariate
```{r LarvDiamStat}
larvlen<-lmer(LarvaeDiamum~ParentTrt*JarSatArg_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarID)+(1|JarSeatable)+(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
steppedlarvlen<-step(larvlen)
print(steppedlarvlen)
#final model chosen by the lme4 step function: y~ParentTrt*JarSatArg_SW +(1|FemaleID)+(1|MaleID)+(1|JarID)+ (1|JarSeatable)
larvlenfin<- lmer(LarvaeDiamum~ParentTrt*JarSatArg_SW+ (1|FemaleID)+(1|MaleID)+(1|JarID)+ (1|JarSeatable), data=LarvaeDat,control = lmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))

#check the final model to see that it meets assumptions
plot(larvlenfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvlenfin))
qqline(resid(larvlenfin)) #has relatively long tails, I think this should be acceptable
shapiro.test(resid(larvlenfin))
acf(resid(larvlenfin)) #this looks good 
summary(larvlenfin)
```
### Is it okay to just use the visual assessment to look at assumptions? 


Analyze larvae growth with adult trt as fixed effect and jar as covariate
```{r LarvGrowthStat}
larvgrow<-lmer(GrowthPerDay~ParentTrt*JarSatArg_SW+(1|MaleID)+(1|CrossID)+(1|JarID) +(1|JarSeatable)+(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
steppedlarvgrow<- step(larvgrow)
print(steppedlarvgrow)
#final model chosen by the lme4 step function: y~SatCalcite_ATDIC*JarTrt+ (1|CrossID)+ (1|JarID) + (1|JarSeatable)
larvgrowfin<- lmer(GrowthPerDay~ParentTrt*JarSatArg_SW+(1|CrossID)+ (1|JarID) + (1|JarSeatable), data=LarvaeDat)
#check the final model to see that it meets assumptions
plot(larvgrowfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvgrowfin))
qqline(resid(larvgrowfin)) #has relatively long tails, based on scale I think it reasonably meets assumption of normality. 
acf(resid(larvgrowfin))
shapiro.test(resid(larvgrowfin)) #doesn't meet assumption of normality, but it looks fine to me.
summary(larvgrowfin)

ggplot(LarvaeDat, aes(x=GrowthPerDay,color=JarTrt))+
  geom_histogram(alpha=0.5, position="identity")+
  facet_grid(ParentTrt~.)
ProblemCross<- subset(LarvaeDat, CrossID=="CF02_CM02_B1_QG")# I don't know why I labeled this as a problem cross. It may have previously had mislabeled control vs OA jars
ggplot(ProblemCross, aes(y=GrowthPerDay,x=JarID, color=JarTrt))+
  geom_boxplot(alpha=0.5, position="identity")
```

Look at larvae area; Parent Trt as fixed effect and Jar Treatment as covariate
```{r SurfaceAreaStat}
larvareat<-lmer(sqrt(LarvaeAreaum2)~ParentTrt*JarSatArg_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|TrtTankID)+(1|MaleTrtTankID) , data=LarvaeDat)
steppedlarvareat<-step(larvareat)
print(steppedlarvareat)
#final model chosen by the lme4 step function: y~SatArag_ATDIC*JarTrt+(1|FemaleID)+(1|MaleID)+(1|JarID)+(1|JarSeatable)
larvareafint<- lmer(sqrt(LarvaeAreaum2)~ParentTrt*JarSatArg_SW+(1|FemaleID)+(1|MaleID)+(1|JarID)+(1|JarSeatable), data=LarvaeDat, control = lmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))
plot(larvareafint) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvareafint))
qqline(resid(larvareafint)) #this seems to have improved it
shapiro.test(resid(larvareafint)) #doesn't meet assumption according to shapiro. 
acf(resid(larvareafint)) #this looks fine to me. 
summary(larvareafint)
```

Characterize the larvae by the ratio of major and minor axis
```{r LarvAxesStat}
larvrat<-lmer(MajMinRat~ParentTrt*JarSatArg_SW+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) +(1|CrossID)++(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
larvratstep<- step(larvrat)
print(larvratstep)
larvratfin<- lmer(MajMinRat~ParentTrt*JarSatArg_SW +(1|FemaleID)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larvratfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvratfin))
qqline(resid(larvratfin)) #has relatively long tails, doesn't seem to meet assumption of normality. try a transformation
acf(resid(larvratfin)) #this looks fine.
summary(larvratfin)
```

Look at the ratio of perimeter to diameter with Parent as a covariate and JarTrt as a fixed effect. For a circle, this would be equal to pi
```{r PerimeterToLengthStat}
PerDiRat<- lmer((LarvaePerimeterum/LarvaeDiamum)~ParentTrt*JarSatArg_SW+(1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID) +(1|CrossID)+(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
PerDiStep<- step(PerDiRat)
print(PerDiStep)
#final model chosen PerimDiamRat ~ SatArag_ATDIC * JarTrt + (1 | FemaleID) + (1 | JarID)
PerDiRatFin<- lmer((LarvaePerimeterum/LarvaeDiamum) ~ ParentTrt*JarSatArg_SW + (1 | FemaleID) + (1 | JarID)+(1|JarSeatable), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(PerDiRatFin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(PerDiRatFin))
qqline(resid(PerDiRatFin)) #this looks fine
acf(resid(PerDiRatFin))
summary(PerDiRatFin)
```

Analyze eccentricity with parent as a covariate and jar as fixed effect
```{r EccentricityStat}
larveccen<-lmer(LarvaeEccentricity^3~ParentTrt*JarSatArg_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
steppedlarveccen<- step(larveccen)
print(steppedlarveccen)
#final model
larveccenfin<- lmer(LarvaeEccentricity^3~ParentTrt*JarSatArg_SW+ (1|FemaleID)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larveccenfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larveccenfin))
qqline(resid(larveccenfin)) #has relatively long tails, doesn't meet assumption of normality. 
acf(resid(larveccenfin))
summary(larveccenfin)
```

Test Larvae perimeter
```{r PerimeterStat}
larvperim<-lmer(LarvaePerimeterum~ParentTrt*JarSatArg_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
steppedlarvperim<- step(larvperim)
print(steppedlarvperim)
#final model chosen by the step function: LarvaePerimeterum~SatArag_ATDIC*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)
larvperimfin<- lmer(LarvaePerimeterum~ParentTrt*JarSatArg_SW+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larvperimfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvperimfin))
qqline(resid(larvperimfin)) #has relatively long tails, doesn't meet assumption of normality. 
acf(resid(larvperimfin)) #this looks fine to me.
summary(larvperimfin)
```

Analyze survival data with parent as a fixed effect
```{r SurvivalStat}
surv1<- lmer(RatSurv~ParentTrt+ (1|FemaleID)+(1|MaleID)+(1|TrtTankID)+(1|MaleTrtTankID), data=SurvWideDat)
plot(surv1)
qqnorm(resid(surv1))
qqline(resid(surv1))
acf(resid(surv1))
survstepFE<- step(surv1)
print(survstepFE)
survfin<- lmer(RatSurv~ (1|FemaleID), data=SurvWideDat)#final model chosen by step function: RatSurv~ (1|FemaleID)
plot(survfin)
qqnorm(resid(survfin))
qqline(resid(survfin))
acf(resid(survfin))
summary(survfin)

#try to see if this is significantly different for exposed vs control
SurvOA<- subset(SurvWideDat, ParentTrt=="2600")
SurvCon<- subset(SurvWideDat, ParentTrt=="400")
t.test(SurvOA$RatSurv, mu = 1, alternative = "two.sided")
t.test(SurvCon$RatSurv, mu = 1, alternative = "two.sided")

```

Examine coefficient of variation for each jar
```{r CoeffVar Stats}
GPDcoeffvar<-lmer(coeffvar_GrowthPerDay~ParentTrt*JarSatArg_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|TrtTankID)+(1|MaleTrtTankID), data=LarvByJarDat)
steppedGPDCoeffvar<-step(GPDcoeffvar) 
print(GPDcoeffvar)
#looks like it doesn't choose a model with any random effects. 

GPDCoeffVarlm<- lm(log(coeffvar_GrowthPerDay)~ParentTrt*JarSatArg_SW, data=LarvByJarDat, na.action=na.exclude)
GPDCoeffVarlm1<- lm(log(coeffvar_GrowthPerDay)~ParentTrt+JarSatArg_SW, data=LarvByJarDat, na.action=na.exclude)
anova(GPDCoeffVarlm1, GPDCoeffVarlm)
par(mfcol=c(2,2))
plot(GPDCoeffVarlm1)
par(mfcol=c(1,1))
shapiro.test(resid(GPDCoeffVarlm1))
summary(GPDCoeffVarlm1)
acf(resid(GPDCoeffVarlm1)) #this looks good 
#does not look like it meets assumptions.
```

Final Models to Use
```{r FinalModelsWithSameRandomEffects}
summary(larvgrowfin)
summary(larvareafint)
summary(larvlenfin)
summary(larvperimfin)
summary(PerDiRatFin)
```

```{r LegendEnhanced functioin, include=FALSE}
legend.enhanced <- function (x, y = NULL, legend, fill = NULL, col = par("col"), 
          border = "black", lty, lwd, pch, angle = 45, density = NULL, 
          bty = "o", bg = par("bg"), box.lwd = par("lwd"), box.lty = par("lty"), 
          box.col = par("fg"), pt.bg = NA, cex = 1, pt.cex = cex, 
          pt.lwd = lwd, xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1, 
          adj = c(0, 0.5), text.width = NULL, text.col = par("col"), 
          text.font = NULL, merge = do.lines && has.pch, trace = FALSE, 
          plot = TRUE, ncol = 1, horiz = FALSE, title = NULL, inset = 0, 
          xpd, title.col = text.col, title.adj = c(0.5,0), seg.len = 2) 
{
  if (missing(legend) && !missing(y) && (is.character(y) || 
                                         is.expression(y))) {
    legend <- y
    y <- NULL
  }
  mfill <- !missing(fill) || !missing(density)
  if (!missing(xpd)) {
    op <- par("xpd")
    on.exit(par(xpd = op))
    par(xpd = xpd)
  }
  title <- as.graphicsAnnot(title)
  if (length(title) > 1) 
    stop("invalid 'title'")
  legend <- as.graphicsAnnot(legend)
  n.leg <- if (is.call(legend)) 
    1
  else length(legend)
  if (n.leg == 0) 
    stop("'legend' is of length 0")
  auto <- if (is.character(x)) 
    match.arg(x, c("bottomright", "bottom", "bottomleft", 
                   "left", "topleft", "top", "topright", "right", "center"))
  else NA
  if (is.na(auto)) {
    xy <- xy.coords(x, y)
    x <- xy$x
    y <- xy$y
    nx <- length(x)
    if (nx < 1 || nx > 2) 
      stop("invalid coordinate lengths")
  }
  else nx <- 0
  xlog <- par("xlog")
  ylog <- par("ylog")
  rect2 <- function(left, top, dx, dy, density = NULL, angle, 
                    ...) {
    r <- left + dx
    if (xlog) {
      left <- 10^left
      r <- 10^r
    }
    b <- top - dy
    if (ylog) {
      top <- 10^top
      b <- 10^b
    }
    rect(left, top, r, b, angle = angle, density = density, 
         ...)
  }
  segments2 <- function(x1, y1, dx, dy, ...) {
    x2 <- x1 + dx
    if (xlog) {
      x1 <- 10^x1
      x2 <- 10^x2
    }
    y2 <- y1 + dy
    if (ylog) {
      y1 <- 10^y1
      y2 <- 10^y2
    }
    segments(x1, y1, x2, y2, ...)
  }
  points2 <- function(x, y, ...) {
    if (xlog) 
      x <- 10^x
    if (ylog) 
      y <- 10^y
    points(x, y, ...)
  }
  text2 <- function(x, y, ...) {
    if (xlog) 
      x <- 10^x
    if (ylog) 
      y <- 10^y
    text(x, y, ...)
  }
  if (trace) 
    catn <- function(...) do.call("cat", c(lapply(list(...), 
                                                  formatC), list("\n")))
  cin <- par("cin")
  Cex <- cex * par("cex")
  if (is.null(text.width)) 
    text.width <- max(abs(strwidth(legend, units = "user", 
                                   cex = cex, font = text.font)))
  else if (!is.numeric(text.width) || text.width < 0) 
    stop("'text.width' must be numeric, >= 0")
  xc <- Cex * xinch(cin[1L], warn.log = FALSE)
  yc <- Cex * yinch(cin[2L], warn.log = FALSE)
  if (xc < 0) 
    text.width <- -text.width
  xchar <- xc
  xextra <- 0
  yextra <- yc * (y.intersp - 1)
  ymax <- yc * max(1, strheight(legend, units = "user", cex = cex)/yc)
  ychar <- yextra + ymax
  if (trace) 
    catn("  xchar=", xchar, "; (yextra,ychar)=", c(yextra, 
                                                   ychar))
  if (mfill) {
    xbox <- xc * 0.8
    ybox <- yc * 0.5
    dx.fill <- xbox
  }
  do.lines <- (!missing(lty) && (is.character(lty) || any(lty > 
                                                            0))) || !missing(lwd)
  n.legpercol <- if (horiz) {
    if (ncol != 1) 
      warning(gettextf("horizontal specification overrides: Number of columns := %d", 
                       n.leg), domain = NA)
    ncol <- n.leg
    1
  }
  else ceiling(n.leg/ncol)
  has.pch <- !missing(pch) && length(pch) > 0
  if (do.lines) {
    x.off <- if (merge) 
      -0.7
    else 0
  }
  else if (merge) 
    warning("'merge = TRUE' has no effect when no line segments are drawn")
  if (has.pch) {
    if (is.character(pch) && !is.na(pch[1L]) && nchar(pch[1L], 
                                                      type = "c") > 1) {
      if (length(pch) > 1) 
        warning("not using pch[2..] since pch[1L] has multiple chars")
      np <- nchar(pch[1L], type = "c")
      pch <- substr(rep.int(pch[1L], np), 1L:np, 1L:np)
    }
    if (!is.character(pch)) 
      pch <- as.integer(pch)
  }
  if (is.na(auto)) {
    if (xlog) 
      x <- log10(x)
    if (ylog) 
      y <- log10(y)
  }
  if (nx == 2) {
    x <- sort(x)
    y <- sort(y)
    left <- x[1L]
    top <- y[2L]
    w <- diff(x)
    h <- diff(y)
    w0 <- w/ncol
    x <- mean(x)
    y <- mean(y)
    if (missing(xjust)) 
      xjust <- 0.5
    if (missing(yjust)) 
      yjust <- 0.5
  }
  else {
    h <- (n.legpercol + (!is.null(title))) * ychar + yc
    w0 <- text.width + (x.intersp + 1) * xchar
    if (mfill) 
      w0 <- w0 + dx.fill
    if (do.lines) 
      w0 <- w0 + (seg.len + x.off) * xchar
    w <- ncol * w0 + 0.5 * xchar
    if (!is.null(title) && (abs(tw <- strwidth(title, units = "user", 
                                               cex = cex) + 0.5 * xchar)) > abs(w)) {
      xextra <- (tw - w)/2
      w <- tw
    }
    if (is.na(auto)) {
      left <- x - xjust * w
      top <- y + (1 - yjust) * h
    }
    else {
      usr <- par("usr")
      inset <- rep_len(inset, 2)
      insetx <- inset[1L] * (usr[2L] - usr[1L])
      left <- switch(auto, bottomright = , topright = , 
                     right = usr[2L] - w - insetx, bottomleft = , 
                     left = , topleft = usr[1L] + insetx, bottom = , 
                     top = , center = (usr[1L] + usr[2L] - w)/2)
      insety <- inset[2L] * (usr[4L] - usr[3L])
      top <- switch(auto, bottomright = , bottom = , bottomleft = usr[3L] + 
                      h + insety, topleft = , top = , topright = usr[4L] - 
                      insety, left = , right = , center = (usr[3L] + 
                                                             usr[4L] + h)/2)
    }
  }
  if (plot && bty != "n") {
    if (trace) 
      catn("  rect2(", left, ",", top, ", w=", w, ", h=", 
           h, ", ...)", sep = "")
    rect2(left, top, dx = w, dy = h, col = bg, density = NULL, 
          lwd = box.lwd, lty = box.lty, border = box.col)
  }
  xt <- left + xchar + xextra + (w0 * rep.int(0:(ncol - 1), 
                                              rep.int(n.legpercol, ncol)))[1L:n.leg]
  yt <- top - 0.5 * yextra - ymax - (rep.int(1L:n.legpercol, 
                                             ncol)[1L:n.leg] - 1 + (!is.null(title))) * ychar
  if (mfill) {
    if (plot) {
      if (!is.null(fill)) 
        fill <- rep_len(fill, n.leg)
      rect2(left = xt, top = yt + ybox/2, dx = xbox, dy = ybox, 
            col = fill, density = density, angle = angle, 
            border = border)
    }
    xt <- xt + dx.fill
  }
  if (plot && (has.pch || do.lines)) 
    col <- rep_len(col, n.leg)
  if (missing(lwd) || is.null(lwd)) 
    lwd <- par("lwd")
  if (do.lines) {
    if (missing(lty) || is.null(lty)) 
      lty <- 1
    lty <- rep_len(lty, n.leg)
    lwd <- rep_len(lwd, n.leg)
    ok.l <- !is.na(lty) & (is.character(lty) | lty > 0) & 
      !is.na(lwd)
    if (trace) 
      catn("  segments2(", xt[ok.l] + x.off * xchar, ",", 
           yt[ok.l], ", dx=", seg.len * xchar, ", dy=0, ...)")
    if (plot) 
      segments2(xt[ok.l] + x.off * xchar, yt[ok.l], dx = seg.len * 
                  xchar, dy = 0, lty = lty[ok.l], lwd = lwd[ok.l], 
                col = col[ok.l])
    xt <- xt + (seg.len + x.off) * xchar
  }
  if (has.pch) {
    pch <- rep_len(pch, n.leg)
    pt.bg <- rep_len(pt.bg, n.leg)
    pt.cex <- rep_len(pt.cex, n.leg)
    pt.lwd <- rep_len(pt.lwd, n.leg)
    ok <- !is.na(pch)
    if (!is.character(pch)) {
      ok <- ok & (pch >= 0 | pch <= -32)
    }
    else {
      ok <- ok & nzchar(pch)
    }
    x1 <- (if (merge && do.lines) 
      xt - (seg.len/2) * xchar
      else xt)[ok]
    y1 <- yt[ok]
    if (trace) 
      catn("  points2(", x1, ",", y1, ", pch=", pch[ok], 
           ", ...)")
    if (plot) 
      points2(x1, y1, pch = pch[ok], col = col[ok], cex = pt.cex[ok], 
              bg = pt.bg[ok], lwd = pt.lwd[ok])
  }
  xt <- xt + x.intersp * xchar
  if (plot) {
    if (is.na(title.adj[2])) { title.adj[2] <- 0}
    if (!is.null(title)) 
      text2(left + w * title.adj[1], top - ymax, labels = title, 
            adj = c(title.adj[1], title.adj[2]), cex = cex, col = title.col)
    text2(xt, yt, labels = legend, adj = adj, cex = cex, 
          col = text.col, font = text.font)
  }
  invisible(list(rect = list(w = w, h = h, left = left, top = top), 
                 text = list(x = xt, y = yt)))
}
```

Create multipanel figure
```{r MultipanelFigure}
par(mfcol=c(1,1), mar=c(4,4.3, .75,1))
p1=visreg(larvgrowfin, "JarSatArg_SW", by="ParentTrt", overlay=TRUE, ylab=expression(paste("Growth (", mu, "m/day)")), xlab=expression(paste("Larvae Treatment ", Omega[Aragonite])), legend=FALSE)
par(mfcol=c(2,2))
p2<-visreg(larvlenfin, "JarSatArg_SW", by="ParentTrt", overlay=TRUE, ylab=expression(paste("Length (", mu, "m)")), xlab="", legend=FALSE)
p3<-visreg(larvareafint, "JarSatArg_SW", by="ParentTrt", overlay=TRUE, ylab=expression(sqrt(paste("Area (", mu, m^2,")"))), xlab=expression(paste("Larvae Treatment ", Omega[Aragonite])), legend=FALSE)
p4<-visreg(larvperimfin, "JarSatArg_SW", by="ParentTrt", overlay=TRUE, ylab=expression(paste("Perimeter (", mu, "m)")), xlab="", legend=FALSE)
p5<-visreg(PerDiRatFin, "JarSatArg_SW", by="ParentTrt", overlay=TRUE, ylab="Perimeter :Length", xlab=expression(paste("Larvae Treatment ", Omega[Aragonite])), legend=FALSE)

```

# Get ratio of OA to control to understand differences in calcification
```{r OA:Control Calcification}
#get means of the data for each family for each treatment
FamMeans<- ddply(LarvByJarDat, .(CrossID, JarTrt), numcolwise(mean, na.rm=T))
#use spread function in tidyr to make the families in wide form
FamMeansWide<- pivot_wider(FamMeans, id_cols= CrossID, names_from= JarTrt, values_from= c(GrowthPerDay, LarvaeAreaum2, LarvaeDiamum, LarvaePerimeterum))

FamMeansWide$OACont_GPD<- (FamMeansWide$GrowthPerDay_Exposed-FamMeansWide$GrowthPerDay_Control)/FamMeansWide$GrowthPerDay_Control
FamMeansWide$OACont_Len<- (FamMeansWide$LarvaeDiamum_Exposed-FamMeansWide$LarvaeDiamum_Control)/FamMeansWide$LarvaeDiamum_Control
FamMeansWide$OACont_SA<- (FamMeansWide$LarvaeAreaum2_Exposed-FamMeansWide$LarvaeAreaum2_Control)/FamMeansWide$LarvaeAreaum2_Control
FamMeansWide$OACont_P<- (FamMeansWide$LarvaePerimeterum_Exposed-FamMeansWide$LarvaePerimeterum_Control)/FamMeansWide$LarvaePerimeterum_Control

write.csv(FamMeansWide,"~/R/GitHub/LarvaeTransGen2018/data/CalcPercDiffs.csv", row.names = FALSE)
```

# Visualize variance among families: Calcification
Right now this is using the mean for each jar (not all larvae measured could do that with LarvaeDat)
```{r Visualize Calcification Variance}
#EF01_EM01 only has one cross, so a boxplot won't work. Remove from the data before trying to make the plots: 
LarvByJarVarVizDat<- subset(LarvByJarDat, CrossID !="EF01_EM01_B1_QG")
ggplot(LarvByJarVarVizDat, aes(y=GrowthPerDay,x=as.factor(MaleIDPlain), color=JarTrt))+
  geom_boxplot(alpha=0.5, position="identity") +
  facet_grid(~FemIDPlain, scale="free_x")+ 
  labs(x = "Male ID", y= expression(paste("Growth (", mu, "m/day)")), title = "Female ID") + 
  scale_color_manual(name = "Larval Treatment", labels = c("Control", "OA"), values = c("#00BFC4", "#F8766D"))+
  geom_hline(yintercept=0)+
    theme_classic() +
  theme(axis.title.y.right = element_blank(),                # hide right axis title
        axis.text.y.right = element_blank(),                 # hide right axis labels
        axis.ticks.y = element_blank(),                      # hide left/right axis ticks
        axis.text.y = element_text(margin = margin(r = 0)),  # move left axis labels closer to axis
        panel.spacing = unit(0.7, "mm"),                    # remove spacing between facets
        strip.background = element_rect(size = 0.5), axis.text.x=element_text(angle=90),plot.title = element_text(size=10, hjust = 0.5)) 
  
ggplot(LarvByJarVarVizDat, aes(y=LarvaeAreaum2,x=MaleIDPlain, color=JarTrt))+
  geom_boxplot(alpha=0.5, position="identity") +
  facet_grid(~FemIDPlain, scale="free_x")+ 
  labs(x = "Male ID", y= expression(paste("Surface Area (", mu, m^2,")")), title = "Female ID") + 
  scale_color_manual(name = "Larval Treatment", labels = c("Control", "OA"), values = c("#00BFC4", "#F8766D"))+
    theme_classic() +
  theme(axis.title.y.right = element_blank(),                # hide right axis title
        axis.text.y.right = element_blank(),                 # hide right axis labels
        axis.ticks.y = element_blank(),                      # hide left/right axis ticks
        axis.text.y = element_text(margin = margin(r = 0)),  # move left axis labels closer to axis
        panel.spacing = unit(0.7, "mm"),                    # remove spacing between facets
        strip.background = element_rect(size = 0.5), axis.text.x=element_text(angle=90),plot.title = element_text(size=10, hjust = 0.5)) 

ggplot(LarvByJarVarVizDat, aes(y=LarvaeDiamum ,x=MaleIDPlain, color=JarTrt))+
  geom_boxplot(alpha=0.5, position="identity") +
  facet_grid(~FemIDPlain, scale="free_x")+ 
  labs(x = "Male ID", y= expression(paste("Length (", mu, "m)")), title = "Female ID") + 
  scale_color_manual(name = "Larval Treatment", labels = c("Control", "OA"), values = c("#00BFC4", "#F8766D"))+
    theme_classic() +
  theme(axis.title.y.right = element_blank(),                # hide right axis title
        axis.text.y.right = element_blank(),                 # hide right axis labels
        axis.ticks.y = element_blank(),                      # hide left/right axis ticks
        axis.text.y = element_text(margin = margin(r = 0)),  # move left axis labels closer to axis
        panel.spacing = unit(0.7, "mm"),                    # remove spacing between facets
        strip.background = element_rect(size = 0.5), axis.text.x=element_text(angle=90),plot.title = element_text(size=10, hjust = 0.5)) 

ggplot(LarvByJarVarVizDat, aes(y=LarvaePerimeterum ,x=MaleIDPlain, color=JarTrt))+
  geom_boxplot(alpha=0.5, position="identity") +
  facet_grid(~FemIDPlain, scale="free_x")+ 
  labs(x = "Male ID", y= expression(paste("Perimeter (", mu, "m)")), title = "Female ID") + 
  scale_color_manual(name = "Larval Treatment", labels = c("Control", "OA"), values = c("#00BFC4", "#F8766D"))+
    theme_classic() +
  theme(axis.title.y.right = element_blank(),                # hide right axis title
        axis.text.y.right = element_blank(),                 # hide right axis labels
        axis.ticks.y = element_blank(),                      # hide left/right axis ticks
        axis.text.y = element_text(margin = margin(r = 0)),  # move left axis labels closer to axis
        panel.spacing = unit(0.7, "mm"),                    # remove spacing between facets
        strip.background = element_rect(size = 0.5), axis.text.x=element_text(angle=90),plot.title = element_text(size=10, hjust = 0.5)) 

ggplot(LarvByJarVarVizDat, aes(y=(LarvaePerimeterum/LarvaeDiamum),x=MaleIDPlain, color=JarTrt))+
  geom_boxplot(alpha=0.5, position="identity") +
  facet_grid(~FemIDPlain, scale="free_x")+ 
  labs(x = "Male ID", y= paste("Perimeter:Length"), title = "Female ID") + 
  scale_color_manual(name = "Larval Treatment", labels = c("Control", "OA"), values = c("#00BFC4", "#F8766D"))+
  geom_hline(yintercept= 3.142)+
    theme_classic() +
  theme(axis.title.y.right = element_blank(),                # hide right axis title
        axis.text.y.right = element_blank(),                 # hide right axis labels
        axis.ticks.y = element_blank(),                      # hide left/right axis ticks
        axis.text.y = element_text(margin = margin(r = 0)),  # move left axis labels closer to axis
        panel.spacing = unit(0.7, "mm"),                    # remove spacing between facets
        strip.background = element_rect(size = 0.5), axis.text.x=element_text(angle=90),plot.title = element_text(size=10, hjust = 0.5)) 
```

Visualize larvae survival
```{r SurvivalVisualization}
SurvDat$SimpleCross<- paste(vapply(strsplit(SurvDat$CrossID, ""), '[', 4, FUN.VALUE = character(1)), vapply(strsplit(SurvDat$CrossID, ""), '[', 9, FUN.VALUE=character(1)), sep="_")
SurvDat$SimpleMale<-paste(vapply(strsplit(SurvDat$CrossID, ""), '[', 9, FUN.VALUE=character(1)))
SurvDat$SimpleFemale<- paste(vapply(strsplit(SurvDat$CrossID, ""), '[', 4, FUN.VALUE = character(1)))
SurvDat$ParentTrt<- factor(SurvDat$ParentTrt, levels=c("400", "2600"))
MeanSurvDat<- ddply(SurvWideDat, .(ParentTrt), numcolwise(mean, na.rm=T))
MeanSurvDat$ParentTrt<- factor(MeanSurvDat$ParentTrt, levels= c("400", "2600"))
SESurvDat<- ddply(SurvWideDat, .(ParentTrt), numcolwise(se, na.rm=T))
SESurvDat$ParentTrt<- factor(SESurvDat$ParentTrt, levels= c("400", "2600"))
#make reaction norms for survival color will be parent treatment. X will be jar treatment. Each cross will be its own line
ParTreat<- c("400"="Control-exposed Parents", "2600"="OA-exposed Parents")
ggplot(SurvDat,aes(x=JarTrt, y=LarvaeSurvived, color = SimpleCross, group= CrossID)) +
    geom_point(aes(color=SimpleCross))+
  geom_line(aes(color=SimpleCross, group=CrossID, linetype=SimpleFemale))+
  geom_errorbar(aes(ymin= LarvaeSurvived-SELarvaeSurvived, ymax=LarvaeSurvived+SELarvaeSurvived),width=0.05, position=position_dodge(0))+
    labs(x = "Larvae Treatment", y = "Number of Larvae Survived") +
  facet_grid(~ParentTrt)+
  theme_classic()
ggplot(SurvDat,aes(x=JarTrt, y=LarvaeSurvived, color = "black45", group= CrossID)) +
    geom_point(aes(color="black45"))+
  geom_line(aes(color="black45", group=CrossID))+
  geom_errorbar(aes(ymin= LarvaeSurvived-SELarvaeSurvived, ymax=LarvaeSurvived+SELarvaeSurvived),width=0.05, position=position_dodge(0))+
    labs(x = "Larvae Treatment", y = "Number of Larvae Survived") +
  scale_color_manual(values="black")+
  facet_grid(.~ParentTrt, labeller =labeller(ParentTrt = ParTreat))+
  theme_classic()
#try to plot survival 
ggplot(SurvWideDat,aes(x=ParentTrt, y=RatSurv, color = "gray45")) +
    geom_point(aes(color="gray45"))+
    labs(x = "Parent Treatment", y = "Proportion of Surviving Larvae") +
  scale_color_manual(values="gray45")+
  theme_classic()
#make bar graph of survival
ggplot(data = MeanSurvDat, aes(x = ParentTrt, y = RatSurv, fill="gray45")) +
  geom_bar(stat="identity", position="dodge")+
  ylim(c(0,1))+
  geom_hline(yintercept=1)+
  geom_errorbar(aes(ymin= RatSurv-SESurvDat$RatSurv, ymax=RatSurv+SESurvDat$RatSurv),width=0.2, position=position_dodge(.9))+
  labs(x = "Parent Treatment", y = "Proportion of Surviving Larvae") +
  scale_x_discrete(breaks=c("400","2600"),
        labels=c("Control", "OA"))+
  scale_fill_manual(values = "gray45") +
  theme_classic()

```

# Get ratio of OA to control to understand differences in calcification
```{r OA:Control Survival}
SurvWideDat$OACont_Perc<- (SurvWideDat$MeanLarvSurvivedExp-SurvWideDat$MeanLarvSurvivedCon)/SurvWideDat$MeanLarvSurvivedCon

write.csv(SurvWideDat,"~/R/GitHub/LarvaeTransGen2018/data/SurvDatPercent.csv", row.names = FALSE)

ggplot(LarvByJarVarVizDat, aes(y=LarvaeSurvived,x=MaleIDPlain, color=JarTrt))+
  geom_boxplot(alpha=0.5, position="identity") +
  facet_grid(~FemIDPlain, scale="free_x")+ 
  labs(x = "Male ID", y= paste("Number of Surviving Larvae"), title = "Female ID") + 
  scale_color_manual(name = "Larval Treatment", labels = c("Control", "OA"), values = c("#00BFC4", "#F8766D"))+
    theme_classic() +
  theme(axis.title.y.right = element_blank(),                # hide right axis title
        axis.text.y.right = element_blank(),                 # hide right axis labels
        axis.ticks.y = element_blank(),                      # hide left/right axis ticks
        axis.text.y = element_text(margin = margin(r = 0)),  # move left axis labels closer to axis
        panel.spacing = unit(0.7, "mm"),                    # remove spacing between facets
        strip.background = element_rect(size = 0.5), axis.text.x=element_text(angle=90),plot.title = element_text(size=10, hjust = 0.5)) 
```



```{r GetDescriptiveStats}
#get the mean length of the oysters collected (n=80)
mean(Adult_Sample$AdultLength*2.54)
sd(Adult_Sample$AdultLength*2.54)
se(Adult_Sample$AdultLength*2.54)
#get means for the data
meanJarTrt<- ddply(LarvaeDat, .(JarTrt, ParentTrt), numcolwise(mean, na.rm=T))
seJarTrt<- ddply(LarvaeDat, .(JarTrt, ParentTrt), numcolwise(se, na.rm=T))
mgpd<- subset(meanJarTrt, select= c(JarTrt, ParentTrt, GrowthPerDay, LarvaeDiamum, LarvaeAreaum2, LarvaePerimeterum))
(mgpd$GrowthPerDay[3]-mgpd$GrowthPerDay[4])/mgpd$GrowthPerDay[4]
(mgpd$GrowthPerDay[1]-mgpd$GrowthPerDay[2])/mgpd$GrowthPerDay[2]
(mgpd$LarvaeDiamum[3]-mgpd$LarvaeDiamum[4])/mgpd$LarvaeDiamum[4]
(mgpd$LarvaeDiamum[1]-mgpd$LarvaeDiamum[2])/mgpd$LarvaeDiamum[2]
mean(mgpd$LarvaeDiamum[1:2])/mean(mgpd$LarvaeDiamum[3:4])
(mgpd$LarvaeAreaum2[3]-mgpd$LarvaeAreaum2[4])/mgpd$LarvaeAreaum2[4]
(mgpd$LarvaeAreaum2[1]-mgpd$LarvaeAreaum2[2])/mgpd$LarvaeAreaum2[2]
mean(mgpd$LarvaeAreaum2[1:2])/mean(mgpd$LarvaeAreaum2[3:4])

(mgpd$LarvaePerimeterum[3]-mgpd$LarvaePerimeterum[4])/mgpd$LarvaePerimeterum[4]
(mgpd$LarvaePerimeterum[1]-mgpd$LarvaePerimeterum[2])/mgpd$LarvaePerimeterum[2]
mean(mgpd$LarvaePerimeterum[1:2])/mean(mgpd$LarvaePerimeterum[3:4])

mean(SurvWideDat$RatSurv, na.rm=TRUE)

#get difference in gpd within parental treatments
(mgpd$GrowthPerDay[4]- mgpd$GrowthPerDay[2])/mgpd$GrowthPerDay[2] #control parents
(mgpd$GrowthPerDay[3]- mgpd$GrowthPerDay[1])/mgpd$GrowthPerDay[1] #exposed parents
(mgpd$GrowthPerDay[3]- mgpd$GrowthPerDay[2])/mgpd$GrowthPerDay[2] #exposed parents exposed larvae vs. control parents control larvae
```

# Try to look at trade offs between growth and survival
```{r GPD and survival}
par(mfcol=c(1,1), mar= c(5.1, 5.1, 4.1, 2.1))
plot(LarvaeSurvived~GrowthPerDay, data=LarvByJarDat)
#Try using the mean number of surviving larvae in the control group for each cross and getting a proportion for each jar. 
LarvByJarDat$MeanControlSurvived<- SurvWideDat$MeanLarvSurvivedCon[match(LarvByJarDat$CrossID, SurvWideDat$CrossID)]
LarvByJarDat$PropSurviving<- LarvByJarDat$LarvaeSurvived/LarvByJarDat$MeanControlSurvived

#now plot that the proportion of surviving larvae by the mean growth per day for that jar
plot(PropSurviving~GrowthPerDay, data=LarvByJarDat)

#since there is a signficant interaction between parental treatment and larval treatment for GPD, try plotting those as different points
EPEL<- subset(LarvByJarDat, ParentTrt=="2600" & JarTrt=="Exposed")
EPCL<- subset(LarvByJarDat, ParentTrt=="2600" & JarTrt=="Control")
CPEL<- subset(LarvByJarDat, ParentTrt=="400" & JarTrt=="Exposed")
CPCL<- subset(LarvByJarDat, ParentTrt=="400" & JarTrt=="Control")

#I don't know how it is best to test for the effects of GPD on survival. Should I do separate tests for each treatment? I feel like that is the best approach. 

#start with tests for exposed parents exposed larvae: 
EPELmod<- lmer(PropSurviving~GrowthPerDay + (1|FemaleID) + (1|MaleID), data=EPEL)
EPELmodStep<- step(EPELmod)
EPELmodStep #note that GPD is marginally significant. 
EPELmod2<- lmer(PropSurviving~GrowthPerDay+ (1|FemaleID), data=EPEL)
summary(EPELmod2)
#check assumptions for EPELmod2
plot(EPELmod2) #looks fine
qqnorm(resid(EPELmod2))
qqline(resid(EPELmod2)) #also looks fine
acf(resid(EPELmod2)) #meets assumption
#meets assumptions
EPELmodStepChosen<- lmer(PropSurviving~(1|FemaleID), data=EPEL)
summary(EPELmodStepChosen)
#check assumptions of EPELmodStepChosen
plot(EPELmodStepChosen) #looks fine
qqnorm(resid(EPELmodStepChosen))
qqline(resid(EPELmodStepChosen)) #also looks fine
acf(resid(EPELmodStepChosen)) #meets assumption
#meets assumptions

#now look at exposed parents contorl larvae
EPCLmod<- lmer(PropSurviving~GrowthPerDay + (1|FemaleID) + (1|MaleID), data=EPCL)
EPCLmodStep<- step(EPCLmod)
EPCLmodStep #not mixed model chosen
EPCL_lm<- lm(PropSurviving~GrowthPerDay, data=EPCL)
summary(EPCL_lm) #not significant
par(mfcol=c(2,2))
plot(EPCL_lm) #looks like it meets assumptions
par(mfcol=c(1,1))
acf(resid(EPCL_lm)) 
#meets assumptions

#now look at control parents control larvae
CPCLmod<- lmer(PropSurviving~GrowthPerDay + (1|FemaleID) + (1|MaleID), data=CPCL)
CPCLmodStep<- step(CPCLmod)
CPCLmodStep #not mixed model chosen
CPCL_lm<- lm(PropSurviving~GrowthPerDay, data=CPCL)
summary(CPCL_lm) #not significant
#check assumptions
par(mfcol=c(2,2))
plot(CPCL_lm) #looks like it meets assumptions
par(mfcol=c(1,1))
acf(resid(CPCL_lm)) 
#meets assumptions

#now look at control parents exposed larvae
CPELmod<- lmer(PropSurviving~GrowthPerDay + (1|FemaleID) + (1|MaleID), data=CPEL)
CPELmodStep<- step(CPELmod)
CPELmodStep 
CPELmodFin<- lmer(PropSurviving~GrowthPerDay+ (1|FemaleID), data=CPEL)
summary(CPELmodFin)
#check assumptions
plot(CPELmodFin) #looks fine
qqnorm(resid(CPELmodFin))
qqline(resid(CPELmodFin)) #also looks fine
acf(resid(CPELmodFin)) #meets assumption
#meets assumptions
plot(PropSurviving~GrowthPerDay, data=EPEL, col="red", pch=19, ylim=c(0.0, 1.5), xlim=c(-4, 7), ylab="Proportion of Surviving Larvae\n (#Survived/Mean #SurvivedControlJar)", xlab="Shell Growth Rate (um/day)")
points(PropSurviving~GrowthPerDay, data=EPCL, col="blue", pch=15)
points(PropSurviving~GrowthPerDay, data=CPEL, col="orange", pch=17)
points(PropSurviving~GrowthPerDay, data=CPCL, col="purple", pch=18)
# Add a legend
legend(6, 0.45, legend=c("EPEL", "EPCL", "CPCL", "CPEL"),
       col=c("red", "blue", "purple", "orange"), pch=c(19,15,18,17), cex=0.6, bty="n")
#add line for significant lmer for CPEL
abline(a=0.87033, b=0.04093, col="orange")
#add line for marginally significant lmer for EPEL
abline(a=0.83220, b=0.05278, col="red", lty=2)
```

EPEL: GPD marginally significant (positive slope)
EPCL: GPD not significant
CPEL: GPD significant (p=0.0239) positive slope. 
CPCL: GPD not significant

Conclusion from the above models: For exposed larvae from control and exposed parents GPD is positively associated with proportion of surviving larvae. 

# Try to look at relationship between survival and growth ratios 
```{r Survival & Growth Ratio Comparison}
#try plotting proportion surviving by proportion growth change
plot(RatSurv~RatGPD, data=SurvWideDat, pch=19, xlab="Family Ratio Growth per Day (OA:Control)", ylab="Family Ratio Surviving Larvae (OA: Control)")
abline(a=0.87146, b=0.171172)
GPDSurvCross<- lmer(RatSurv~RatGPD*ParentTrt + (1|FemaleID) + (1|MaleID), data=SurvWideDat)
GPDSurvCrossStep<- step(GPDSurvCross)
GPDSurvCrossStep
GPDSurvCrossFin<- lmer(RatSurv~RatGPD + (1|FemaleID), data=SurvWideDat)
summary(GPDSurvCrossFin)
plot(GPDSurvCrossFin)
qqnorm(resid(GPDSurvCrossFin))
qqline(resid(GPDSurvCrossFin))
acf(resid(GPDSurvCrossFin))

#since there is an interaction between parental treatment within each parental treatment
ConParents<- subset(SurvWideDat, ParentTrt=="400" )
ExpParents<- subset(SurvWideDat, ParentTrt=="2600")

GPDSurvConParents<- lmer(RatSurv~RatGPD + (1|FemaleID) + (1|MaleID)+ (1|TrtTankID)+ (1|MaleTrtTankID), data=ConParents)
steppedConParents<- step(GPDSurvConParents)
steppedConParents
#only random effect of female ID
plot(RatSurv~RatGPD, data=ConParents, pch=19, xlab="Family Ratio Growth per Day (OA:Control)", ylab="Family Ratio Surviving Larvae (OA: Control)")

GPDSurvExpParents<- lmer(RatSurv~RatGPD + (1|FemaleID) + (1|MaleID)+ (1|TrtTankID)+ (1|MaleTrtTankID), data=ExpParents)
steppedExpParents<- step(GPDSurvExpParents)
steppedExpParents #not a mixed model
GPDSurvExpParentsLM<- lm(RatSurv~RatGPD, data=ExpParents)
summary(GPDSurvExpParentsLM)
plot(RatSurv~RatGPD, data=ExpParents, pch=19, col="red")
```


Export Figures
```{r ExportFigures, include=FALSE}
plots.dir.path<- list.files(tempdir(), pattern="rs-graphics", full.names= TRUE)
plots.png.paths<- list.files(plots.dir.path, pattern=".png", full.names=TRUE)
file.copy(from=plots.png.paths, to="../results")
```

Do simple t-tests on the means for oa-exposed and control exposed parents within each larvae treatment
```{r t-tests}
#get Bonferoni adjusted p-crit: 
0.05/10

Lexp<- subset(LarvaeDat, JarTrt=="Exposed")
Lcon<- subset(LarvaeDat, JarTrt=="Control")

t.test(GrowthPerDay~ParentTrt, data=Lexp)
t.test(GrowthPerDay~ParentTrt, data=Lcon)

t.test(LarvaeDiamum~ParentTrt, data=Lexp)
t.test(LarvaeDiamum~ParentTrt, data=Lcon)#not significantly different for control larvae

t.test(LarvaeAreaum2~ParentTrt, data=Lexp)
t.test(LarvaeAreaum2~ParentTrt, data=Lcon) #not signficantly different for control larvae

t.test(LarvaePerimeterum~ParentTrt, data=Lexp)
t.test(LarvaePerimeterum~ParentTrt, data=Lcon) #not significantly different for control larvae

t.test(PerimLenRat~ParentTrt, data=Lexp)
t.test(PerimLenRat~ParentTrt, data=Lcon)#not significantly different for control larvae
```
