---
title: "Larvae Functional Morphology"
output: html_document
---

# Load libraries
```{r}
library(tidyverse)
library(plyr)
library(sciplot)
library(reshape2)
library(lme4)
library(lmerTest)
library(blme)
library(grid)
library(data.table)
library(gridExtra)
library(lattice)
#library(optimx)
library(car)
library(Hmisc)
library(corrplot)
library(emmeans)


#make a skewness function
skewness<- function (x){
  result<-sum(((x-mean(x))/sd(x))^3, na.rm=TRUE)
  return(result)
}
```

# Load data
Name the data you export from FileMaker Pro by their exact table names and save them as CSVs, e.g. Larvae_Morphology.csv
```{r}
#setwd("~/Repos/LarvaeTransGen2018")
setwd("~/R/GitHub/LarvaeTransGen2018/data") #Elise's working directory
#upload all of the data tables for the larvae experiment
Larvae_Morphology <- read.csv("../data/Larvae_Morphology.csv") #contains data from CellProfiler from the larvae outlines
Barcode_Jar <- read.csv("../data/Barcode_Jar.csv") #contains data on CrossID, seatable, and whether or not larvae were present
Block_ID <- read.csv("../data/Block_ID.csv") #contains data on the block
Cross<- read.csv("../data/Cross.csv") #contains data on the female and male IDs for the cross and whether the cross was for QG or Meth
Fert_QG<- read.csv("../data/Fert_QG.csv") #contains data on fertilization counts, the JarIDs for the crosses
Header_WC<- read.csv("../data/Header_WC.csv") #contains data on the header tanks used to fill the jars
Larvae_Calcification <- read.csv("../data/Larvae_Calcification.csv") #contains data on filter weights and counts
Larvae_Counts <- read.csv("../data/Larvae_Counts.csv") #contains data on sedgewick rafter larvae counts
Larvae_WC <- read.csv("../data/Larvae_WC.csv") #contains data on larvae jar water chemistry
Parent_ID <- read.csv("../data/Parent_ID.csv") #contains data on the adults at the time of shucking. Includes Parent ID assignments
Larvae_Cilia<- read.csv("../data/Larvae_Cilia.csv") #contains data on cilia extrusion of the larvae photographed for morphology
Egg_Morphology<- read.csv("../data/Egg_Morphology.csv") #contains data on egg sizes from CellProfiler
Adult_Sample<- read.csv("../data/Adult_Sample.csv") #contains data on adult oysters at the time of collection
Storage_pH_Checks<- read.csv("../data/Storage_pH_Checks.csv") #contains data on the storage solution pH
WC_Standard<- read.csv("../data/WC_Standard.csv") #contains data on the standars used for water chemistry
ParentInfo<- merge(x=Parent_ID, y=Adult_Sample, by="AdultID", all.x=TRUE, all.y=TRUE)
```

Join the data into dataframes for analysis.
```{r}
#First dataframe will be for examining egg morphology for all eggs traced; multiple values for each adult
Eggs<- merge(x=Egg_Morphology, y=ParentInfo, by.x="FemaleID", by.y="ParentBlockID", all.x=TRUE)
#will have to add tank water chemistry data to this dataframe when it is available
#make ParentTrt a factor
Eggs$ParentTrt<- as.factor(Eggs$ParentTrt)

```

Visualize the data for eggs to start
```{r}
#There were some concerns at first that the eggs that were measured were too large. Elise looked into this in April 2020. The literature lists mature eggs for C. virginica as ~60 um in diameter. Check this against our mean eggs

meanegg<- aggregate(EggDiamum~BlockID, data= Eggs, FUN=mean)
meanegg # these means are in agreement with the average size of a mature egg. 

#now look at the data to find outliers. 
#look at the egg measurements for each female
boxplot(EggDiamum~FemaleID, data=Eggs)
boxplot(EggDiamum~BlockID*ParentTrt, data=Eggs)

#check for outliers using code from https://www.r-bloggers.com/identify-describe-plot-and-remove-the-outliers-from-the-dataset/. The function uses Tukey's method to ID ouliers ranged above and below the 1.5*IQR. 
outlierKD <- function(dt, var) {
     var_name <- eval(substitute(var),eval(dt))
     na1 <- sum(is.na(var_name))
     m1 <- mean(var_name, na.rm = T)
     par(mfrow=c(2, 2), oma=c(0,0,3,0))
     boxplot(var_name, main="With outliers")
     hist(var_name, main="With outliers", xlab=NA, ylab=NA)
     outlier <- boxplot.stats(var_name)$out
     mo <- mean(outlier)
     var_name <- ifelse(var_name %in% outlier, NA, var_name)
     boxplot(var_name, main="Without outliers")
     hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
     title("Outlier Check", outer=TRUE)
     na2 <- sum(is.na(var_name))
     cat("Outliers identified:", na2 - na1, "n")
     cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "n")
     cat("Mean of the outliers:", round(mo, 2), "n")
     m2 <- mean(var_name, na.rm = T)
     cat("Mean without removing outliers:", round(m1, 2), "n")
     cat("Mean if we remove outliers:", round(m2, 2), "n")
     response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
     if(response == "y" | response == "yes"){
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     } else{
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     }
}

outlierKD(Eggs, EggDiamum)
#I opted to remove outliers for now. 
#Outliers identified: 131 nPropotion (%) of outliers: 11.1 nMean of the outliers: 53.51 nMean without removing outliers: 61.97 nMean if we remove outliers: 62.9 n
```


```{r}

par(mfrow=c(1,1))
#plot egg data with outliers removed
ggplot(Eggs, aes(x=EggDiamum, color=ParentTrt))+
  geom_histogram(alpha=0.5, position="identity") +
  facet_grid(~BlockID)
#we can see that the egg data are slightly left skewed 
eggmeans<- aggregate(EggDiamum~ParentTrt, data=Eggs, FUN=mean) #400 = 63.08564; 2800= 62.75694
eggmedians<- aggregate(EggDiamum~ParentTrt, data=Eggs, FUN=median)#400 = 62.54237; 2800= 62.42467
#I think I can use the mean egg size if the outliers are removed

#check the updated data: 
boxplot(EggDiamum~FemaleID, data=Eggs)
boxplot(Eggs$EggDiamum~Eggs$BlockID, xlab="Block ID", ylab="Egg Diameter um")

```

Make datasets without the egg size outliers
```{r}
#start by removing the outliers from Egg_Morphology
outlierKD(Egg_Morphology, EggDiamum)
meanegg<- ddply(Egg_Morphology, .(FemaleID), numcolwise(mean, na.rm=T)) #get mean egg size
FemAdults<- merge(x=ParentInfo, y= meanegg, by.x="ParentBlockID", by.y= "FemaleID", all.y=TRUE)
#will have to add tank water chemistry data to this dataframe when it is available
#make ParentTrt a factor
FemAdults$ParentTrt<- as.factor(FemAdults$ParentTrt)
#Next dataframe will be for one line per traced larva, includes all the data for that jar, including average egg size for that jar
#Eggs dataframe has all of the info that we need for the 
Larvae<-merge(x=Larvae_Morphology, y=Barcode_Jar, by="JarID", all.x=TRUE)
#use gather function in tidyr to make the Fert_QG Jar ID columns in long form vs. wide form
Fert_QG_long <- Fert_QG %>% tidyr::gather(Key, JarID, JarID1:JarID6)
#use the gathered data frame to merge with Larvae
Larvae<-merge(x=Larvae, y=Fert_QG_long, by.x= c("JarID", "CrossID"), by.y=c("JarID", "CrossID"), all.x=TRUE) #did the join by two columns so CrossID column wasn't duplicated
Larvae<- merge(x=Larvae, y=Cross, by="CrossID", all.x=TRUE)
Larvae<- merge(x=Larvae, y=FemAdults, by.x= "FemaleID", by.y="ParentBlockID", all.x=TRUE)
Larvae<- merge(x=Larvae, y=Larvae_WC, by= "JarID", all.x=TRUE)
Larvae$JarSeatable<- as.factor(Larvae$JarSeatable)#make seatable a factor

#Final dataframe will be one line per Jar, includes all data for the jar, including the average larvae size per jar and cilia
LarvByJar<-merge(x=Larvae_Calcification, y=Larvae_Cilia, by="JarID", all.x=TRUE, all.y=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Barcode_Jar, by="JarID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Fert_QG_long, by.x=c("JarID", "CrossID"), by.y=c("JarID","CrossID"), all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Cross, by="CrossID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=FemAdults, by.x="FemaleID", by.y="ParentBlockID", all.x=TRUE)
#get the means for larvae morphology and add to LarvByJar
MeanLarvMorph<- ddply(Larvae_Morphology, .(JarID), numcolwise(mean, na.rm=T))
SELarvMorph<- ddply(Larvae_Morphology, .(JarID), numcolwise(se, na.rm=T)) #get standard error of larvae size
MeanLarvMorph<- full_join(MeanLarvMorph,SELarvMorph, by=c("JarID", "JarID"),suffix=c("","SE")) #join the mean and se larvae data
LarvByJar<- merge(x=LarvByJar, y=MeanLarvMorph, by="JarID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Larvae_WC, by="JarID", all.x=TRUE)
```


Look at autocorrelation of the larvae cilia data
```{r}
cilaut1<- lm(PercentCilia~TrtTankID, data=LarvByJar)
acf(resid(cilaut1)) #this looks fine to me
cilaut2<- lm(PercentCilia~AccTankID, data=LarvByJar)
acf(resid(cilaut2)) #looks fine

cilaut3<- lm(PercentCilia~TimeFilter, data=LarvByJar)
acf(resid(cilaut3)) #looks fine

cilaut4<- lm(PercentCilia~JarSeatable, data=LarvByJar)
acf(resid(cilaut4)) #looks fine

```

Visualize the data for functional morphology
```{r}
ggplot(LarvByJar, aes(x=PercentCilia, color=ParentTrt))+
  geom_histogram(alpha=0.5, position="identity") +
  facet_grid(~JarTrt)

#just do a prelim anova to look for differences
cilaov<- aov(PercentCilia~ParentTrt*JarTrt, data=LarvByJar)
shapiro.test(cilaov$residuals) #doesn't meet assumption
leveneTest(cilaov) # meets assumption
anova(cilaov)

#create most complex model and then use the step function to see which model is best fit for the data; these are for ParentTrt and JarTrt as fixed effects not covariates
larvcil<-lmer(PercentCilia~ParentTrt*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|TimeFilter)+(1|TrtTankID)+(1|AccTankID), data=LarvByJar)

#check assumptions of the model
plot(larvcil) #this looks okay to me
qqnorm(resid(larvcil))
qqline(resid(larvcil)) #look at the two points that are way off the line

summary(larvcil)

```

Plot cilia data
```{r}
meancil1<- aggregate(PercentCilia~JarTrt*ParentTrt, data=LarvByJar, FUN=mean)
secil1<- aggregate(PercentCilia~JarTrt*ParentTrt, data=LarvByJar, FUN=se)

#use a bargraph to plot cilia extrusion +/- SE
ggplot(data = meancil1, aes(x = ParentTrt, y = PercentCilia, fill=JarTrt)) +
  geom_bar(stat="identity", aes(fill=JarTrt), position="dodge")+
  labs(x = "Parent Treatment", y = "Proportion of Larvae with Cilia Extruded") +
  scale_fill_manual(name = "Larvae Treatment", labels = c("Control", "OA"), values = c("skyblue2", "red2")) +
  geom_errorbar(aes(ymin= meancil1$PercentCilia-secil1$PercentCilia, ymax=meancil1$PercentCilia+secil1$PercentCilia),width=0.2, position=position_dodge(.9))+
  theme_classic()

meancil<- aggregate(PercentCilia~JarTrt, data=LarvByJar, FUN=mean)
secil<- aggregate(PercentCilia~JarTrt, data=LarvByJar, FUN=se)
#plot without parent treatment
ggplot(data = meancil, aes(x = JarTrt, y = PercentCilia)) +
  geom_bar(stat="identity", position="dodge")+
  labs(x = "Larvae Treatment", y = "Proportion of Larvae with Cilia Extruded") +
  scale_fill_manual(values = "gray45") +
  geom_errorbar(aes(ymin= meancil$PercentCilia-secil$PercentCilia, ymax=meancil$PercentCilia+secil$PercentCilia),width=0.2, position=position_dodge(.9))+
  theme_classic()

```

Try using generalized lmm for the data with a binomial distribution
```{r}
glcil<- glmer(as.integer(PercentCilia)~ParentTrt*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|TimeFilter)+(1|TrtTankID)+(1|AccTankID), data=LarvByJar, family=binomial)

summary(glcil)
plot(glcil)

```

