---
title: "Larvae Biomineralization Block 1"
output: html_document
---

Load libraries
```{r LoadLibraries, include=FALSE}
library(knitr)
library(cowplot)
library(plyr)
library(tidyverse)
library(sciplot)
library(reshape2)
library(lme4)
library(lmerTest)
library(blme)
library(grid)
library(data.table)
library(gridExtra)
library(lattice)
library(optimx)
library(car)
library(Hmisc)
library(corrplot)
library(emmeans)
library(chron)
library(GLMMadaptive)
library(calibrate)
library(merTools)
library(bootpredictlme4)
library(visreg)
```

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

Load data
Name the data you export from FileMaker Pro by their exact table names and save them as CSVs, e.g. Larvae_Morphology.csv
```{r LoadData}
#setwd("~/Repos/LarvaeTransGen2018")
setwd("~/R/GitHub/LarvaeTransGen2018/data") #Elise's working directory
#upload all of the data tables for the larvae experiment
Larvae_Morphology <- read.csv("../data/Larvae_morphology.csv") #contains data from CellProfiler from the larvae outlines
Barcode_Jar <- read.csv("../data/Barcode_Jar.csv") #contains data on CrossID, seatable, and whether or not larvae were present
Block_ID <- read.csv("../data/Block_ID.csv") #contains data on the block
Cross<- read.csv("../data/Cross.csv") #contains data on the female and male IDs for the cross and whether the cross was for QG or Meth
Fert_QG<- read.csv("../data/Fert_QG.csv") #contains data on fertilization counts, the JarIDs for the crosses
Header_WC<- read.csv("../data/Header_WC.csv") #contains data on the header tanks used to fill the jars
Larvae_Calcification <- read.csv("../data/Larvae_Calcification.csv") #contains data on filter weights and counts
Larvae_Calcification<- subset(Larvae_Calcification, JarID != "")
Larvae_Counts <- read.csv("../data/Larvae_Counts.csv") #contains data on sedgewick rafter larvae counts
Larvae_WC <- read.csv("../data/Larvae_WC.csv") #contains data on larvae jar water chemistry
Parent_ID <- read.csv("../data/Parent_ID.csv") #contains data on the adults at the time of shucking. Includes Parent ID assignments
Larvae_Cilia<- read.csv("../data/Larvae_Cilia.csv") #contains data on cilia extrusion of the larvae photographed for morphology
Egg_Morphology<- read.csv("../data/Egg_Morphology.csv") #contains data on egg sizes from CellProfiler
Egg_Morphology$EggPerimeterum<- Egg_Morphology$EggPerimeterPix/8.892  #add a column to convert egg perimeter to um based on the scaled slide
Egg_Morphology$EggPerimDiamRat<- Egg_Morphology$EggPerimeterum/Egg_Morphology$EggDiamum #add a column for the ratio of perimeter to diameter
Adult_Sample<- read.csv("../data/Adult_Sample.csv") #contains data on adult oysters at the time of collection
WC_Standard<- read.csv("../data/WC_Standard.csv") #contains data on the standards used for water chemistry
Tank_WC<- read.csv("../data/Tank_WC.csv") #contains data on the adult tank water chemistry
CiliaArea<- read.csv("../data/Cilia_AreaLength.csv")
```

Make fields factors: 
```{r MakeFieldsFactors}
Adult_Sample$AccTankID<- as.factor(Adult_Sample$AccTankID)
Adult_Sample$TrtTankID<- as.factor(Adult_Sample$TrtTankID)
Barcode_Jar$JarSeatable<- as.factor(Barcode_Jar$JarSeatable)
Barcode_Jar<- subset(Barcode_Jar, JarSeatable =="2" | JarSeatable =="4" | JarSeatable =="6")
Tank_WC$ParentTrt<- as.factor(Tank_WC$ParentTrt)
Tank_WC$Tank<- as.factor(Tank_WC$Tank)
WC_Standard$CRM<- as.factor(WC_Standard$CRM)
Larvae_WC$DateFilter<- as.factor(Larvae_WC$DateFilter)
CiliaArea$LarvaeMorphID<- as.factor(CiliaArea$LarvaeMorphID)
CiliaArea$CilType<- as.factor(CiliaArea$CilType)
Larvae_Morphology$LarvaeMorphID<- as.factor(Larvae_Morphology$LarvaeMorphID)
```

Look at the water chemistry data for adult tanks
```{r AdultTankChemistry}
#subset the tank data to only include B1 
Tank_WCB1<- subset(Tank_WC, UseForCalcite == "TRUE")
Tank_WCB1$Tank<-droplevels(Tank_WCB1)$Tank #drop unused levels
Tank_WCB1$ParentTrt<- droplevels(Tank_WCB1)$ParentTrt
#test to see if tanks are significantly different. 
boxplot(SatCalcite_ATDIC~Tank, data=Tank_WCB1, na.omit=TRUE)
#check for significant differences in saturation state between tanks within treatments
Ca1<- lmer(SatCalcite_ATDIC~ParentTrt + (1|Tank), data=Tank_WCB1)
Ca2<- lmer(SatCalcite_ATDIC~ 1 + (1|Tank), data= Tank_WCB1)
anova(Ca1, Ca2) #select Ca1
Ca3<- lm(SatCalcite_ATDIC~ParentTrt, data=Tank_WCB1)
anova(Ca1, Ca3) #select Ca3
#check assumptions
par(mfcol=c(2,2))
plot(Ca3) #seems to meet assumptions of homoscedasticity and linearity. Normality could look better. 
par(mfcol=c(1,1))
acf(resid(Ca3)) #this looks good. 
summary(Ca3) #There is a significant effect of ParentTrt, but not tank according to best model. 
#try a transformation to see if it helps normality
Ca1log<- lmer(log(SatCalcite_ATDIC)~ParentTrt+(1|Tank), data=Tank_WCB1)
Ca3log<- lm(log(SatCalcite_ATDIC)~ParentTrt, data=Tank_WCB1)
anova(Ca1log, Ca3log) #still choose Ca3log
par(mfcol=c(2,2))
plot(Ca3log) #now meets assumptions 
par(mfcol=c(1,1))
acf(resid(Ca3log))#looks good
summary(Ca3log)
#create dataframe that has the mean values for each tank
TankMeans<- ddply(Tank_WCB1, .(Tank, Block, ParentTrt), numcolwise(mean, na.rm=T))#can't use this for pH and salinity, remember, it is just for carb chem
```

Join the data into dataframe for Egg analysis
```{r JoinDataForEggAnalysis}
#First dataframe will be for examining egg morphology for all eggs traced; multiple values for each adult
AdultSampParID<- merge(x=Adult_Sample, y=Parent_ID, by="AdultID", all.x=TRUE)
AdultSampParID<- merge(x=AdultSampParID, y=TankMeans, by.x="TrtTankID", by.y="Tank", all.x=TRUE)
AdultSampParID<- subset(AdultSampParID, ParentBlockID !="NA")
AdultSampParID<- subset(AdultSampParID, Block =="B1")
MaleInfo<- subset(AdultSampParID, Sex=="M")
FemaleInfo<- subset(AdultSampParID, Sex=="F")
colnames(MaleInfo)<- paste("Male", colnames(MaleInfo), sep="")
Eggs<- merge(x=Egg_Morphology, y=FemaleInfo, by.x="FemaleID", by.y="ParentBlockID", all.x=TRUE)
Eggs$FemaleID<- as.factor(Eggs$FemaleID)
#subset the data to only include B1
EggsB1<- subset(Eggs, Block =="B1")
```

Visualize the data for eggs to start
```{r EggVisualization}
#get the original range of the eggs
range(EggsB1$EggDiamum)
#look at the egg measurements to see outliers
boxplot(EggDiamum~FemaleID, data=EggsB1)
boxplot(EggDiamum~ParentTrt, data=EggsB1)
```

Remove egg outliers
```{r EggOutlierRemoval}
#check for outliers using code from https://www.r-bloggers.com/identify-describe-plot-and-remove-the-outliers-from-the-dataset/. The function uses Tukey's method to ID ouliers ranged above and below the 1.5*IQR. 
outlierKD <- function(dt, var) {
     var_name <- eval(substitute(var),eval(dt))
     na1 <- sum(is.na(var_name))
     m1 <- mean(var_name, na.rm = T)
     par(mfrow=c(2, 2), oma=c(0,0,3,0))
     boxplot(var_name, main="With outliers")
     hist(var_name, main="With outliers", xlab=NA, ylab=NA)
     outlier <- boxplot.stats(var_name)$out
     mo <- mean(outlier)
     var_name <- ifelse(var_name %in% outlier, NA, var_name)
     boxplot(var_name, main="Without outliers")
     hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
     title("Outlier Check", outer=TRUE)
     na2 <- sum(is.na(var_name))
     cat("Outliers identified:", na2 - na1, "n")
     cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "n")
     cat("Mean of the outliers:", round(mo, 2), "n")
     m2 <- mean(var_name, na.rm = T)
     cat("Mean without removing outliers:", round(m1, 2), "n")
     cat("Mean if we remove outliers:", round(m2, 2), "n")
     response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
     if(response == "y" | response == "yes"){
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     } else{
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     }
}
outlierKD(EggsB1, EggDiamum) #I opted to remove outliers for now. 
#Outliers identified: 48 nPropotion (%) of outliers: 11.5 nMean of the outliers: 57.63 nMean without removing outliers: 62.4 nMean if we remove outliers: 62.95 n
yes #to remove outliers from EggsB1
#count how many eggs are left per female after removing the outliers. 
outrem<- aggregate(EggDiamum~FemaleID, data=EggsB1, FUN=length)
#EF07_B1 had 10 eggs that were included the rest were removed. Keep EF07_B1 in the analysis
```

Visualize eggs without outliers
```{r EggVisualizationNoOutliers}
#Plot again without outliers
par(mfcol=c(1,1))
boxplot(EggDiamum~FemaleID, data=EggsB1) 
boxplot(EggDiamum~ParentTrt, data=EggsB1)
#get the range now
range(EggsB1$EggDiamum, na.rm=TRUE)
#note that the eggs were filtered through a 70 um filter, but it appears that larger ones made it through.We are going to keep these in for now because they aren't that outrageously large. 
#plot histograms of the egg data with outliers removed
ggplot(EggsB1, aes(x=EggDiamum, color=ParentTrt))+
  geom_histogram(alpha=0.5, position="identity")
eggmeans<- aggregate(EggDiamum~ParentTrt, data=EggsB1, FUN=mean) #400 = 63.45709; 2800= 62.36829
eggmedians<- aggregate(EggDiamum~ParentTrt, data=EggsB1, FUN=median)#400 = 62.91200; 2800= 62.06727
view(eggmeans)
view(eggmedians)
#I will use the mean egg size. 
```

Make datasets without the egg size outliers
```{r MakeLarvaeDataframes}
meanegg<- ddply(EggsB1, .(FemaleID), numcolwise(mean, na.rm=T)) #get mean egg size
#select only columns that we need for egg morph
meanegg<- subset(meanegg, select=c("FemaleID", "EggDiamum"))
FemAdults<- merge(x=FemaleInfo, y= meanegg, by.x="ParentBlockID", by.y= "FemaleID", all.y=TRUE, all.x=TRUE)
#get the cilia area and extent data ready
CiliaExtent<- subset(CiliaArea, CilType=="Extent")
CiliaArea<- subset(CiliaArea, CilType=="Area")
#now use ddply to get only one value for each morphID
CiliaExtentAll<- ddply(CiliaExtent, .(LarvaeMorphID), numcolwise(mean, na.rm=TRUE))
CiliaAreaAll<- ddply(CiliaArea, .(LarvaeMorphID), numcolwise(mean, na.rm=TRUE))
colnames(CiliaExtentAll)<- paste(colnames(CiliaExtentAll), "ciliaextent", sep="")
colnames(CiliaAreaAll)<- paste(colnames(CiliaAreaAll), "ciliaarea", sep="")
#merge the cilia info to larvae morphology
Larvae_Morphology<- merge(Larvae_Morphology, CiliaExtentAll, by.x="LarvaeMorphID", by.y="LarvaeMorphIDciliaextent", all.x=TRUE)
Larvae_Morphology<- merge(Larvae_Morphology, CiliaAreaAll, by.x="LarvaeMorphID", by.y="LarvaeMorphIDciliaarea", all.x=TRUE)
#Next dataframe will be for one line per traced larva, includes all the data for that jar, including average egg size for that jar 
Larvae<-merge(x=Larvae_Morphology, y=Barcode_Jar, by="JarID", all.x=TRUE)
Larvae<- merge(x=Larvae, y=Larvae_WC, by="JarID", all.x=TRUE, all.y=TRUE)
#use gather function in tidyr to make the Fert_QG Jar ID columns in long form vs. wide form
Fert_QG_long <- Fert_QG %>% tidyr::gather(Key, JarID, JarID1:JarID6)
#use the gathered data frame to merge with Larvae
Larvae<-merge(x=Larvae, y=Fert_QG_long, by.x= c("JarID", "CrossID"), by.y=c("JarID", "CrossID"), all.x=TRUE) #did the join by two columns so CrossID column wasn't duplicated
Larvae<- merge(x=Larvae, y=Cross, by="CrossID", all.x=TRUE)
Larvae<- merge(x=Larvae, y=FemAdults, by.x= "FemaleID", by.y="ParentBlockID", all.x=TRUE)
Larvae<- merge(x=Larvae, y=MaleInfo, by.x= "MaleID", by.y="MaleParentBlockID", all.x=TRUE)
Larvae<- merge(x=Larvae, y= Larvae_Counts, by="JarID", all.x=TRUE)
#Remove jars without larvae
Larvae<- subset(Larvae, Larvae=="TRUE")
#get the growth per day for the larva
Larvae$GrowthPerDay<- (Larvae$LarvaeDiamum-Larvae$EggDiamum)/3
Larvae$Areaciliaareaum<- Larvae$Areaciliaarea/Larvae$LarvaePixPerum^2
Larvae$Perimeterciliaareaum<- Larvae$Perimeterciliaarea/Larvae$LarvaePixPerum
Larvae$Perimeterciliaextentum<- Larvae$Perimeterciliaextent/Larvae$LarvaePixPerum
Larvae$MajorAxisLengthciliaextentum<- Larvae$MajorAxisLengthciliaextent/Larvae$LarvaePixPerum
#Final dataframe will be one line per Jar, includes all data for the jar, including the average larva size per jar and cilia
LarvByJar<- merge(x=Larvae_Calcification, y=Larvae_Cilia, by="JarID", all.x=TRUE, all.y=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Barcode_Jar, by="JarID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Larvae_WC, by="JarID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Fert_QG_long, by.x=c("JarID", "CrossID"), by.y=c("JarID","CrossID"), all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Cross, by="CrossID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=FemAdults, by.x="FemaleID", by.y="ParentBlockID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=MaleInfo, by.x="MaleID", by.y="MaleParentBlockID", all.x=TRUE)
LarvByJar<- merge(x=LarvByJar, y=Larvae_Counts, by="JarID", all.x=TRUE)
#get the means for larvae morphology and add to LarvByJar
MeanLarvMorph<- ddply(Larvae_Morphology, .(JarID), numcolwise(mean, na.rm=T))
SELarvMorph<- ddply(Larvae_Morphology, .(JarID), numcolwise(se, na.rm=T)) #get standard error of larva size
MeanLarvMorph<- full_join(MeanLarvMorph,SELarvMorph, by=c("JarID", "JarID"),suffix=c("","SE")) #join the mean and se larvae data
LarvByJar<- merge(x=LarvByJar, y=MeanLarvMorph, by="JarID", all.x=TRUE)
#Now let's add the survival data. See notes on "Larvae Survival" for more info on how we decided to count total larvae in the jar
#Make columns for the two larvae counts for 1_5 and 2
LarvByJar$v1_5SurvCount<- rowSums(LarvByJar[,c("SWRTotal", "F2LarvaeCount")], na.rm=TRUE)
LarvByJar$v2SurvCount<- rowSums(LarvByJar[,c("F1LarvaeCount", "F2LarvaeCount")], na.rm=TRUE)
#use ifelse statement to define the LarvaeSurvived column
LarvByJar$LarvaeSurvived<- ifelse(LarvByJar$ProtocolVersion =="1", paste(LarvByJar$TotalLarvae), ifelse(LarvByJar$ProtocolVersion =="1_5", paste(LarvByJar$v1_5SurvCount), paste(LarvByJar$v2SurvCount)))
LarvByJar$LarvaeSurvived<- as.numeric(LarvByJar$LarvaeSurvived)
#get the total larvae for the control jars and exposed jars for each family and then get the ratio of exposed to control to get an idea of how good a male and female pair are at larvae surviving. Can also use this to create reaction norms. 
SurvDat<- ddply(LarvByJar, .(CrossID, ParentTrt, JarTrt, MaleID, FemaleID, TrtTankID,MaleTrtTankID, AccTankID, Block), numcolwise(mean, na.rm=TRUE))
SESurvDat<- ddply(LarvByJar, .(CrossID, ParentTrt, JarTrt, Block), numcolwise(se, na.rm=TRUE)) #get the standard errors for each mean count
SESurvDat<- subset(SESurvDat, select= c(CrossID, JarTrt, LarvaeSurvived, SatCalcite_ATDIC, JarSatCalcite_SW)) #rename the columns
colnames(SESurvDat)<- paste("SE", colnames(SESurvDat), sep="")
#merge the standard error data to the SurvDat dataframe
SurvDat<- merge(SurvDat, SESurvDat, by.x= c("CrossID", "JarTrt"), by.y= c("SECrossID", "SEJarTrt"))
#only keep the relevant fields because otherwise this is a huge dataframe
SurvDat<- subset(SurvDat, select= c(CrossID, JarTrt, ParentTrt, MaleID, FemaleID, TrtTankID,MaleTrtTankID, AccTankID, LarvaeSurvived, Block, SELarvaeSurvived, SatCalcite_ATDIC)) 
SurvDat<- subset(SurvDat, Block=="B1")
SurvDat<- subset(SurvDat, CrossID!="")
#use the spread function to go from long form to wide form for the survival data
SurvDatToWide<- unite(SurvDat, Value, c(LarvaeSurvived, SELarvaeSurvived), sep="_")
SurvWideDat<- spread(SurvDatToWide, JarTrt, value= Value)
#now split the mean and SE apart again
SurvWideDat<- separate(SurvWideDat, Control, into= c("MeanLarvSurvivedCon", "SELarvSurvivedCon"), sep="_")
SurvWideDat<- separate(SurvWideDat, Exposed, into= c("MeanLarvSurvivedExp", "SELarvSurvivedExp"), sep="_")
#make the means and ses numeric
SurvWideDat$MeanLarvSurvivedExp<- as.numeric(SurvWideDat$MeanLarvSurvivedExp)
SurvWideDat$MeanLarvSurvivedCon<- as.numeric(SurvWideDat$MeanLarvSurvivedCon)
SurvWideDat$SELarvSurvivedExp<- as.numeric(SurvWideDat$SELarvSurvivedExp)
SurvWideDat$SELarvSurvivedCon<- as.numeric(SurvWideDat$SELarvSurvivedCon)
SurvWideDat$RatSurv<- SurvWideDat$MeanLarvSurvivedExp/SurvWideDat$MeanLarvSurvivedCon
SurvWideDat$SurvChange<- (SurvWideDat$MeanLarvSurvivedCon-SurvWideDat$MeanLarvSurvivedExp)/SurvWideDat$MeanLarvSurvivedCon
SurvWideDat$CrossID<- as.factor(SurvWideDat$CrossID)
SurvWideDat$FemaleID<- as.factor(SurvWideDat$FemaleID)
SurvWideDat$MaleID<- as.factor(SurvWideDat$MaleID)
LarvByJar<- subset(LarvByJar, Larvae=="TRUE")
#now remove the data we do not want to analyze.
LarvaeDat<- subset(Larvae, Block == "B1")
LarvaeDat$JarID<- as.factor(LarvaeDat$JarID)
LarvaeDat$JarID<-droplevels(LarvaeDat)$JarID
LarvaeDat$AccTankID<- as.factor(LarvaeDat$AccTankID)
LarvaeDat$AccTankID<-droplevels(LarvaeDat)$AccTankID
LarvaeDat$TrtTankID<- as.factor(LarvaeDat$TrtTankID)
LarvaeDat$TrtTankID<-droplevels(LarvaeDat)$TrtTankID
LarvaeDat$MaleTrtTankID<- as.factor(LarvaeDat$MaleTrtTankID)
LarvaeDat$MaleTrtTankID<-droplevels(LarvaeDat)$MaleTrtTankID
LarvaeDat$MaleAccTankID<- as.factor(LarvaeDat$MaleAccTankID)
LarvaeDat$MaleAccTankID<-droplevels(LarvaeDat)$MaleAccTankID
LarvaeDat$JarSeatable<- as.factor(LarvaeDat$JarSeatable)
LarvaeDat$JarSeatable<-droplevels(LarvaeDat)$JarSeatable
LarvaeDat$JarTrt<- as.factor(LarvaeDat$JarTrt)
LarvaeDat$ParentTrt<- as.factor(LarvaeDat$ParentTrt)
LarvaeDat$FemaleID<- as.factor(LarvaeDat$FemaleID)
LarvaeDat$FemaleID<-droplevels(LarvaeDat)$FemaleID
LarvaeDat$MaleID<- as.factor(LarvaeDat$MaleID)
LarvaeDat$MaleID<-droplevels(LarvaeDat)$MaleID
LarvaeDat$CrossID<- as.factor(LarvaeDat$CrossID)
LarvaeDat$CrossID<-droplevels(LarvaeDat)$CrossID
LarvByJarDat<- subset(LarvByJar, Block == "B1")
LarvByJarDat$JarID<- as.factor(LarvByJarDat$JarID)
LarvByJarDat$AccTankID<- as.factor(LarvByJarDat$AccTankID)
LarvByJarDat$TrtTankID<- as.factor(LarvByJarDat$TrtTankID)
LarvByJarDat$MaleAccTankID<- as.factor(LarvByJarDat$MaleAccTankID)
LarvByJarDat$MaleTrtTankID<- as.factor(LarvByJarDat$MaleTrtTankID)
LarvByJarDat$JarSeatable<- as.factor(LarvByJarDat$JarSeatable)
LarvByJarDat$ParentTrt<- as.factor(LarvByJarDat$ParentTrt)
LarvByJarDat$FemaleID<- as.factor(LarvByJarDat$FemaleID)
LarvByJarDat$MaleID<- as.factor(LarvByJarDat$MaleID)
LarvByJarDat$CrossID<- as.factor(LarvByJarDat$CrossID)
LarvByJarDat$JarTrt<- as.factor(LarvByJarDat$JarTrt)
#make a column for the ratio of major to minor axis
LarvaeDat$MajMinRat<- LarvaeDat$LarvaeMajorAxisLengthPix/LarvaeDat$LarvaeMinorAxisLengthPix
LarvaeDat$PerimLenRat<-LarvaeDat$LarvaePerimeterum/LarvaeDat$LarvaeDiamum
LarvByJarDat<- subset(LarvByJarDat, JarSatCalcite_SW !="NA")
```

Analyze jar chemistry data
```{r JarChemistryStat}
CalJar<- subset(LarvByJarDat, BottleChemistry=="TRUE" )#only use those that had their chem measured
#look at the saturation state for jars
boxplot(JarSatCalcite_SW~JarTrt*JarSeatable, data=CalJar)
#I was having trouble with making this model: JarSatCalcite_SW~JarTrt + (1|ParentTrt)+ (1|TimeFilter), so I looked at TimeFilter because I think that is the problem. 
Jar1<- lm(JarSatCalcite_SW~TimeFilter, data=CalJar)
par(mfcol=c(2,2))
plot(Jar1)
par(mfcol=c(1,1))
#looks like we can't use timefilter because it is too specific to each jar. Get the error message "observations with leverage one"; try using seatable instead
Jar2<- lmer(JarSatCalcite_SW~JarTrt + (1|JarSeatable), data=CalJar)
Jar2step<- step(Jar2)
#get error model not of class lmerMod. I looked up this issue and turns out that error results when all random effects are dropped. See https://stackoverflow.com/questions/55638476/how-to-do-stepwise-model-with-random-effect-lme4-lmertest for more info if you want. 
JarFin<- lm(JarSatCalcite_SW~JarTrt, data=CalJar)
#check assumptions 
par(mfcol=c(2,2))
plot(JarFin)
par(mfcol=c(1,1))
acf(resid(JarFin))
summary(JarFin)
#Conclusion: JarSeatable doesn't need to be accounted for as a random effect in terms of water chem. 
```

Test for significant effect of parent treatment on egg size. Start with measured water chem
```{r EggSizeStat}
CalEggs<- subset(EggsB1, SatCalcite_ATDIC != "NA")
CalEggs$TrtTankID<-droplevels(CalEggs)$TrtTankID
#start with models that have the response variable EggDiamum
eggdiam1<- lmer(EggDiamum~SatCalcite_ATDIC*AdultLength + (1|FemaleID)  + (1|AccTankID), data=CalEggs)
eggstep<-step(eggdiam1)
print(eggstep)
#final model chosen: EggDiamum~ 1 + (1|FemaleID)
eggdiamfin<- lmer(EggDiamum~1+(1|FemaleID), data=CalEggs)
#Check assumptions
par(mfcol=c(1,1))
qqnorm(resid(eggdiamfin))
qqline(resid(eggdiamfin))
plot(eggdiamfin)
acf(resid(eggdiamfin))
#the above looks fine, seems to meet all assumptions. 
summary(eggdiamfin)
#Conclusion: Female matters but not treatment or adult length. Acclimation tank and treatment tank also don't matter.  
plot(EggDiamum~FemaleID, data=CalEggs) #we need to account for FemaleID, but not egg size in Larvae models. Can also make parallel models one with EggDiamum as a covariate and one with Female ID as a random effect

#also test eccentricity
eggperimdiam1<- lmer(EggPerimDiamRat~SatCalcite_ATDIC*AdultLength + (1|FemaleID)  + (1|AccTankID), data=CalEggs)
eggperimdiamstep<-step(eggperimdiam1)
print(eggperimdiamstep)
eggperimdiamfin<- lmer(log(EggPerimDiamRat)~(1|FemaleID), data=CalEggs)
par(mfcol=c(1,1))
qqnorm(resid(eggperimdiamfin))
qqline(resid(eggperimdiamfin))
plot(eggperimdiamfin)
acf(resid(eggperimdiamfin))
summary(eggperimdiamfin)
plot(EggPerimDiamRat~FemaleID, data=CalEggs)
```

```{r EggSizeStatFixedEffect, include=FALSE}
eggdiam1f<- lmer(EggDiamum~ParentTrt*AdultLength + (1|FemaleID) + (1|TrtTankID)+ (1|AccTankID), data=CalEggs)
eggstepf<-step(eggdiam1f)
print(eggstepf)
#final model chosen: EggDiamum~ 1 + (1|FemaleID), which is the same as the eggdiamfin model
summary(eggdiamfin)
```

Choosing mean or median for larvae length data
```{r LarvaeDiamMeanOrMedian}
ggplot(LarvaeDat, aes(x=LarvaeDiamum, color=JarTrt))+
  geom_histogram(alpha=0.5, position="identity") +
  facet_grid(~ParentTrt)
larvmeans<- aggregate(LarvaeDiamum~ParentTrt*JarTrt, data=LarvaeDat, FUN=mean)
larvmedians<- aggregate(LarvaeDiamum~ParentTrt*JarTrt, data=LarvaeDat, FUN=median)
view(larvmeans)
view(larvmedians)
#I think we can use the mean for these data 

#look for the number of jars in each treatment
larvjarn<- aggregate(LarvaeDiamum~ParentTrt*JarTrt, data=LarvByJarDat, FUN=length)
```

Look at the larvae diameter data
Use fixed effect for adult treatment and larvae treatment as covariate
```{r LarvDiamStat}
larvlen<-lmer(LarvaeDiamum~ParentTrt*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarID)+(1|JarSeatable)+(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
steppedlarvlen<-step(larvlen)
print(steppedlarvlen)
#final model chosen by the lme4 step function: y~ParentTrt*JarSatCalcite_SW +(1|FemaleID)+(1|MaleID)+(1|JarID)+ (1|JarSeatable)
larvlenfin<- lmer(LarvaeDiamum~ParentTrt*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|JarID)+ (1|JarSeatable), data=LarvaeDat)

#check the final model to see that it meets assumptions
plot(larvlenfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvlenfin))
qqline(resid(larvlenfin)) #has relatively long tails, I think this should be acceptable
acf(resid(larvlenfin)) #this looks good 
summary(larvlenfin)
```

Analyze larvae growth with adult trt as fixed effect and jar as covariate
```{r LarvGrowthStat}
larvgrow<-lmer(GrowthPerDay~ParentTrt*JarSatCalcite_SW+(1|MaleID)+(1|CrossID)+(1|JarID) +(1|JarSeatable)+(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
steppedlarvgrow<- step(larvgrow)
print(steppedlarvgrow)
#final model chosen by the lme4 step function: y~SatCalcite_ATDIC*JarTrt+ (1|CrossID)+ (1|JarID) + (1|JarSeatable)
larvgrowfin<- lmer(GrowthPerDay~ParentTrt*JarSatCalcite_SW+(1|CrossID)+ (1|JarID) + (1|JarSeatable), data=LarvaeDat)
#check the final model to see that it meets assumptions
plot(larvgrowfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvgrowfin))
qqline(resid(larvgrowfin)) #has relatively long tails, based on scale I think it reasonably meets assumption of normality. 
acf(resid(larvgrowfin))
summary(larvgrowfin)

ggplot(LarvaeDat, aes(x=GrowthPerDay,color=JarTrt))+
  geom_histogram(alpha=0.5, position="identity")+
  facet_grid(ParentTrt~.)
ProblemCross<- subset(LarvaeDat, CrossID=="CF02_CM02_B1_QG")
ggplot(ProblemCross, aes(y=GrowthPerDay,x=JarID, color=JarTrt))+
  geom_boxplot(alpha=0.5, position="identity")
```

Look at larvae area; Parent Trt as fixed effect and Jar Treatment as covariate
```{r SurfaceAreaStat}
larvareat<-lmer(sqrt(LarvaeAreaum2)~ParentTrt*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|TrtTankID)+(1|MaleTrtTankID) , data=LarvaeDat)
steppedlarvareat<-step(larvareat)
print(steppedlarvareat)
#final model chosen by the lme4 step function: y~SatCalcite_ATDIC*JarTrt+(1|FemaleID)+(1|MaleID)+(1|JarID)+(1|JarSeatable)
larvareafint<- lmer(sqrt(LarvaeAreaum2)~ParentTrt*JarSatCalcite_SW+(1|FemaleID)+(1|MaleID)+(1|JarID)+(1|JarSeatable), data=LarvaeDat)
plot(larvareafint) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvareafint))
qqline(resid(larvareafint)) #this seems to have improved it
acf(resid(larvareafint)) #this looks fine to me. 
summary(larvareafint)
```

Characterize the larvae by the ratio of major and minor axis
```{r LarvAxesStat}
larvrat<-lmer(MajMinRat~ParentTrt*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)+(1|AccTankID) +(1|CrossID)++(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
larvratstep<- step(larvrat)
print(larvratstep)
larvratfin<- lmer(MajMinRat~ParentTrt*JarSatCalcite_SW +(1|FemaleID)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larvratfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvratfin))
qqline(resid(larvratfin)) #has relatively long tails, doesn't seem to meet assumption of normality. try a transformation
acf(resid(larvratfin)) #this looks fine.
summary(larvratfin)
```

Look at the ratio of perimeter to diameter with Parent as a covariate and JarTrt as a fixed effect. For a circle, this would be equal to pi
```{r PerimeterToLengthStat}
PerDiRat<- lmer((LarvaePerimeterum/LarvaeDiamum)~ParentTrt*JarSatCalcite_SW+(1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID) +(1|CrossID)+(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
PerDiStep<- step(PerDiRat)
print(PerDiStep)
#final model chosen PerimDiamRat ~ SatCalcite_ATDIC * JarTrt + (1 | FemaleID) + (1 | JarID)
PerDiRatFin<- lmer((LarvaePerimeterum/LarvaeDiamum) ~ ParentTrt*JarSatCalcite_SW + (1 | FemaleID) + (1 | JarID)+(1|JarSeatable), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(PerDiRatFin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(PerDiRatFin))
qqline(resid(PerDiRatFin)) #this looks fine
acf(resid(PerDiRatFin))
summary(PerDiRatFin)
```

Analyze eccentricity with parent as a covariate and jar as fixed effect
```{r EccentricityStat}
larveccen<-lmer(LarvaeEccentricity^3~ParentTrt*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
steppedlarveccen<- step(larveccen)
print(steppedlarveccen)
#final model
larveccenfin<- lmer(LarvaeEccentricity^3~ParentTrt*JarSatCalcite_SW+ (1|FemaleID)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larveccenfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larveccenfin))
qqline(resid(larveccenfin)) #has relatively long tails, doesn't meet assumption of normality. 
acf(resid(larveccenfin))
summary(larveccenfin)
```

Test Larvae perimeter
```{r PerimeterStat}
larvperim<-lmer(LarvaePerimeterum~ParentTrt*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|CrossID)+(1|JarSeatable)+(1|JarID)+(1|TrtTankID)+(1|MaleTrtTankID), data=LarvaeDat)
steppedlarvperim<- step(larvperim)
print(steppedlarvperim)
#final model chosen by the step function: LarvaePerimeterum~SatCalcite_ATDIC*JarTrt+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID)
larvperimfin<- lmer(LarvaePerimeterum~ParentTrt*JarSatCalcite_SW+ (1|FemaleID)+(1|MaleID)+(1|JarSeatable)+(1|JarID), data=LarvaeDat)
#check the model to see that it meets assumptions
plot(larvperimfin) #seems to meet assumption of linearity and homoscedasticity
qqnorm(resid(larvperimfin))
qqline(resid(larvperimfin)) #has relatively long tails, doesn't meet assumption of normality. 
acf(resid(larvperimfin)) #this looks fine to me.
summary(larvperimfin)
```

Analyze survival data with parent as a fixed effect
```{r SurvivalStat}
surv1<- lmer(RatSurv~ParentTrt+ (1|FemaleID)+(1|MaleID)+(1|TrtTankID)+(1|MaleTrtTankID), data=SurvWideDat)
plot(surv1)
qqnorm(resid(surv1))
qqline(resid(surv1))
acf(resid(surv1))
survstepFE<- step(surv1)
print(survstepFE)
survfin<- lmer(RatSurv~ (1|FemaleID), data=SurvWideDat)#final model chosen by step function: RatSurv~ (1|FemaleID)
plot(survfin)
qqnorm(resid(survfin))
qqline(resid(survfin))
acf(resid(survfin))
summary(survfin)

#try to see if this is significantly different for exposed vs control
SurvOA<- subset(SurvWideDat, ParentTrt=="2600")
SurvCon<- subset(SurvWideDat, ParentTrt=="400")
t.test(SurvOA$RatSurv, mu = 1, alternative = "two.sided")
t.test(SurvCon$RatSurv, mu = 1, alternative = "two.sided")

```

Final Models to Use
```{r FinalModelsWithSameRandomEffects}
summary(larvgrowfin)
summary(larvareafint)
summary(larvlenfin)
summary(larvperimfin)
```

```{r LegendEnhanced functioin, include=FALSE}
legend.enhanced <- function (x, y = NULL, legend, fill = NULL, col = par("col"), 
          border = "black", lty, lwd, pch, angle = 45, density = NULL, 
          bty = "o", bg = par("bg"), box.lwd = par("lwd"), box.lty = par("lty"), 
          box.col = par("fg"), pt.bg = NA, cex = 1, pt.cex = cex, 
          pt.lwd = lwd, xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1, 
          adj = c(0, 0.5), text.width = NULL, text.col = par("col"), 
          text.font = NULL, merge = do.lines && has.pch, trace = FALSE, 
          plot = TRUE, ncol = 1, horiz = FALSE, title = NULL, inset = 0, 
          xpd, title.col = text.col, title.adj = c(0.5,0), seg.len = 2) 
{
  if (missing(legend) && !missing(y) && (is.character(y) || 
                                         is.expression(y))) {
    legend <- y
    y <- NULL
  }
  mfill <- !missing(fill) || !missing(density)
  if (!missing(xpd)) {
    op <- par("xpd")
    on.exit(par(xpd = op))
    par(xpd = xpd)
  }
  title <- as.graphicsAnnot(title)
  if (length(title) > 1) 
    stop("invalid 'title'")
  legend <- as.graphicsAnnot(legend)
  n.leg <- if (is.call(legend)) 
    1
  else length(legend)
  if (n.leg == 0) 
    stop("'legend' is of length 0")
  auto <- if (is.character(x)) 
    match.arg(x, c("bottomright", "bottom", "bottomleft", 
                   "left", "topleft", "top", "topright", "right", "center"))
  else NA
  if (is.na(auto)) {
    xy <- xy.coords(x, y)
    x <- xy$x
    y <- xy$y
    nx <- length(x)
    if (nx < 1 || nx > 2) 
      stop("invalid coordinate lengths")
  }
  else nx <- 0
  xlog <- par("xlog")
  ylog <- par("ylog")
  rect2 <- function(left, top, dx, dy, density = NULL, angle, 
                    ...) {
    r <- left + dx
    if (xlog) {
      left <- 10^left
      r <- 10^r
    }
    b <- top - dy
    if (ylog) {
      top <- 10^top
      b <- 10^b
    }
    rect(left, top, r, b, angle = angle, density = density, 
         ...)
  }
  segments2 <- function(x1, y1, dx, dy, ...) {
    x2 <- x1 + dx
    if (xlog) {
      x1 <- 10^x1
      x2 <- 10^x2
    }
    y2 <- y1 + dy
    if (ylog) {
      y1 <- 10^y1
      y2 <- 10^y2
    }
    segments(x1, y1, x2, y2, ...)
  }
  points2 <- function(x, y, ...) {
    if (xlog) 
      x <- 10^x
    if (ylog) 
      y <- 10^y
    points(x, y, ...)
  }
  text2 <- function(x, y, ...) {
    if (xlog) 
      x <- 10^x
    if (ylog) 
      y <- 10^y
    text(x, y, ...)
  }
  if (trace) 
    catn <- function(...) do.call("cat", c(lapply(list(...), 
                                                  formatC), list("\n")))
  cin <- par("cin")
  Cex <- cex * par("cex")
  if (is.null(text.width)) 
    text.width <- max(abs(strwidth(legend, units = "user", 
                                   cex = cex, font = text.font)))
  else if (!is.numeric(text.width) || text.width < 0) 
    stop("'text.width' must be numeric, >= 0")
  xc <- Cex * xinch(cin[1L], warn.log = FALSE)
  yc <- Cex * yinch(cin[2L], warn.log = FALSE)
  if (xc < 0) 
    text.width <- -text.width
  xchar <- xc
  xextra <- 0
  yextra <- yc * (y.intersp - 1)
  ymax <- yc * max(1, strheight(legend, units = "user", cex = cex)/yc)
  ychar <- yextra + ymax
  if (trace) 
    catn("  xchar=", xchar, "; (yextra,ychar)=", c(yextra, 
                                                   ychar))
  if (mfill) {
    xbox <- xc * 0.8
    ybox <- yc * 0.5
    dx.fill <- xbox
  }
  do.lines <- (!missing(lty) && (is.character(lty) || any(lty > 
                                                            0))) || !missing(lwd)
  n.legpercol <- if (horiz) {
    if (ncol != 1) 
      warning(gettextf("horizontal specification overrides: Number of columns := %d", 
                       n.leg), domain = NA)
    ncol <- n.leg
    1
  }
  else ceiling(n.leg/ncol)
  has.pch <- !missing(pch) && length(pch) > 0
  if (do.lines) {
    x.off <- if (merge) 
      -0.7
    else 0
  }
  else if (merge) 
    warning("'merge = TRUE' has no effect when no line segments are drawn")
  if (has.pch) {
    if (is.character(pch) && !is.na(pch[1L]) && nchar(pch[1L], 
                                                      type = "c") > 1) {
      if (length(pch) > 1) 
        warning("not using pch[2..] since pch[1L] has multiple chars")
      np <- nchar(pch[1L], type = "c")
      pch <- substr(rep.int(pch[1L], np), 1L:np, 1L:np)
    }
    if (!is.character(pch)) 
      pch <- as.integer(pch)
  }
  if (is.na(auto)) {
    if (xlog) 
      x <- log10(x)
    if (ylog) 
      y <- log10(y)
  }
  if (nx == 2) {
    x <- sort(x)
    y <- sort(y)
    left <- x[1L]
    top <- y[2L]
    w <- diff(x)
    h <- diff(y)
    w0 <- w/ncol
    x <- mean(x)
    y <- mean(y)
    if (missing(xjust)) 
      xjust <- 0.5
    if (missing(yjust)) 
      yjust <- 0.5
  }
  else {
    h <- (n.legpercol + (!is.null(title))) * ychar + yc
    w0 <- text.width + (x.intersp + 1) * xchar
    if (mfill) 
      w0 <- w0 + dx.fill
    if (do.lines) 
      w0 <- w0 + (seg.len + x.off) * xchar
    w <- ncol * w0 + 0.5 * xchar
    if (!is.null(title) && (abs(tw <- strwidth(title, units = "user", 
                                               cex = cex) + 0.5 * xchar)) > abs(w)) {
      xextra <- (tw - w)/2
      w <- tw
    }
    if (is.na(auto)) {
      left <- x - xjust * w
      top <- y + (1 - yjust) * h
    }
    else {
      usr <- par("usr")
      inset <- rep_len(inset, 2)
      insetx <- inset[1L] * (usr[2L] - usr[1L])
      left <- switch(auto, bottomright = , topright = , 
                     right = usr[2L] - w - insetx, bottomleft = , 
                     left = , topleft = usr[1L] + insetx, bottom = , 
                     top = , center = (usr[1L] + usr[2L] - w)/2)
      insety <- inset[2L] * (usr[4L] - usr[3L])
      top <- switch(auto, bottomright = , bottom = , bottomleft = usr[3L] + 
                      h + insety, topleft = , top = , topright = usr[4L] - 
                      insety, left = , right = , center = (usr[3L] + 
                                                             usr[4L] + h)/2)
    }
  }
  if (plot && bty != "n") {
    if (trace) 
      catn("  rect2(", left, ",", top, ", w=", w, ", h=", 
           h, ", ...)", sep = "")
    rect2(left, top, dx = w, dy = h, col = bg, density = NULL, 
          lwd = box.lwd, lty = box.lty, border = box.col)
  }
  xt <- left + xchar + xextra + (w0 * rep.int(0:(ncol - 1), 
                                              rep.int(n.legpercol, ncol)))[1L:n.leg]
  yt <- top - 0.5 * yextra - ymax - (rep.int(1L:n.legpercol, 
                                             ncol)[1L:n.leg] - 1 + (!is.null(title))) * ychar
  if (mfill) {
    if (plot) {
      if (!is.null(fill)) 
        fill <- rep_len(fill, n.leg)
      rect2(left = xt, top = yt + ybox/2, dx = xbox, dy = ybox, 
            col = fill, density = density, angle = angle, 
            border = border)
    }
    xt <- xt + dx.fill
  }
  if (plot && (has.pch || do.lines)) 
    col <- rep_len(col, n.leg)
  if (missing(lwd) || is.null(lwd)) 
    lwd <- par("lwd")
  if (do.lines) {
    if (missing(lty) || is.null(lty)) 
      lty <- 1
    lty <- rep_len(lty, n.leg)
    lwd <- rep_len(lwd, n.leg)
    ok.l <- !is.na(lty) & (is.character(lty) | lty > 0) & 
      !is.na(lwd)
    if (trace) 
      catn("  segments2(", xt[ok.l] + x.off * xchar, ",", 
           yt[ok.l], ", dx=", seg.len * xchar, ", dy=0, ...)")
    if (plot) 
      segments2(xt[ok.l] + x.off * xchar, yt[ok.l], dx = seg.len * 
                  xchar, dy = 0, lty = lty[ok.l], lwd = lwd[ok.l], 
                col = col[ok.l])
    xt <- xt + (seg.len + x.off) * xchar
  }
  if (has.pch) {
    pch <- rep_len(pch, n.leg)
    pt.bg <- rep_len(pt.bg, n.leg)
    pt.cex <- rep_len(pt.cex, n.leg)
    pt.lwd <- rep_len(pt.lwd, n.leg)
    ok <- !is.na(pch)
    if (!is.character(pch)) {
      ok <- ok & (pch >= 0 | pch <= -32)
    }
    else {
      ok <- ok & nzchar(pch)
    }
    x1 <- (if (merge && do.lines) 
      xt - (seg.len/2) * xchar
      else xt)[ok]
    y1 <- yt[ok]
    if (trace) 
      catn("  points2(", x1, ",", y1, ", pch=", pch[ok], 
           ", ...)")
    if (plot) 
      points2(x1, y1, pch = pch[ok], col = col[ok], cex = pt.cex[ok], 
              bg = pt.bg[ok], lwd = pt.lwd[ok])
  }
  xt <- xt + x.intersp * xchar
  if (plot) {
    if (is.na(title.adj[2])) { title.adj[2] <- 0}
    if (!is.null(title)) 
      text2(left + w * title.adj[1], top - ymax, labels = title, 
            adj = c(title.adj[1], title.adj[2]), cex = cex, col = title.col)
    text2(xt, yt, labels = legend, adj = adj, cex = cex, 
          col = text.col, font = text.font)
  }
  invisible(list(rect = list(w = w, h = h, left = left, top = top), 
                 text = list(x = xt, y = yt)))
}
```

Create multipanel figure
```{r MultipanelFigure}
par(mfcol=c(1,1), mar=c(4,4.3, .75,1))
p1=visreg(larvgrowfin, "JarSatCalcite_SW", by="ParentTrt", overlay=TRUE, ylab=expression(paste("Growth (", mu, "m/day)")), xlab=expression(paste("Larvae Treatment ", Omega[Calcite])), legend=FALSE)
par(mfcol=c(2,2))
p2<-visreg(larvlenfin, "JarSatCalcite_SW", by="ParentTrt", overlay=TRUE, ylab=expression(paste("Length (", mu, "m)")), xlab="", legend=FALSE)
p3<-visreg(larvareafint, "JarSatCalcite_SW", by="ParentTrt", overlay=TRUE, ylab=expression(sqrt(paste("Area (", mu, m^2,")"))), xlab=expression(paste("Larvae Treatment ", Omega[Calcite])), legend=FALSE)
p4<-visreg(larvperimfin, "JarSatCalcite_SW", by="ParentTrt", overlay=TRUE, ylab=expression(paste("Perimeter (", mu, "m)")), xlab="", legend=FALSE)
p5<-visreg(PerDiRatFin, "JarSatCalcite_SW", by="ParentTrt", overlay=TRUE, ylab="Perimeter :Length", xlab=expression(paste("Larvae Treatment ", Omega[Calcite])), legend=FALSE)

```

Visualize larvae survival
```{r SurvivalVisualization}
ggplot(data = meansurv, aes(x = JarSatCalcite_SW, y = LarvaeSurvived, group=ParentTrt, fill=ParentTrt, color=ParentTrt)) +
  geom_point(aes(colour=ParentTrt), size=2)+
  geom_ribbon( aes(ymin = survlwrBT, ymax = survuprBT, fill = ParentTrt, color = NULL), alpha = .15) +
     geom_line( aes(y = survfitBT, color=ParentTrt), size = 1)+
  geom_errorbar(aes(ymin= LarvaeSurvived-sesurv$LarvaeSurvived, ymax=LarvaeSurvived+sesurv$LarvaeSurvived),width=0)+
   geom_errorbarh(aes(xmin= JarSatCalcite_SW-sesurv$JarSatCalcite_SW, xmax=JarSatCalcite_SW+sesurv$JarSatCalcite_SW),height=0, position=position_dodge(0))+
  labs(x = "Larvae Jar Saturation State (Calcite)", y = "Larvae Survived") +
  scale_color_manual(values = c("red2","skyblue2"), name="Parent Treatment", labels= c("OA Exposed", "Control"))+
  scale_fill_manual(values = c("red2","skyblue2"), name="Parent Treatment", labels= c("OA Exposed", "Control"))+
theme_classic()
ggplot(data = meansurvsub, aes(x = JarSatCalcite_SW, y = LarvaeSurvived)) +
  geom_point(size=2)+
  geom_ribbon( aes(ymin = survlwrBT, ymax = survuprBT, color = NULL), alpha = .15) +
     geom_line( aes(y = survfitBT), size = 1)+
  ylim(c(200,1400))+
  geom_errorbar(aes(ymin= LarvaeSurvived-sesurvsub$LarvaeSurvived, ymax=LarvaeSurvived+sesurvsub$LarvaeSurvived),width=0)+
   geom_errorbarh(aes(xmin= JarSatCalcite_SW-sesurvsub$JarSatCalcite_SW, xmax=JarSatCalcite_SW+sesurvsub$JarSatCalcite_SW),height=0, position=position_dodge(0))+
  labs(x =expression(paste(Omega[Calcite])),
                  y = expression(paste("Number of Surviving Larvae")))+
theme_classic()
#this figure doesn't account for parent treatment because it wasn't significant

boxplot(RatSurv~as.factor(ParentTrt), data=SurvWideDat)
boxplot(SurvChange~as.factor(ParentTrt), data= SurvWideDat)
#make reaction norms for survival color will be parent treatment. X will be jar treatment. Each cross will be its own line
ggplot(SurvDat,aes(x=JarTrt, y=LarvaeSurvived, color = ParentTrt, group= CrossID)) +
    geom_point(aes(color=ParentTrt))+
  geom_line(aes(color=ParentTrt, group=CrossID))+
  geom_errorbar(aes(ymin= LarvaeSurvived-SELarvaeSurvived, ymax=LarvaeSurvived+SELarvaeSurvived),width=0.05, position=position_dodge(0))+
    labs(x = "Larvae Treatment", y = "Number of Larvae Survived") +
  scale_color_manual(values = c("skyblue2", "red2")) +
  theme_classic()
ggplot(SurvWideDat,
       aes(y = SurvChange, x = ParentTrt)) +
  geom_point() +
  ggtitle("Larvae survival by parent treatment")+ 
  ylab("Proportion change in the number of larvae that survived")+
  theme_classic()
meansurv2<- ddply(SurvWideDat, .(ParentTrt), numcolwise(mean, na.rm=T))
sesurv<- ddply(SurvWideDat, .(ParentTrt), numcolwise(se, na.rm=T))
#make a shorter name for the CrossIDs in SurvWideDat
SurvWideDat$Cross<- substr(SurvWideDat$CrossID, 1, 9)
ggplot(data = SurvWideDat, aes(x = Cross, y = RatSurv, fill=ParentTrt)) +
  geom_bar(stat="identity", position="dodge")+
  labs(x = "Cross ID", y = "Proportion of Larvae that Survived") +
  scale_fill_manual(name = "Parent Treatment", labels = c("Control", "OA Exposed"), values = c("skyblue2", "red2")) +
  theme(axis.text.x=element_text(angle=90))
ggplot(data = meansurv2, aes(x = ParentTrt, y = RatSurv)) +
  geom_bar(stat="identity", position="dodge")+
  labs(x = "Parent Treatment", y = "Proportion of Larvae that Survived")+
  geom_errorbar(aes(ymin= RatSurv-sesurv$RatSurv, ymax=RatSurv+sesurv$RatSurv),width=0.2, position=position_dodge(.9))+
  theme_classic()
```

```{r GetDescriptiveStats}
#get means for the data
meanJarTrt<- ddply(LarvaeDat, .(JarTrt, ParentTrt), numcolwise(mean, na.rm=T))
seJarTrt<- ddply(LarvaeDat, .(JarTrt, ParentTrt), numcolwise(se, na.rm=T))
mgpd<- subset(meanJarTrt, select= c(JarTrt, ParentTrt, GrowthPerDay, LarvaeDiamum, LarvaeAreaum2, LarvaePerimeterum))
(mgpd$GrowthPerDay[3]-mgpd$GrowthPerDay[4])/mgpd$GrowthPerDay[4]
(mgpd$GrowthPerDay[1]-mgpd$GrowthPerDay[2])/mgpd$GrowthPerDay[2]
(mgpd$LarvaeDiamum[3]-mgpd$LarvaeDiamum[4])/mgpd$LarvaeDiamum[4]
(mgpd$LarvaeDiamum[1]-mgpd$LarvaeDiamum[2])/mgpd$LarvaeDiamum[2]
mean(mgpd$LarvaeDiamum[1:2])/mean(mgpd$LarvaeDiamum[3:4])
(mgpd$LarvaeAreaum2[3]-mgpd$LarvaeAreaum2[4])/mgpd$LarvaeAreaum2[4]
(mgpd$LarvaeAreaum2[1]-mgpd$LarvaeAreaum2[2])/mgpd$LarvaeAreaum2[2]
mean(mgpd$LarvaeAreaum2[1:2])/mean(mgpd$LarvaeAreaum2[3:4])

(mgpd$LarvaePerimeterum[3]-mgpd$LarvaePerimeterum[4])/mgpd$LarvaePerimeterum[4]
(mgpd$LarvaePerimeterum[1]-mgpd$LarvaePerimeterum[2])/mgpd$LarvaePerimeterum[2]
mean(mgpd$LarvaePerimeterum[1:2])/mean(mgpd$LarvaePerimeterum[3:4])

mean(SurvWideDat$RatSurv, na.rm=TRUE)
```

Export Figures
```{r ExportFigures, include=FALSE}
plots.dir.path<- list.files(tempdir(), pattern="rs-graphics", full.names= TRUE)
plots.png.paths<- list.files(plots.dir.path, pattern=".png", full.names=TRUE)
file.copy(from=plots.png.paths, to="../results")
```

Do simple t-tests on the means for oa-exposed and control exposed parents within each larvae treatment
```{r t-tests}
#get Bonferoni adjusted p-crit: 
0.05/10

Lexp<- subset(LarvaeDat, JarTrt=="Exposed")
Lcon<- subset(LarvaeDat, JarTrt=="Control")

t.test(GrowthPerDay~ParentTrt, data=Lexp)
t.test(GrowthPerDay~ParentTrt, data=Lcon)

t.test(LarvaeDiamum~ParentTrt, data=Lexp)
t.test(LarvaeDiamum~ParentTrt, data=Lcon)#not significantly different for control larvae

t.test(LarvaeAreaum2~ParentTrt, data=Lexp)
t.test(LarvaeAreaum2~ParentTrt, data=Lcon) #not signficantly different for control larvae

t.test(LarvaePerimeterum~ParentTrt, data=Lexp)
t.test(LarvaePerimeterum~ParentTrt, data=Lcon) #not significantly different for control larvae

t.test(PerimLenRat~ParentTrt, data=Lexp)
t.test(PerimLenRat~ParentTrt, data=Lcon)#not significantly different for control larvae
```
