---
title: "McNally_etal_Supplemental_Code"
author: "Elise McNally"
date: "11/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r LoadLibraries, include=FALSE}
library(knitr)
library(cowplot)
library(plyr)
library(tidyverse)
library(sciplot)
library(reshape2)
library(lme4)
library(lmerTest)
library(blme)
library(grid)
library(data.table)
library(gridExtra)
library(lattice)
library(optimx)
library(car)
library(Hmisc)
library(DescTools)
library(corrplot)
library(emmeans)
library(chron)
library(GLMMadaptive)
library(calibrate)
library(merTools)
library(bootpredictlme4)
library(visreg)
library(lubridate)
library(marelac)
```

# Load Data
```{r Load Data}
setwd("~/R/GitHub/LarvaeTransGen2018/data") #Elise's working directory
BARN<- read.csv("Barn_Botts.csv")
LoggerRawData<- read.csv("BARN_pHLogger.csv")
LoggerCalib<- read.csv("BARN_pHLoggerCalib.csv")
CondRawData<- read.csv("BARN_CondLogger.csv")
KnownSals<- read.csv("BARN_KnownSals.csv")

```

# Work on getting salinity from conductivity logger
We measure salinity directly from water samples taken at the time of launching the logger. To correct for drift we need to: 
1) ID times on the logger that correspond to times that salinity was measured
2) Convert known salinities to conductivity
3) Get drift of the conductivity logger between salinity measurement points
4) Convert measured conductivity to actual conductivity using the drift correction
5) convert the corrected conductivity to salinity
6) Match the logged salinity values to the pH logger data

The function convert_StoR provides a salinity to conductivity ratio conversion. Equation is valid for salinity 2-42 in the world ocean. Here is the usage for the code: 
convert_StoR(S= SALINITYppt, t= TEMPCels, p=PRESSURE, P=applied pressure)
1bar= seasurface

The conductivity ratio is the measured conductivity divided by the conductivity of 35ppt seawater at 15C and 0 pressure, which is 42.914 mS/cm = 42914 uS/cm

Load libraries
```{r}
library(marelac)
library(lubridate)
```

Load data
```{r}
#set working directory
setwd("~/R/GitHub/FieldSamples/data")
sal<- read.csv("KnownSal.csv")
```

Get conductivity from salinity
```{r}
#Create foreloop to get the ratios for each of the bottles that were collected
for(i in 1:length(sal$Sal)){
  sal$CondRatio[i]<- convert_StoR(S=sal$Sal[i], t= sal$TempC[i], p=sal$Pressure[i])
}

sal$Conductivity<- 42914*sal$CondRatio
#now use that conductivity to correct the logger data. 

write.csv(sal, "~/R/GitHub/FieldSamples/data/KnownCond.csv")

```


Code from Adrienne to get salinities
```{r}
#if salinity is not converted from the HydroCAT here is the formula to do it
library(oce)

test$cond=test$`Spec Conductivity`*0.001

test$Salinity=swSCTp(test$cond, T90fromT68(test$Temperature), pressure = (test$Pressure), conductivityUnit = "mS/cm", eos = "unesco")
```




# Work on getting pH to the NBS and Total scales accounting for liquid junction effects

1) ID times that the CRM was measured; noting if Cu was on or off
2) Get measured pH from the logger
3) Use CO2SYS to calculate the certified pH of the CRM on the NBS scale for the measured temperature
4) Calculate the correction factor for liquid junction effects (Cert-Meas pH)/Crm salinity
5) Remove correction factor outliers
6) Get the average correction factor per unit of salinity
7) Add the LJCorr factor for each electrode to the LoggerRawData 
8) Get the correction to the total scale (not salinity dependent)
9) Use the LJCorr and total correction along with the salinity measurements to get the pH data to the NBSD and total scales. 



#Subset data
```{r Subset to only include May-July}
#Make dates in the date format for R
BARN$Date<- ymd(BARN$Date)
LoggerRawData$DateTimeGMT<- ymd_hms(LoggerRawData$DateTimeGMT)
Logger<- LoggerRawData
#get the months
BARN$Month<- month(BARN$Date)
Logger$Month<- month(Logger$DateTimeGMT)
Logger$Day<- day(Logger$DateTimeGMT) #make a simple month and day field
Logger$DayMonth<- as.factor(paste(Logger$Month, Logger$Day, sep="-"))
Logger$Year<- year(Logger$DateTimeGMT)
#subset to only include may-july and only tidal samples for bottles
BARN$SampType<- as.factor(BARN$SampType)
BARNsub<- subset(BARN, SampType=="Tidal" & Month!=10)
LoggerSub<- subset(Logger, Year==2018 | Year== 2019) %>% 
  subset(Month == 5 |Month== 6 | Month == 7 | Month == 8)

LoggerSub$SimpleDate<- date(LoggerSub$DateTimeGMT)
LoggerSub$pH<- as.numeric(LoggerSub$pH)
LogAgg<- aggregate(pH ~ SimpleDate, data=LoggerSub, FUN= mean)
plot(pH~SimpleDate, data=LogAgg)

```

# Make figures using bottle data
```{r Bottle Figures}
dev.off()
#round relative tide
BARNsub$RelTideRounded<- round(BARNsub$RelTide)
BottMeans<- ddply(BARNsub, .(RelTideRounded), numcolwise(mean, na.rm=TRUE))
BottSes<- ddply(BARNsub, .(RelTideRounded), numcolwise(se, na.rm=TRUE))

plot(OmegaAr~RelTideRounded, data=BottMeans, pch=19, xlab="Relative Tide Time", ylab= expression(paste( Omega[Aragonite])), ylim=c(0, 2.0))
abline(b=0, a=1)
arrows(x0=BottMeans$RelTideRounded, x1=BottMeans$RelTideRounded,y0=BottMeans$OmegaAr-BottSes$OmegaAr, y1=BottMeans$OmegaAr+BottSes$OmegaAr, angle=90, code=3, length=0.05, col="black")

plot(pCO2~RelTideRounded, data=BottMeans, pch=19, xlab="Relative Tide Time", ylab= expression(paste(pCO[2], " (ppm)")), ylim=c(0, 3000))
arrows(x0=BottMeans$RelTideRounded, x1=BottMeans$RelTideRounded,y0=BottMeans$pCO2-BottSes$pCO2, y1=BottMeans$pCO2+BottSes$pCO2, angle=90, code=3, length=0.05, col="black")

plot(CO3~RelTideRounded, data=BottMeans, pch=19, xlab="Relative Tide Time", ylab= expression(paste("[", CO[3], "] (", mu, "m)")), ylim=c(0, 125))
arrows(x0=BottMeans$RelTideRounded, x1=BottMeans$RelTideRounded,y0=BottMeans$CO3-BottSes$CO3, y1=BottMeans$CO3+BottSes$CO3, angle=90, code=3, length=0.05, col="black")

plot(SalCorr~RelTideRounded, data=BottMeans, pch=19, xlab="Relative Tide Time", ylab="Salinity (ppt)", ylim= c(20,32))
arrows(x0=BottMeans$RelTideRounded, x1=BottMeans$RelTideRounded,y0=BottMeans$SalCorr-BottSes$SalCorr, y1=BottMeans$SalCorr+BottSes$SalCorr, angle=90, code=3, length=0.05, col="black")

```

# Make figures using pH logger data
```{r Logger figures}

LoggerSub$Yearday<- yday(LoggerSub$DateTimeGMT) #get the year day of the data
LoggerSub$pHTempF<- as.numeric(LoggerSub$pHTempF)
LoggerSub$pHTempC<- (LoggerSub$pHTempF-32)*5/9
LogMeans<- ddply(LoggerSub, .(Yearday), numcolwise(mean, na.rm=TRUE))

LogSes<- ddply(LoggerSub, .(Yearday), numcolwise(se, na.rm=TRUE))
par(mar = c(5, 4, 4, 4) + 0.3)
Plot<- plot(pH~Yearday, data=LogMeans, pch=19, type="p", xaxt="n", xlab="", ylim=c(7.4, 8.05), ylab= "pH (NBS)")
axis(1,at=LoggerSub$Yearday, labels=LoggerSub$DayMonth)
arrows(x0=LogMeans$Yearday, x1=LogMeans$Yearday, y0=LogMeans$pH - LogSes$pH, y1=LogMeans$pH + LogSes$pH, code=3, angle=90, length= 0.1)
par(new=T)
plot(pHTempC~Yearday, data=LogMeans, pch=17, type="p", axes=FALSE, bty="n", xlab="Date", ylab= "", col="blue", ylim=c(19, 27))
axis(side=4,at=pretty(range(LoggerSub$pHTempC)), col.axis="blue")
arrows(x0=LogMeans$Yearday, x1=LogMeans$Yearday, y0=LogMeans$pHTempC - LogSes$pHTempC, y1=LogMeans$pHTempC + LogSes$pHTempC, code=3, angle=90, length= 0.1, col="blue")
mtext("Temperature (C)", side= 4, line=3, col="blue")

```


