---
title: "VisualizationLarvae"
author: "Katie Lotterhos"
date: "6/5/2019"
output: html_document
---

setwd("~/Desktop/Repos/LarvaeTransGen2018")

# Install libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
```

## Load data

```{r }
morph <- fread("data/Larvae_morphology.csv")
head(morph)
```

## Plots of overall effect

```{r }
morph$CrossID_LarTrt <- paste(morph$CrossID, morph$JarTrt, sep="_")

morph$AdultTrt <- substr(morph$CrossID, 1, 1)

par(mar = c(5, 4, 0.5, 0.5), bg = "transparent")
boxplot(morph$MaxFeretDiamum ~ morph$JarTrt * morph$AdultTrt, 
        xlim = c(0, 5),
        xlab = "Larvae Treatment",
        ylab = "Diameter (um)",
        names = c("Control", "OA", "Control", "OA"), 
        col = c("powderblue", "red"))
rect(xleft = -0.2, 
     ybottom = 0, 
     xright = 2.5, 
     ytop = 85,
     border = FALSE,
     col = rgb(0, 0, 0.5, alpha = 0.1))
rect(xleft = 2.5, 
     ybottom = 0, 
     xright = 5.2, 
     ytop = 85,
     border = FALSE,
     col = rgb(0.5, 0, 0, alpha = 0.1))
text(x = 1.25, y = 84, "Control Parents")
text(x = 3.75, y = 84, "OA Parents")
boxplot(morph$MaxFeretDiamum ~ morph$JarTrt * morph$AdultTrt, 
        xlim = c(0, 5),
        xlab = "Larvae Treatment",
        ylab = "Diameter (um)",
        names = c("Control", "OA", "Control", "OA"), 
        col = c("powderblue", "red"),
        add = TRUE)
dev.copy(png, "results/Diameter_LarvaeAdultTreatmentContrast_Boxplot.png")
dev.off()

boxplot(morph$SurfaceAreaum2~morph$AdultTrt*morph$JarTrt)

boxplot(morph$Perimeterum~morph$AdultTrt*morph$JarTrt)
```

## Plots of overall effect

Area
```{r }

par(mar=c(10,4,0.5,0.5))
boxplot(morph$SurfaceAreaum2~morph$CrossID_LarTrt, las=2)

difftab <- data.frame(tapply(morph$SurfaceAreaum2, list(morph$CrossID, morph$JarTrt), median))
difftab$AdultTrt <- substr(rownames(difftab), 1,1)
difftab$diff <- difftab$Exposed-difftab$Control

par(mar=c(4,6,0.5,0.5))
boxplot(difftab$diff~difftab$AdultTrt, xlab="Adult Treatment", names=c("Control", "OA"), ylab="Decrease in median shell area per family \n from control to OA conditions")

t.test(difftab$diff~difftab$AdultTrt)
# don't trust this p-value b/c non-independence among parents,
# use permutation to test null H0

difftab$Fem <- substr(rownames(difftab), 1, 4)
difftab$Male <- substr(rownames(difftab), 6, 9)

difftab$diff~difftab$AdultTrt

(difftab$diff~difftab$AdultTrt*difftab$Fem*difftab$Male)

plot(NULL, xlim=c(0,3), ylim=c(2500,max(difftab$Control)))
arrows(x0 = 0.5, y0=difftab$Control[difftab$AdultTrt=="C"], x1=1.5, y1=difftab$Exposed[difftab$AdultTrt=="C"], col="blue")

arrows(x0 = 1.0, y0=difftab$Control[difftab$AdultTrt=="E"], x1=2.0, y1=difftab$Exposed[difftab$AdultTrt=="E"], col="red")
```



cond <- morph$Female=="EF03"
morph_sub <- morph[cond,]
table(morph_sub$CrossID_LarTrt,morph_sub$JarID)
table(morph_sub$CrossID_JarID,morph_sub$JarTrt)


# need to rename jar ID
morph$CrossID_JarID <- paste(morph$CrossID, morph$JarID, sep="_")
par(mar=c(10,4,0.5,0.5))
boxplot(morph_sub$MaxFeretDiamum~morph_sub$CrossID_JarID*morph_sub$JarTrt, las=2)

```

## Plots of female

```{r }
morph$Female <- substr(morph$CrossID, 1, 4)
morph$Male <- substr(morph$CrossID, 6, 9)

boxplot(morph$MaxFeretDiamum~morph$Female*morph$JarTrt, las=2)
(female_means <- tapply(morph$MaxFeretDiamum, list(morph$Female, morph$AdultTrt), mean))
colMeans(female_means, na.rm=TRUE)

boxplot(morph$MaxFeretDiamum~morph$Male*morph$JarTrt, las=2)
```
